---
title: "5TGM1_Stromal_QC_Integration_StandardWorkflow"
output: html_document
date: "2023-06-01"
---
# R Markdown used to preprocess raw data, perform quality check, filter out low quality cells, integrate libraries, and perform standward workflow on 5TGM1 stromal single-cell data.

# Created 6/1/2023 by Dylan Gagler, MSc - dylan.gagler@nyulangone.org

# Load up libraries
```{r}
library(Seurat)
library(ggplot2)
library(dplyr)
library(ggridges)
library(ggforce)
library(cowplot)
library(Matrix)

options(future.globals.maxSize=1000000000000000)
```
# Read in data
```{r}
lib1.data <- Read10X('/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/082322-1/filtered_feature_bc_matrix')
lib2.data <- Read10X('/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/082322-2/filtered_feature_bc_matrix')
lib3.data <- Read10X('/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/082322-3/filtered_feature_bc_matrix')
lib5.data <- Read10X('/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/090122-1/filtered_feature_bc_matrix')
lib6.data <- Read10X('/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/090122-2/filtered_feature_bc_matrix')
lib4.data <- Read10X('/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/sixth/filtered_feature_bc_matrix')
```
# Fixing lib4 hashtag names
```{r}
lib4.data$`Antibody Capture`@Dimnames[[1]][1:8] <- c("hashtag1", "hashtag2", "hashtag3", "hashtag4",
                                                     "hashtag5", "hashtag6", "hashtag7", "hashtag8")
```
# Make Seurat object from GE assay
```{r}
lib1 <- CreateSeuratObject(counts = lib1.data$`Gene Expression`, min.cells=3, min.features=200)
lib2 <- CreateSeuratObject(counts = lib2.data$`Gene Expression`, min.cells=3, min.features=200)
lib3 <- CreateSeuratObject(counts = lib3.data$`Gene Expression`, min.cells=3, min.features=200)
lib4 <- CreateSeuratObject(counts = lib4.data$`Gene Expression`, min.cells=3, min.features=200)
lib5 <- CreateSeuratObject(counts = lib5.data$`Gene Expression`, min.cells=3, min.features=200)
lib6 <- CreateSeuratObject(counts = lib6.data$`Gene Expression`, min.cells=3, min.features=200)
```
# Fixing mismatched matrix sizes due to cells lost from above filtering
```{r}
# Lib1
lib1.diff <- setdiff(colnames(lib1.data$`Antibody Capture`), colnames(lib1)) # Getting the unmatched cell
lib1.data$`Antibody Capture` <- lib1.data$`Antibody Capture`[, !colnames(lib1.data$`Antibody Capture`) %in% lib1.diff]
# Lib2
lib2.diff <- setdiff(colnames(lib2.data$`Antibody Capture`), colnames(lib2)) # Getting the unmatched cell
lib2.data$`Antibody Capture` <- lib2.data$`Antibody Capture`[, !colnames(lib2.data$`Antibody Capture`) %in% lib2.diff]
# Lib3
lib3.diff <- setdiff(colnames(lib3.data$`Antibody Capture`), colnames(lib3)) # Getting the unmatched cell
lib3.data$`Antibody Capture` <- lib3.data$`Antibody Capture`[, !colnames(lib3.data$`Antibody Capture`) %in% lib3.diff]
# Lib4
lib4.diff <- setdiff(colnames(lib4.data$`Antibody Capture`), colnames(lib4)) # Getting the unmatched cell
lib4.data$`Antibody Capture` <- lib4.data$`Antibody Capture`[, !colnames(lib4.data$`Antibody Capture`) %in% lib4.diff]
# Lib5
lib5.diff <- setdiff(colnames(lib5.data$`Antibody Capture`), colnames(lib5)) # Getting the unmatched cell
lib5.data$`Antibody Capture` <- lib5.data$`Antibody Capture`[, !colnames(lib5.data$`Antibody Capture`) %in% lib5.diff]
# Lib6
lib6.diff <- setdiff(colnames(lib6.data$`Antibody Capture`), colnames(lib6)) # Getting the unmatched cell
lib6.data$`Antibody Capture` <- lib6.data$`Antibody Capture`[, !colnames(lib6.data$`Antibody Capture`) %in% lib6.diff]
```
# Create hashtag assay
```{r}
lib1[["HTO"]] <- CreateAssayObject(counts = lib1.data$`Antibody Capture` + 1, colnames = (x = lib1)) # needed to add 1 to make HTODemux() run
lib2[["HTO"]] <- CreateAssayObject(counts = lib2.data$`Antibody Capture` + 1, colnames = (x = lib2))
lib3[["HTO"]] <- CreateAssayObject(counts = lib3.data$`Antibody Capture` + 1, colnames = (x = lib3))
lib4[["HTO"]] <- CreateAssayObject(counts = lib4.data$`Antibody Capture` + 1, colnames = (x = lib4))
lib5[["HTO"]] <- CreateAssayObject(counts = lib5.data$`Antibody Capture` + 1, colnames = (x = lib5))
lib6[["HTO"]] <- CreateAssayObject(counts = lib6.data$`Antibody Capture` + 1, colnames = (x = lib6))
```
# Add library info
```{r}
lib1[["library"]] <- "lib1"
lib2[["library"]] <- "lib2"
lib3[["library"]] <- "lib3"
lib4[["library"]] <- "lib4"
lib5[["library"]] <- "lib5"
lib6[["library"]] <- "lib6"

lib1[["case.control"]] <- "control"
lib2[["case.control"]] <- "control"
lib3[["case.control"]] <- "myeloma"
lib4[["case.control"]] <- "myeloma"
lib5[["case.control"]] <- "control"
lib6[["case.control"]] <- "myeloma"
```
# Get %mito
```{r}
# Lib 1
lib1[["percent.mt"]] <- PercentageFeatureSet(lib1, pattern="^mt-")
# Lib 2
lib2[["percent.mt"]] <- PercentageFeatureSet(lib2, pattern="^mt-")
# Lib 3
lib3[["percent.mt"]] <- PercentageFeatureSet(lib3, pattern="^mt-")
# Lib 4
lib4[["percent.mt"]] <- PercentageFeatureSet(lib4, pattern="^mt-")
# Lib 5
lib5[["percent.mt"]] <- PercentageFeatureSet(lib5, pattern="^mt-")
# Lib 6
lib6[["percent.mt"]] <- PercentageFeatureSet(lib6, pattern="^mt-")
```
# Visualize genes per cell
```{r}
genes.for.RidgePlot <- data.frame(lib = c(rep("Lib 1", length(lib1$nFeature_RNA)),
                              rep("Lib 2", length(lib2$nFeature_RNA)),
                              rep("Lib 3", length(lib3$nFeature_RNA)),
                              rep("Lib 4", length(lib4$nFeature_RNA)),
                              rep("Lib 5", length(lib5$nFeature_RNA)),
                              rep("Lib 6", length(lib6$nFeature_RNA))),
                              
                              counts = c(lib1$nFeature_RNA,
                                        lib2$nFeature_RNA,
                                        lib3$nFeature_RNA,
                                        lib4$nFeature_RNA,
                                        lib5$nFeature_RNA,
                                        lib6$nFeature_RNA)
)

genes.for.RidgePlot$lib <- factor(genes.for.RidgePlot$lib, levels = c("Lib 1", "Lib 2", "Lib 3", "Lib 4", "Lib 5", "Lib 6"), ordered = T)

ggplot(data = genes.for.RidgePlot) +
  geom_density_ridges2(aes(x = counts, y = lib, group = lib, fill = lib, alpha=0.8)) + 
  labs(title = "Unique Genes per Cell",
       x = "Count",
       y = "Density") +
  theme(plot.title = element_text(hjust = 0.7)) +
  geom_vline(xintercept=500) +
  scale_y_discrete(expand=c(0,0)) +
  scale_x_continuous(limits = c(0,7000), expand = c(0.025,0)) +
  theme_minimal() + theme(legend.position = "none")
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/GenesPerCell_LibraryRidgePlot.png")
```
# Visualize transcripts per cell
```{r}
transcripts.for.RidgePlot <- data.frame(lib = c(rep("Lib 1", length(lib1$nCount_RNA)),
                              rep("Lib 2", length(lib2$nCount_RNA)),
                              rep("Lib 3", length(lib3$nCount_RNA)),
                              rep("Lib 4", length(lib4$nCount_RNA)),
                              rep("Lib 5", length(lib5$nCount_RNA)),
                              rep("Lib 6", length(lib6$nCount_RNA))),
                              
                              counts = c(lib1$nCount_RNA,
                                        lib2$nCount_RNA,
                                        lib3$nCount_RNA,
                                        lib4$nCount_RNA,
                                        lib5$nCount_RNA,
                                        lib6$nCount_RNA)
)

transcripts.for.RidgePlot$lib <- factor(transcripts.for.RidgePlot$lib, levels = c("Lib 1", "Lib 2", "Lib 3", "Lib 4", "Lib 5", "Lib 6"), ordered = T)

ggplot(data = transcripts.for.RidgePlot) +
  geom_density_ridges2(aes(x = counts, y = lib, group = lib, fill = lib, alpha=0.8)) + 
  labs(title = "Unique Transcripts per Cell",
       x = "Count",
       y = "Density") +
  theme(plot.title = element_text(hjust = 0.7)) +
  geom_vline(xintercept=500) +
  scale_y_discrete(expand=c(0,0)) +
  scale_x_continuous(limits = c(0,15000), expand = c(0.025,0)) +
  theme_minimal() + theme(legend.position = "none")
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/TranscriptsPerCell_LibraryRidgePlot.png")
```
# % Mito ridges
```{r}
percmito.for.RidgePlot <- data.frame(lib = c(rep("Lib 1", length(lib1$percent.mt)),
                              rep("Lib 2", length(lib2$percent.mt)),
                              rep("Lib 3", length(lib3$percent.mt)),
                              rep("Lib 4", length(lib4$percent.mt)),
                              rep("Lib 5", length(lib5$percent.mt)),
                              rep("Lib 6", length(lib6$percent.mt))),
                              
                              counts = c(lib1$percent.mt,
                                        lib2$percent.mt,
                                        lib3$percent.mt,
                                        lib4$percent.mt,
                                        lib5$percent.mt,
                                        lib6$percent.mt)
)

percmito.for.RidgePlot$lib <- factor(percmito.for.RidgePlot$lib, levels = c("Lib 1", "Lib 2", "Lib 3", "Lib 4", "Lib 5", "Lib 6"), ordered = T)

ggplot(data = percmito.for.RidgePlot) +
  geom_density_ridges2(aes(x = counts, y = lib, group = lib, fill = lib, alpha=0.8)) + 
  labs(title = "Mitochondrial Transcript % per Cell",
       x = "Count",
       y = "Density") +
  theme(plot.title = element_text(hjust = 0.7)) +
  geom_vline(xintercept=500) +
  scale_y_discrete(expand=c(0,0)) +
  scale_x_continuous(limits = c(0,10), expand = c(0.025,0)) +
  theme_minimal() + theme(legend.position = "none")
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/PercMito_LibraryRidgePlot.png")
```
# Normalizing and demultiplexing HTOs
```{r}
# Lib 1
lib1 <- NormalizeData(object = lib1,
                           assay = "HTO",
                           normalization.method = "CLR")
lib1 <- HTODemux(object = lib1,
                      assay = "HTO",
                      kfunc =  "kmeans",
                      positive.quantile = 0.999)
# Lib 2
lib2 <- NormalizeData(object = lib2,
                           assay = "HTO",
                           normalization.method = "CLR")
lib2 <- HTODemux(object = lib2,
                      assay = "HTO",
                      kfunc =  "kmeans",
                      positive.quantile = 0.999)
# Lib 3
lib3 <- NormalizeData(object = lib3,
                           assay = "HTO",
                           normalization.method = "CLR")
lib3 <- HTODemux(object = lib3,
                      assay = "HTO",
                      kfunc =  "kmeans",
                      positive.quantile = 0.999)
# Lib 4
lib4 <- NormalizeData(object = lib4,
                           assay = "HTO",
                           normalization.method = "CLR")
lib4 <- HTODemux(object = lib4,
                      assay = "HTO",
                      kfunc =  "kmeans",
                      positive.quantile = 0.999)
# Lib 5
lib5 <- NormalizeData(object = lib5,
                           assay = "HTO",
                           normalization.method = "CLR")
lib5 <- HTODemux(object = lib5,
                      assay = "HTO",
                      kfunc =  "kmeans",
                      positive.quantile = 0.999)
# Lib 6
lib6 <- NormalizeData(object = lib6,
                           assay = "HTO",
                           normalization.method = "CLR")
lib6 <- HTODemux(object = lib6,
                      assay = "HTO",
                      kfunc =  "kmeans",
                      positive.quantile = 0.999)
```
# Hashtag heatmaps
```{r}
HTOHeatmap(lib1, assay = "HTO", ncells = 5000)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Library1_HTO_Heatmap.png", width = 5, height = 3)

HTOHeatmap(lib2, assay = "HTO", ncells = 5000)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Library2_HTO_Heatmap.png", width = 5, height = 3)

HTOHeatmap(lib3, assay = "HTO", ncells = 5000)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Library3_HTO_Heatmap.png", width = 5, height = 3)

HTOHeatmap(lib4, assay = "HTO", ncells = 5000)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Library4_HTO_Heatmap.png", width = 5, height = 3)

HTOHeatmap(lib5, assay = "HTO", ncells = 5000)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Library5_HTO_Heatmap.png", width = 5, height = 3)

HTOHeatmap(lib6, assay = "HTO", ncells = 5000)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Library6_HTO_Heatmap.png", width = 5, height = 3)
```
# Hashtag pie charts - needs to be altered for each library. For some reason, the order of the classification outputs can be different between libraries. Make sure you are
# properly identifying things as singlets, doublets, etc.
```{r}
table(lib5$HTO_classification.global)
# Get values
#lib5.doublets <- as.numeric(table(lib5$HTO_classification.global)[1])
lib5.negatives <- as.numeric(table(lib5$HTO_classification.global)[1])
lib5.singlets <- as.numeric(table(lib5$HTO_classification.global)[2])
lib5.sum <- sum(table(lib5$HTO_classification.global))
lib5.values <- c(1, lib5.negatives, lib5.singlets)


# Get fracs
#lib5.doublet.frac <- lib5.doublets/lib5.sum
lib5.negative.frac <- lib5.negatives/lib5.sum
lib5.singlet.frac <- lib5.singlets/lib5.sum
lib5.fracs <- c(0.0001, lib5.negative.frac, lib5.singlet.frac)

# Get labels
#lib5.doublet.label <- paste(toString(round(lib5.doublet.frac, digits=2)*100), "%", sep="")
lib5.negative.label <- paste(toString(round(lib5.negative.frac, digits=2)*100), "%", sep="")
lib5.singlet.label <- paste(toString(round(lib5.singlet.frac, digits=2)*100), "%", sep="")
lib5.labels <- c("0", lib5.negative.label, lib5.singlet.label)

new.label <- c(0, 5649, 1411)
# 
# # Breakdown per hashtag 
table(lib5$HTO_classification)

lib5.pie <- data.frame(group=c("Doublet", "Negative", "Singlet"), value=lib5.values, #informed by HTO_classification.global table
                       perc=lib5.fracs, labels=new.label)

lib5.pie <- lib5.pie %>%
  arrange(desc(group)) %>%
  mutate(prop = value / sum(lib5.pie$value) *100) %>%
  mutate(ypos = cumsum(prop)- 0.5*prop )

ggplot(lib5.pie, aes(x = "", y = perc, fill = group)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels), color = c(1, "white", "white"),
            position = position_stack(vjust = 0.5),
            show.legend = FALSE) +
  guides(fill = guide_legend(title = "")) +
  scale_fill_viridis_d() +
  coord_polar(theta = "y") +
  ggtitle("Library 6") +
  theme_void()
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/lib5_SingletPie_Counts.png")
```
# Hashtag ridges
```{r}
RidgePlot(lib1, assay = "HTO", features = rownames(lib1[["HTO"]])[1:3], ncol = 3)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Lib1_HTO_RidgePlot.png", height = 4, width = 10)

RidgePlot(lib2, assay = "HTO", features = rownames(lib2[["HTO"]])[1:3], ncol = 3)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Lib2_HTO_RidgePlot.png", height = 4, width = 10)

RidgePlot(lib3, assay = "HTO", features = c("hashtag4", "hashtag5", "hashtag6", "hashtag8"), ncol = 4)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Lib3_HTO_RidgePlot.png", height = 4, width = 12)

RidgePlot(lib4, assay = "HTO", features = c("hashtag4", "hashtag5", "hashtag6", "hashtag8"), ncol = 4)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Lib4_HTO_RidgePlot.png", height = 4, width = 12)

RidgePlot(lib5, assay = "HTO", features = "hashtag7", ncol = 1)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Lib5_HTO_RidgePlot.png", height = 4, width = 4)

RidgePlot(lib6, assay = "HTO", features = c("hashtag8"), ncol = 1)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Lib6_HTO_RidgePlot.png", height = 4, width = 4)
```
# Filtering out low quality cells
# Library 4 removed because failed -- very few cells
```{r}
prefilter_total <- sum(length(colnames(lib1)), length(colnames(lib2)), length(colnames(lib3)), length(colnames(lib5)), length(colnames(lib6)))

lib1 <- subset(x = lib1,
                        subset = percent.mt < 5.00 & 
                        nCount_RNA > 500 & nFeature_RNA > 500 &
                        nCount_RNA < quantile(nCount_RNA, prob = 1 - 5/100) & 
                        nFeature_RNA < quantile(nFeature_RNA, prob = 1 - 2/100)
                        )
lib2 <- subset(x = lib2,
                        subset = percent.mt < 5.00 & 
                        nCount_RNA > 500 & nFeature_RNA > 500 &
                        nCount_RNA < quantile(nCount_RNA, prob = 1 - 5/100) & 
                        nFeature_RNA < quantile(nFeature_RNA, prob = 1 - 2/100)
                        )
lib3 <- subset(x = lib3,
                        subset = percent.mt < 5.00 & 
                        nCount_RNA > 500 & nFeature_RNA > 500 &
                        nCount_RNA < quantile(nCount_RNA, prob = 1 - 5/100) & 
                        nFeature_RNA < quantile(nFeature_RNA, prob = 1 - 2/100)
                        )
#lib4 <- subset(x = lib4,
#                        subset = percent.mt < 5.00 & 
#                        nCount_RNA > 500 & nFeature_RNA > 500 &
#                        nCount_RNA < quantile(nCount_RNA, prob = 1 - 5/100) & 
#                        nFeature_RNA < quantile(nFeature_RNA, prob = 1 - 2/100)
#                        )

lib5 <- subset(x = lib5,
                        subset = percent.mt < 5.00 & 
                        nCount_RNA > 500 & nFeature_RNA > 500 &
                        nCount_RNA < quantile(nCount_RNA, prob = 1 - 5/100) & 
                        nFeature_RNA < quantile(nFeature_RNA, prob = 1 - 2/100)
                        )

lib6 <- subset(x = lib6,
                        subset = percent.mt < 5.00 & 
                        nCount_RNA > 500 & nFeature_RNA > 500 &
                        nCount_RNA < quantile(nCount_RNA, prob = 1 - 5/100) & 
                        nFeature_RNA < quantile(nFeature_RNA, prob = 1 - 2/100)
                        )

postfilter_total <- sum(length(colnames(lib1)), length(colnames(lib2)), length(colnames(lib3)), length(colnames(lib5)), length(colnames(lib6)))
print(paste0("pre-filtering cell counts = ", prefilter_total))
print(paste0("post-filtering cell counts = ", postfilter_total))
print(paste0("fraction lost = ", 1 - postfilter_total/prefilter_total))
```
# Setting up libraries for integration
```{r}
lib.list <- c(lib1, lib2, lib3, lib5, lib6)

# Normalize and find variable features for each lib
lib.list <- lapply(X = lib.list, FUN = function(x) {
    x <- NormalizeData(x) 
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})

# select features that are repeatedly variable across datasets for integration run PCA on each
# dataset using these features
features <- SelectIntegrationFeatures(object.list = lib.list)
lib.list <- lapply(X = lib.list, FUN = function(x) {
    x <- ScaleData(x, features = features, verbose = FALSE)
    x <- RunPCA(x, features = features, verbose = FALSE)
})
```
# Find anchors
```{r}
lib.anchors <- FindIntegrationAnchors(object.list = lib.list, anchor.features = features, reduction = "cca")
```
# Integrate data
```{r}
# this command creates an 'integrated' data assay
lib.combined <- IntegrateData(anchorset = lib.anchors)
```
# Create new variable and dump old
```{r}
mouseBM <- lib.combined

rm(lib.combined, lib1, lib2, lib3, lib4, lib5, lib6)
```
# Standard workflow time
```{r}
DefaultAssay(mouseBM) <- "integrated"

mouseBM <- ScaleData(mouseBM)
mouseBM <- RunPCA(mouseBM, npcs = 30)
ElbowPlot(
  object = mouseBM, 
  ndims = 30
) +
  geom_abline(
    aes(intercept = 1.75, slope = 0, color = "red"),
    show.legend = FALSE
  )
mouseBM <- FindNeighbors(mouseBM, dims = 1:30)
mouseBM <- FindClusters(mouseBM, resolution = 0.2)
mouseBM <- RunUMAP(mouseBM, dims = 1:30)
DimPlot(mouseBM, label = T, group.by = "integrated_snn_res.0.2")
```
# Save out object
```{r}
saveRDS(lib.combined, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/5TGM1_IntegratedObject_StandardWorkflow.rds")
```
