---
title: "5TGM1_AllCells_ManualAnnotation"
output: html_document
date: "2023-06-01"
---
################################################################################

Created 6/1/2023 by Dylan Gagler, MSc - dylan.gagler@nyulangone.org

Code used to manually annotate both the stromal and non-stromal compartment of the mouse BM. Accepts QC filtered and integrated data. After annotation, this excludes non-stromal cells and saves out an rds object with stromals only.

Stromal markers derived from Baryawno et al., 2019 - https://www.cell.com/cell/pdf/S0092-8674(19)30459-3.pdf

Immune and erythroid markers derived from general apriori knowledge.

################################################################################

Load libraries
```{r}
library(Seurat)
library(ggplot2)
```
Read in data
```{r}
mouseBM <- readRDS("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/objects/Mouse_5TGM1_QC_IntegratedObject_Final.rds")
DefaultAssay(mouseBM) <- "RNA"
```
########################################
### Stromal Compartment Marker Genes ###
########################################
```{r}
# Mesenchymal stem cells (MSCs)
FeaturePlot(mouseBM, features = "Lepr")
FeaturePlot(mouseBM, features = "Adipoq")
FeaturePlot(mouseBM, features = "Cxcl12")

# Osteo-lineage cells (OLCs)
FeaturePlot(mouseBM, features = "Bglap")
FeaturePlot(mouseBM, features = "Spp1")
FeaturePlot(mouseBM, features = "Sp7")

# Chondrocytes 
FeaturePlot(mouseBM, features = "Col2a1")
FeaturePlot(mouseBM, features = "Sox9")
FeaturePlot(mouseBM, features = "Acan")

# Pericytes
FeaturePlot(mouseBM, features = "Acta2")
FeaturePlot(mouseBM, features = "Myh11")
FeaturePlot(mouseBM, features = "Mcam")

# Endothelial Cells (BMECs)
FeaturePlot(mouseBM, features = "Cdh5")
FeaturePlot(mouseBM, features = "Kdr")
FeaturePlot(mouseBM, features = "Emcn")

# Fibroblasts
FeaturePlot(mouseBM, features = "S100a4")
FeaturePlot(mouseBM, features = "Fn1")
FeaturePlot(mouseBM, features = "Dcn")
```
############################################
### Non-Stromal Compartment Marker Genes ###
############################################
```{r}
# B cells
FeaturePlot(mouseBM, features = "Cd79a")

# Hematopoietic Stem Cells
FeaturePlot(mouseBM, features = "Elane")
FeaturePlot(mouseBM, features = "Mpo")

# Monocyte/macrophages
FeaturePlot(mouseBM, features = "Ccl3")
FeaturePlot(mouseBM, features = "Tyrobp")

# Platelets
FeaturePlot(mouseBM, features = "Ppbp")

# RBCs
FeaturePlot(mouseBM, features = "Hba-a1")

# Neutrophil/GMPs
FeaturePlot(mouseBM, features = "S100a9")
FeaturePlot(mouseBM, features = "Ngp")

# T cells (none found)
FeaturePlot(mouseBM, features = "Cd4")
FeaturePlot(mouseBM, features = "Nkg7")

# Cycling cells
FeaturePlot(mouseBM, features = "Car1")
FeaturePlot(mouseBM, features = "Cdk6")
FeaturePlot(mouseBM, features = "Slc25a21")

# Plasma cells (none found)
FeaturePlot(mouseBM, features = "Ly6k")
FeaturePlot(mouseBM, features = "Slamf7")

# Oligodendrocytes
FeaturePlot(mouseBM, features = "Cdh19")
FeaturePlot(mouseBM, features = "Plp1")
```
Check out clusters on UMAP
```{r}
DimPlot(mouseBM, group.by = "seurat_clusters", label = T)
```
Changing idents
```{r}
Idents(mouseBM) <- mouseBM@meta.data$seurat_clusters

mouseBM <- RenameIdents(mouseBM, "0" = "B cells")
mouseBM <- RenameIdents(mouseBM, "1" = "MoMacs")
mouseBM <- RenameIdents(mouseBM, "2" = "Cycling")
mouseBM <- RenameIdents(mouseBM, "3" = "B cells")
mouseBM <- RenameIdents(mouseBM, "4" = "AEC")
mouseBM <- RenameIdents(mouseBM, "5" = "Fibroblasts")
mouseBM <- RenameIdents(mouseBM, "6" = "MSC")
mouseBM <- RenameIdents(mouseBM, "7" = "Platelets")
mouseBM <- RenameIdents(mouseBM, "8" = "SEC")
mouseBM <- RenameIdents(mouseBM, "9" = "B cells")
mouseBM <- RenameIdents(mouseBM, "10" = "Pericytes")
mouseBM <- RenameIdents(mouseBM, "11" = "Cycling")
mouseBM <- RenameIdents(mouseBM, "12" = "OLC")
mouseBM <- RenameIdents(mouseBM, "13" = "MoMacs")
mouseBM <- RenameIdents(mouseBM, "14" = "B cells")
mouseBM <- RenameIdents(mouseBM, "15" = "B cells")
mouseBM <- RenameIdents(mouseBM, "16" = "RBC")
mouseBM <- RenameIdents(mouseBM, "17" = "Platelets")
mouseBM <- RenameIdents(mouseBM, "18" = "Chondrocytes")
mouseBM <- RenameIdents(mouseBM, "19" = "Oligodendrocytes")

mouseBM@meta.data$celltypes <- Idents(mouseBM)
DimPlot(mouseBM, group.by = "celltypes", label = T)
```
Subsetting out stromals
```{r}
stromals <- subset(mouseBM, subset = celltypes %in% c("Chondrocytes", "OLC", "MSC", "AEC", "SEC", "Fibroblasts", "Pericytes"))
stromals@meta.data$celltypes <- droplevels(stromals@meta.data$celltypes)
```
Reclustering stromals
```{r}
stromals <- ScaleData(stromals)
stromals <- FindVariableFeatures(stromals, nfeatures = 2000)
stromals <- RunPCA(stromals, npcs = 25)
ElbowPlot(
  object = stromals, 
  ndims = 25
) +
  geom_abline(
    aes(intercept = 1.75, slope = 0, color = "red"),
    show.legend = FALSE
  )

# Determine percent of variation associated with each PC
pct <- stromals[["pca"]]@stdev / sum(stromals[["pca"]]@stdev) * 100
cumu <- cumsum(pct) # Calculate cumulative percents for each PC

# Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5
co1 <- which(cumu > 90 & pct < 5)[1]
# Determine the difference between variation of PC and subsequent PC
co2 <- sort(which((pct[1:length(pct) - 1] - pct[2:length(pct)]) > 0.1), decreasing = T)[1] + 1

if (co1 < co2) {
  n_pc <- co1
  print(paste0("the elbow PC is ", co1))
} else {
  n_pc <- 2
  print(paste0("the elbow point is ", co2))
}

stromals <- FindNeighbors(stromals, dims = 1:21)
stromals <- FindClusters(stromals, resolution = 0.12)
stromals <- RunUMAP(stromals, dims = 1:21)

DimPlot(stromals, label = T)
DimPlot(stromals, group.by = "celltypes", label = T)
```
Identifying B cell contam
```{r}
FeaturePlot(stromals, features = c("Cd79a"))
```
Subsetting B cell contam and running final workflow
```{r}
stromals.subbed <- subset(stromals, subset = seurat_clusters %in% c("0", "1", "2", "3", "4", "5", "6", "9"))

stromals.subbed <- ScaleData(stromals.subbed)
stromals.subbed <- FindVariableFeatures(stromals.subbed, nfeatures = 2000)
stromals.subbed <- RunPCA(stromals.subbed, npcs = 21)
stromals.subbed <- FindNeighbors(stromals.subbed, dims = 1:21)
stromals.subbed <- FindClusters(stromals.subbed, resolution = 0.12)
stromals.subbed <- RunUMAP(stromals.subbed, dims = 1:21)

DimPlot(stromals.subbed, label = T)
DimPlot(stromals.subbed, group.by = "celltypes", label = T)
table(stromals.subbed@meta.data$celltypes)
```
Save out
```{r}
saveRDS(stromals.subbed, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/objects/5TGM1_StromalsOnly_Annotated_Object.rds")
```

