---
title: "ComplexHeatmap_Prep"
output: html_document
date: "2023-02-14"
---

# Libraries
```{r}
library(Seurat)
library(ggplot2)
library(circlize)
library(ComplexHeatmap)
library(ggpubr)
library(grid)
library(gridExtra)
library(tidyverse)
library(scales)
show_col(hue_pal()(6))

colors <- c("#F8766D", "#B79F00", "#00BA38", "#00BFC4", "#619CFF", "#F564E3")
```
# Load in stromal data
```{r}
stromals <- readRDS("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Rerun_Mouse5TG_StromalsOnly_AnnotatedObject6_21PCs_NoLib_FIXEDLIBRARIES_6.rds")
```
# Subset celltype
```{r}
subtype <- subset(stromals, subset = celltypes_ECsubbed %in% c("Sinusoidal ECs", "Arterial ECs"))
```
# Standard workflow
# Doing this because we want the most variable genes for the fibroblasts specifically, not for the entire dataset
```{r}
subtype <- ScaleData(subtype)
subtype <- FindVariableFeatures(subtype, nfeatures = 2000)
subtype <- RunPCA(subtype, npcs = 5)
subtype <- FindNeighbors(subtype, dims = 1:5)
subtype <- FindClusters(subtype, resolution = 0.12)
subtype <- RunUMAP(subtype, dims = 1:5)
DimPlot(subtype, split.by = "case.control")
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCOLCs_Subclustered_ConditionSplit_UMAP.png", height = 4, width = 8)
```

# Identifying and removing B cell contam
```{r}
FeaturePlot(subtype, features = c("Cd79a"))
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCOLCs_Cd79aContam_UMAP.png", height = 4, width = 8)

FeaturePlot(subtype, features = c("Cdh5"))
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCOLCs_Cdh5Contam_UMAP.png", height = 4, width = 8)

FeaturePlot(subtype, features = c("Epcam"))
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCOLCs_EpcamContam_UMAP.png", height = 4, width = 8)

FeaturePlot(subtype, features = c("Lilr4b"))
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCOLCs_Lilr4bContam_UMAP.png", height = 4, width = 8)

subtype <- subset(subtype, subset = seurat_clusters %in% c("0", "1", "2", "3", "4", "5"))
Idents(subtype) <- "seurat_clusters"
DimPlot(subtype, cols = colors, 
        split.by = "case.control")
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCOLCs_Subclustered_ConditionSplit_UMAP_ContamRemoved.png", height = 5, width = 8)
```
# Get differentially expressed genes
```{r}
Idents(subtype) <- "case.control" 
all.markers <- FindAllMarkers(subtype, assay = "RNA")
all.markers <- all.markers[order(-all.markers$avg_log2FC),]
```
# Get DA
```{r}
healthy.sub <- subset(subtype, subset = case.control == "Healthy")
myeloma.sub <- subset(subtype, subset = case.control == "Myeloma")

table(healthy.sub@meta.data$seurat_clusters)/sum(table(healthy.sub@meta.data$seurat_clusters))
table(myeloma.sub@meta.data$seurat_clusters)/sum(table(myeloma.sub@meta.data$seurat_clusters))
table(subtype@meta.data$seurat_clusters)/sum(table(subtype@meta.data$seurat_clusters))

```

# Top N logFC genes
```{r}
# top100_pos_lfc_genes <- all.markers %>%
#                           dplyr::arrange(p_val_adj) %>%
#                           dplyr::filter(avg_log2FC > 0) %>%
#                           dplyr::select(gene) %>%
#                           head(n = 300) %>%
#                           purrr::as_vector()
# 
# top100_neg_lfc_genes <- all.markers %>%
#                           dplyr::arrange(p_val_adj) %>%
#                           dplyr::filter(avg_log2FC < 0) %>%
#                           dplyr::select(gene) %>%
#                           head(n = 300) %>%
#                           purrr::as_vector()
```
### For use when getting gene order for different clusters
```{r}
output_gene_order <- c()

for(i in c("0", "1", "2", "3", "4", "5")) {
  print(i)

  cluster_genes <- all.markers %>% filter(cluster == i) %>% dplyr::arrange(p_val_adj) %>%
                      dplyr::filter(avg_log2FC > 0) %>%
                      dplyr::select(gene) %>%
                      head(n = 15) %>%
                      purrr::as_vector()
  print(cluster_genes)

  output_gene_order <- c(output_gene_order, cluster_genes)
}

output_gene_order <- unique(output_gene_order)
```
### For use when using case/control markers and gene orders
```{r}
output_gene_order <- c()

for(i in c("Healthy", "Myeloma")) {
  print(i)

  cluster_genes <- all.markers %>% dplyr::filter(cluster == i) %>% dplyr::arrange(p_val_adj) %>%
                      dplyr::filter(avg_log2FC > 0) %>%
                      dplyr::select(gene) %>%
                      head(n = 80) %>%
                      purrr::as_vector()
  print(cluster_genes)

  output_gene_order <- c(output_gene_order, cluster_genes)
}

output_gene_order <- unique(output_gene_order)
```
# Subset matrix by top genes
```{r}
matrix <- subtype@assays$RNA@scale.data
#matrix.subset <- matrix[rownames(matrix) %in% c(top100_pos_lfc_genes, top100_neg_lfc_genes),]
matrix.roworder.subset <- matrix[rownames(matrix) %in% output_gene_order,]

output_gene_order_sub <- output_gene_order[output_gene_order %in% rownames(matrix.roworder.subset)]
```
# Get annotation metadata
```{r}
annotation <- HeatmapAnnotation("Disease Status" = subtype@meta.data$case.control,
                                "Seurat Cluster" = subtype@meta.data$seurat_clusters,
                                simple_anno_size = unit(3, "mm"),
                                col = list("Disease Status" = c("Healthy" = "green", 
                                                                "Myeloma" = "purple"),
                                           "Seurat Cluster" = c("0" = "#F8766D",
                                                         "1" = "#B79F00",
                                                         "2" = "#00BA38",
                                                         "3" = "#00BFC4",
                                                         "4" = "#619CFF",
                                                         "5" = "#F564E3"
                                                         )))

col_fun = colorRamp2(c(-2, 0, 2), c("blue", "white", "red"))
```
# Separate legend to include?
```{r}
lgd1 = Legend(col_fun = col_fun, title = "Scaled Expression")
lgd2 = Legend(title = "Disease Status", labels = c("Healthy", "Myeloma"), legend_gp = gpar(fill = c("green", "purple")))
lgd3 = Legend(title = "Seurat Cluster", labels = c("0", "1", "2", "3", "4", "5"),
              legend_gp = gpar(fill = colors))
pd = packLegend(lgd1, lgd2, lgd3, direction = "vertical")
draw(pd, row_title = "bals")
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCLineage_VerticalLegend_ComplexHeatmap.png", width = 1, height = 5)

```

# Complex heatmap
```{r}
Heatmap(matrix.roworder.subset, cluster_rows = F, cluster_columns = TRUE,
        row_order = output_gene_order_sub,
        show_column_dend = F,
        col = col_fun,
        column_names_gp = gpar(fontsize = 0), row_names_gp = gpar(fontsize = 4.5),
        clustering_distance_columns = "euclidean", clustering_method_columns = "average",
        column_dend_height = unit(40,"mm"), row_dend_width = unit(40, "mm"),
        top_annotation = annotation,
        heatmap_legend_param = list(title = "Scaled Expression")
        )
#ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCLineage_Monocle_SeuratClusters_ComplexHeatmap.png", width = 10, height = 20)

#DimPlot(subtype, cols = c("#F8766D", "#B79F00", "#00BA38", "#00BFC4", "#619CFF", "#F564E3", "#C77CFF"))
#ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCLineage_SeuratClusters_UMAP.png", width = 10, height = 20)
```

# Differential abundance
```{r}
table(subtype@meta.data$seurat_clusters)/sum(table(subtype@meta.data$seurat_clusters))
```


# Testoo
```{r}
# Adhesions
VlnPlot(subtype, features = c("Jag1", "Vcam1", "Icam1"), split.by = "case.control", group.by = "seurat_clusters")
FeaturePlot(subtype, features = c("Jag1", "Icam1", "Vcam1"), split.by = "case.control")

# Next step
VlnPlot(subtype, features = c("Flt4", "Cdh5", "Pecam1", "Ly6a", "Emcn"), split.by = "case.control", group.by = "seurat_clusters")
FeaturePlot(subtype, features = c("Flt1", "Kdr", "Flt4"), split.by = "case.control")

VlnPlot(subtype, features = c("Il6"), split.by = "case.control", group.by = "seurat_clusters")

#
VlnPlot(subtype, features = c("Fgf23", "Dkk2", "Stab2", "Rgs4", 
                              "Axin2", "Apcdd1", "Lef1", "Tcf7",
                              "Rgs4", "Angptl4", "Gpr182", "Flt4",
                              "Cd34", "Sparcl1"), split.by = "case.control", group.by = "seurat_clusters")
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/BothEC_HealthyFunction_CaseControl_ViolinPlots.png", width = 18, height = 10.5)

VlnPlot(subtype, features = c("Nrp1", "Plxna1", "Flt4"), split.by = "case.control")

# "Healthy" EC genes
VlnPlot(subtype, features = c("Ang2", "Il3ra", "Il6", "Kitl", 
                              "Jag1", "Cxcl12", "Il1b", 
                              "Vegfa", "Vegfb",
                              "Tgfb1", "Tgfb2", "Tgfb3"), split.by = "case.control", group.by = "seurat_clusters")
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/BothEC_HealthyFunction_CaseControl_ViolinPlots.png", width = 18, height = 10.5)

# Antigen presenting/immunomodulating
VlnPlot(subtype, features = c("Cd40", "Cd80", "Cd86", "Icosl", 
                              "Mb1", "Psmb9", "Psmb8", "Psmb10", 
                              "Bat1b", "Rfx1", "Cd74", "H2-Aa"), ncol = 5, split.by = "case.control", group.by = "seurat_clusters")
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/SEC_ImmunomodulatingGenes_CaseControl_ViolinPlots.png", width = 21, height = 7)

VlnPlot(subtype, features = c("Mgp", "Lpl", "Wif1", "Spp1", "Ibsp",
                                  "Col16a1", "Tnn",
                                  "Fbn1", "Igf1",
                                  "Bglap", "Car3"), ncol = 4, split.by = "case.control")
```

```{r}
fibrochondro <- subset(stromals, subset = celltypes_ECsubbed %in% c("Fibroblasts", "Chondrocytes"))
saveRDS(fibrochondro, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/5TG_Mouse_FibroblastsChondrocyte_Subset.rds")
```
# Tikhonova and Dolgalev based MSC differentiation exploration
```{r}
FeaturePlot(subtype, features = "Lepr")
FeaturePlot(stromals, features = "Lepr")
FeaturePlot(stromals, features = "Col1a1")
FeaturePlot(stromals, features = "Cdh5")

DimPlot(stromals, group.by = "celltypes_ECsubbed")

VlnPlot(subtype, features = c("Mgp", "Lpl", "Wif1", "Spp1", "Ibsp"), ncol = 5)
VlnPlot(subtype, features = c("Mgp", "Lpl", "Wif1", "Spp1", "Ibsp"), ncol = 5, split.by = "case.control")

# P - Cluster 1
VlnPlot(subtype, features = c("Mgp", "Gpx3", "Serping1", "Lepr", "Tmem176b", 
                              "Igfbp5", "Malat1", "C1ra", "C4b", "Epas1"), ncol = 5)
# P - Cluster 2
VlnPlot(subtype, features = c("Lpl", "Scp2", "Fstl1", "Rgcc", "Mrps6", 
                              "Pdzrn4", "Mmd", "Np2", "Slc5a3", "Angpt1"), ncol = 5)
# P - Cluster 3
VlnPlot(subtype, features = c("Col8a1", "Kcnk2", "Ndnf", "Palld", "Tnfrsf19", 
                              "Cldn10", "Slc20a2", "Limch1", "Fhl1"), ncol = 5)
# P - Cluster 4
VlnPlot(subtype, features = c("Spp1", "Col1a1", "Col13a1", "Mmp13", "Ifitm5", 
                              "Serpine2", "Mef2c", "Ibsp", "Itgb5", "Aqp1"), ncol = 5)
# Osteo - Cluster 1
VlnPlot(subtype, features = c("Tnn", "Postn", "Col16a1", "Lrrc15", "Angptl2", 
                              "Mmp14", "Alpl", "Ncam1", "Mmp13", "Col5a1"), ncol = 5)
# Osteo - Cluster 2
VlnPlot(subtype, features = c("Igfbp7", "Igfbp6", "Mgp", "Fbn1", "Thbs4", 
                              "Prelp", "Igf1", "Lbp", "Meg3", "Fbln2"), ncol = 5)
# Osteo - Cluster 3
VlnPlot(subtype, features = c("Ifitm", "Col11a2", "Car3", "Cgref1", "Bglap", 
                              "Slc36a2", "Cpz", "Sgm2", "Smpd3", "Mef2c"), ncol = 5)

DimPlot(subtype)

# Full set of bone cells
bone.subset <- subset(stromals, subset = celltypes_ECsubbed %in% c("MSCs", "OLCs"))
bone.subset <- ScaleData(bone.subset)
bone.subset <- FindVariableFeatures(bone.subset, nfeatures = 2000)
bone.subset <- RunPCA(bone.subset, npcs = 5)

bone.subset <- FindNeighbors(bone.subset, dims = 1:5)
bone.subset <- FindClusters(bone.subset, resolution = 0.12)
bone.subset <- RunUMAP(bone.subset, dims = 1:5)

DimPlot(bone.subset, cols = c("#F8766D", "#B79F00", "#00BA38", "#00BFC4", "#619CFF", "#F564E3", "#C77CFF"), split.by = "case.control")

# Pseudotime directions - row 1 = bone, row 2 = adipo
VlnPlot(subtype, features = c("Bglap", "Runx2", "Col1a1", "Sparc",
                                  "Lepr", "Adipoq", "Lpl", "Cebpa"), ncol = 4)
# Igor cluster vs our cluster parallel
VlnPlot(subtype, features = c("Mgp", "Lpl", "Wif1", "Spp1", "Ibsp",
                                  "Col16a1", "Tnn", "Fbn1", "Igf1",
                                  "Bglap", "Car3"), ncol = 4)
DimPlot(bone.subset, cols = c("#F8766D", "#B79F00", "#00BA38", "#00BFC4", "#619CFF", "#F564E3", "#C77CFF"), split.by = "case.control")
DimPlot(bone.subset, group.by = "celltypes_ECsubbed", split.by = "case.control")


VlnPlot(bone.subset, features = c("Mgp", "Lpl", "Wif1", "Spp1", "Ibsp"), ncol = 5, split.by = "case.control")

# Osteo - Cluster 1
VlnPlot(subtype, features = c("Tnn", "Postn", "Col16a1", "Lrrc15", "Angptl2", 
                              "Mmp14", "Alpl", "Ncam1", "Mmp13", "Col5a1"), ncol = 5)
# Osteo - Cluster 2
VlnPlot(subtype, features = c("Igfbp7", "Igfbp6", "Mgp", "Fbn1", "Thbs4", 
                              "Prelp", "Igf1", "Lbp", "Meg3", "Fbln2"), ncol = 5)
# Osteo - Cluster 3
VlnPlot(subtype, features = c("Sparc", "Col11a2", "Car3", "Cgref1", "Bglap", 
                              "Runx2", "Bglap2", "Col1a2", "Smpd3", "Mef2c"), ncol = 5)

VlnPlot(bone.subset, features = c("Tnn", "Postn", "Col16a1", "Lrrc15", "Angptl2", 
                              "Mmp14", "Alpl", "Ncam1", "Mmp13", "Col5a1"), ncol = 5, split.by = "case.control")
# Osteo - Cluster 2
VlnPlot(bone.subset, features = c("Igfbp7", "Igfbp6", "Mgp", "Fbn1", "Thbs4", 
                              "Prelp", "Igf1", "Lbp", "Meg3", "Fbln2"), ncol = 5, split.by = "case.control")
# Osteo - Cluster 3
VlnPlot(bone.subset, features = c("Ifitm", "Col11a2", "Car3", "Cgref1", "Bglap", 
                              "Slc36a2", "Cpz", "Sgm2", "Smpd3", "Mef2c"), ncol = 5, split.by = "case.control")

# Aggregated good bone markers
# O1
VlnPlot(subtype, features = c("Bglap", "Bglap2", "Sparc", "Col1a2", "Mef2c"), split.by = "case.control")
# O2
VlnPlot(subtype, features = c("Mmp14", "Alpl", "Ncam1", "Mmp13"), split.by = "case.control")
# O3
VlnPlot(subtype, features = c("Igfbp7", "Igf1", "Fbn1", "Mgp", "Prelp", "Lbp"))
VlnPlot(subtype, features = c("Igfbp7", "Igf1", "Fbn1", "Mgp", "Prelp", "Lbp"), split.by = "case.control")

# Get cluster subset %s
table(healthy.mscolcs@meta.data$seurat_clusters)/sum(table(healthy.mscolcs@meta.data$seurat_clusters))
table(myeloma.mscolcs@meta.data$seurat_clusters)/sum(table(myeloma.mscolcs@meta.data$seurat_clusters))

```

