---
title: "EndoMT_Exploration"
output: html_document
date: "2023-06-12"
---

# Libraries
```{r}
library(Seurat)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(patchwork)
library(monocle)
library(monocle3)

#colors <- c("#F8766D", "#B79F00", "#00BA38", "#00BFC4", "#619CFF", "#F564E3")
```
# Load in stromal data
```{r}
stromals <- readRDS("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/objects/Rerun_Mouse5TG_StromalsOnly_AnnotatedObject6_21PCs_NoLib_FIXEDLIBRARIES_6.rds")
```
# Subset celltype
```{r}
subtype <- subset(stromals, subset = celltypes_ECsubbed %in% c("Sinusoidal ECs", "Arterial ECs", "MSCs", "OLCs"))
```
# Standard workflow
# Doing this because we want the most variable genes for the fibroblasts specifically, not for the entire dataset
```{r}
subtype <- ScaleData(subtype)
subtype <- FindVariableFeatures(subtype, nfeatures = 2000)
subtype <- RunPCA(subtype, npcs = 10)
subtype <- FindNeighbors(subtype, dims = 1:10)
subtype <- FindClusters(subtype, resolution = 0.2)
subtype <- RunUMAP(subtype, dims = 1:10)
DimPlot(subtype)

#subtype <- subset(subtype, subset = seurat_clusters %in% c("0", "1", "2", "3", "4", "5"))
#DimPlot(subtype, split.by = "case.control")
#ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/BMECs_Subclustered_NotSplit_UMAP.png", height = 5, width = 8)
```
# Explore genes
```{r}
# Subsetting out cells with Cdh5 expression (this is what they did in the paper)
cdh5 <- subset(stromals, subset = Cdh5 > 0)

cdh5 <- ScaleData(cdh5)
cdh5 <- FindVariableFeatures(cdh5, nfeatures = 2000)
cdh5 <- RunPCA(cdh5, npcs = 10)
cdh5 <- FindNeighbors(cdh5, dims = 1:10)
cdh5 <- FindClusters(cdh5, resolution = 0.2)
cdh5 <- RunUMAP(cdh5, dims = 1:10)
DimPlot(cdh5)

VlnPlot(cdh5, features = c("Cdh5", "Pecam1", "Eng", "Emcn", "Prrx1", "Col1a1", "Snai2", "Twist1", "Twist2", "Zeb2", "Yap1", "Cxcl12"), ncol = 4)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Cdh5_Pos_Cells_EndoMT_Genes_ViolinPlots.png", height = 12, width = 18)
DimPlot(cdh5)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Cdh5_Pos_Cells_UMAP.png", height = 4, width = 5)
DimPlot(cdh5, group.by = "celltypes_ECsubbed")
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Cdh5_Pos_Cells__CellTypes_UMAP.png", height = 4, width = 5)
DimPlot(cdh5, split.by = "case.control")
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Cdh5_Pos_Cells_ConditionSplit_UMAP.png", height = 4, width = 8)

```

# Selecting endoMT cells
```{r}
# Based on violin plots above
endoMTcells <- subset(cdh5, subset = seurat_clusters %in% c("4", "6"))
endoMTcellnames <- colnames(endoMTcells)

# Select cells with true endoMT expression
plot <- DimPlot(endoMTcells)
select.cells <- CellSelector(plot = plot)

# Assign to objects
Idents(cdh5) <- "seurat_clusters"
Idents(cdh5, cells = select.cells) <- "EndoMT"
cdh5@meta.data$endoMT <- Idents(cdh5)
DimPlot(cdh5)

endoMT.object <- subset(cdh5, subset = endoMT == "EndoMT")

DimPlot(object = cdh5, cells.highlight = select.cells, cols.highlight = "red", cols = "gray", order = TRUE, label = T) + NoLegend()
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Cdh5Pos_EndoMT_Highlighted_UMAP.png", height = 4, width = 5)

saveRDS(cdh5, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/objects/5TGM1_Cdh5Pos_EndoMT_Annotated_FromOG_Object.rds")

# Backtracking Endo-MT cells to original object
Idents(subtype) <- "seurat_clusters"
Idents(subtype, cells = select.cells) <- "Endo-MT"
subtype@meta.data$endoMT <- Idents(subtype)
DimPlot(subtype)
saveRDS(subtype, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/objects/5TGM1_BMEC_MSCOLC_EndoMT_Annotated_FromOG_Object.rds")
```
#############################################################
################# Generating "Final Figure" #################
#############################################################

# Emulating format from original plot -

AAABBBBB where A = UMAP of BMEC + MSCOLC, B = feature plots of top marker genes for those pops, C = Cdh5 UMAP, D = highlighting EndoMT cells UMAP, E = GSEA, and F = EndoMT violin plots 
AAABBBBB
CCCDDDEE
CCCDDDEE
FFFFFFFF
FFFFFFFF
# Figure A
```{r}
figA <- DimPlot(subtype, group.by = "seurat_clusters") + ggtitle("BMECs and MSC/OLCs")
```
# Figure B
```{r}
p1.ft <- FeaturePlot(subtype, features = c("Cdh5")) + theme(axis.title.x = element_blank()) + NoLegend()
p2.ft <- FeaturePlot(subtype, features = c("Pecam1")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p3.ft <- FeaturePlot(subtype, features = c("Eng")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p4.ft <- FeaturePlot(subtype, features = c("Col1a1"))  + NoLegend()
p5.ft <- FeaturePlot(subtype, features = c("Prrx1")) + theme(
                                                             axis.title.y = element_blank()) + NoLegend()
p6.ft <- FeaturePlot(subtype, features = c("Ngfr")) + theme(
                                                             axis.title.y = element_blank()) + NoLegend()

figB <- (p1.ft | p2.ft | p3.ft) / (p4.ft | p5.ft | p6.ft)
```
# Figure C
```{r}
figC <- DimPlot(cdh5, group.by = "seurat_clusters") + ggtitle("Cdh5 expressing cells")
```
# Figure D
```{r}
figD <- DimPlot(object = cdh5, cells.highlight = select.cells, cols.highlight = "red", cols = "gray", order = TRUE, label = T, label.size = 5, repel = T) + NoLegend()
```
# Figure D
```{r}
#VlnPlot(cdh5, features = c("Cdh5", "Pecam1", "Eng", "Emcn", "Prrx1", "Col1a1", "Snai2", "Twist1", "Twist2", "Zeb2", "Yap1", "Cxcl12"), ncol = 4)
Idents(cdh5) <- "seurat_clusters"
p1.vln <- VlnPlot(cdh5, features = c("Cdh5")) + theme(axis.title.x = element_blank(),
                                                      axis.ticks.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
p2.vln <- VlnPlot(cdh5, features = c("Pecam1")) + theme(axis.title.x = element_blank(),
                                                        axis.ticks.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
p3.vln <- VlnPlot(cdh5, features = c("Emcn")) + theme(axis.title.x = element_blank(),
                                                      axis.ticks.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
p4.vln <- VlnPlot(cdh5, features = c("Prrx1")) + theme(axis.title.x = element_blank(),
                                                       axis.ticks.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
p5.vln <- VlnPlot(cdh5, features = c("Col1a1")) + theme(axis.title.x = element_blank(),
                                                        axis.ticks.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
p6.vln <- VlnPlot(cdh5, features = c("Snai2")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
p7.vln <- VlnPlot(cdh5, features = c("Twist1")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
p8.vln <- VlnPlot(cdh5, features = c("Zeb2")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
p9.vln <- VlnPlot(cdh5, features = c("Yap1")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
p10.vln <- VlnPlot(cdh5, features = c("Cxcl12")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()

figE <- (p1.vln | p2.vln | p3.vln | p4.vln | p5.vln ) / (p6.vln | p7.vln | p8.vln | p9.vln | p10.vln)
```
# Figure 6 Patchwork
```{r fig.height = 11, fig.width = 11}
layout <- "
AAABBBBBBB
AAABBBBBBB
CCCDDDD###
CCCDDDD###
EEEEEEEEEE
EEEEEEEEEE
"

panel <- figA + figB + figC + figD + figE +
  plot_layout(design = layout)
panel
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Figure6_EndoMT.png", height = 11, width = 11)
```
```{r}
VlnPlot(cdh5, features = c("Tgfbr1", "Tgfbr2", "Tgfbr3")) + NoLegend()
```


# Senscence
```{r}
stromals2 <- readRDS("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/objects/Rerun_Mouse5TG_StromalsOnly_AnnotatedObject6_21PCs_NoLib_FIXEDLIBRARIES_6.rds")
msc <- subset(stromals2, subset = celltypes_ECsubbed %in% c("MSCs", "OLCs"))

msc <- ScaleData(msc)
msc <- FindVariableFeatures(msc, nfeatures = 2000)
msc <- RunPCA(msc, npcs = 10)
msc <- FindNeighbors(msc, dims = 1:10)
msc <- FindClusters(msc, resolution = 0.2)
msc <- RunUMAP(msc, dims = 1:10)
msc <- subset(msc, subset = seurat_clusters %in% c("0", "1", "2", "3", "4", "5"))
DimPlot(msc)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/MSCOLC_UMAP.png", height = 4, width = 5)
```
# Checking senescence expression as per the table Faith sent
```{r}
# only included genes with substantial expression in at least 1 cluster
VlnPlot(msc, features = c("Il6st", "Il7","Il15", "Cxcl1", "Csf2", "Ifngr1", "Cxcl13", "Mif", "Fgf7", "Vegfa",  "Kitl", "Cxcl12", "Igfbp3",
                          "Igfbp4", "Igfbp7", "Icam1", "Icam3", "Tnfrsf1a", "Il6st", "Fgfr2", "Mmp13", "Mmp14", "Timp1", "Timp2", "Serpine1", "Serpine2", "Ctsb",
                          "Fn1", "Lama1", "Lamb1"), ncol=7)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/AllPos_SenescenceGenes_VlnPlots.png", height = 12, width = 21)

```
# Find markers for cluster 5
```{r}
aberrant.cluster <- FindAllMarkers(msc)
cluster5 <- aberrant.cluster[aberrant.cluster$cluster == "5",]
cluster5 <- cluster5[cluster5$p_val_adj < 0.05,]
cluster5 <- cluster5 %>% arrange((p_val_adj))
rownames(cluster5) <- cluster5$gene
write.csv(cluster5, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun//MMaberrantCluster_DEgenes.csv")
```

