---
title: "Generate_DESeq2_GSEA_Input"
output: html_document
date: "2023-01-31"
---

```{r}
library(Seurat)
library(Matrix)
library(tibble)
library(SummarizedExperiment)
library(DESeq2)
```
# Read in stromals
```{r}
stromals <- readRDS("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Rerun_Mouse5TG_StromalsOnly_AnnotatedObject6_21PCs_NoLib_FIXEDLIBRARIES_6.rds")
celltype <- "Fibroblasts"
healthy.celltype <- subset(stromals, subset = celltype_cond == paste0("Healthy ", celltype))
myeloma.celltype <- subset(stromals, subset = celltype_cond == paste0("MM ", celltype))
```
# Write matrices
```{r}
matrix <- as.data.frame(GetAssayData(object = healthy.celltype, slot = "counts"))
matrix <- matrix + 1
write.table(matrix, paste0('/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/healthy', celltype, '_counts.txt'), sep = "\t")

matrix <- as.data.frame(GetAssayData(object = myeloma.celltype, slot = "counts"))
matrix <- matrix + 1
write.table(matrix, paste0('/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/myeloma', celltype, '_counts.txt'), sep = "\t")
```
# Make other metadata
# Need a metadata file with 3 columns...1st is cell names, 2nd is the condition (so in this case, just "myeloma" or "healthy", and 3rd is "paired end")
```{r}
metadata <- as.data.frame(colnames(healthy.celltype))
colnames(metadata) <- "V1"
metadata$V2 <- "Healthy"
metadata$V3 <- "paired-end"
head(metadata)

write.table(metadata, paste0("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/healthy", celltype, "_conditions.txt"))

###

metadata <- as.data.frame(colnames(myeloma.celltype))
colnames(metadata) <- "cellnames"
metadata$condition <- "Myeloma"
metadata$sequencing <- "paired-end"
head(metadata)

write.table(metadata, paste0("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/myeloma", celltype, "_conditions.txt"))
```
####################################################################################################
################################# Randomized pseudobulking version #################################
# DESeq2 (which is employed for GSEA) requires pseudobulking (cells collated into samples) to properly do DE gene expression.
# Unfortunately, we dont have samples. So we're gonna try to fake it by randomly subsampling cells down into 3 (and 5) samples per condition.
# Thinking that the best way to do this is to apply the pseudobulk to the healthies and myelomas separately and then remerging the objects

# THIS ISNT IT. Take another look at example counts....need to generate matrix wherein each row is the pseudobulk. And then the condition file doesn't need to be fully populated, I think? Just 1 row for each subsample?

```{r}
# Case/control breakdown
stromals <- readRDS("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Rerun_Mouse5TG_StromalsOnly_AnnotatedObject6_21PCs_NoLib_FIXEDLIBRARIES_6.rds")
healthy <- subset(stromals, subset = case.control == "Healthy")
myeloma <- subset(stromals, subset = case.control == "Myeloma")

# Randomly assigning cells to N samples
n <- 3
healthy@meta.data$pseudobulk <- sample(factor(rep(1:n, length.out = nrow(healthy@meta.data)), labels = paste0("Healthy", 1:n)))
table(healthy@meta.data$pseudobulk)
myeloma@meta.data$pseudobulk <- sample(factor(rep(1:n, length.out = nrow(myeloma@meta.data)), labels = paste0("Myeloma", 1:n)))
table(myeloma@meta.data$pseudobulk)
```
# Get subsample
```{r}
celltype <- "Fibroblasts"
healthy.subtype <- subset(healthy, subset = celltypes_ECsubbed == celltype)
myeloma.subtype <- subset(myeloma, subset = celltypes_ECsubbed == celltype)
```
# Aggregating for pseudobulk
```{r}
sce <- as.SingleCellExperiment(healthy.subtype)
# Generate sample level metadata

# Named vector of sample names
sids <- purrr::set_names(levels(sce$pseudobulk))

## Determine the number of cells per sample
table(sce$pseudobulk)

## Turn named vector into a numeric vector of number of cells per sample
n_cells <- as.numeric(table(sce$pseudobulk))

## Determine how to reoder the samples (rows) of the metadata to match the order of sample names in sids vector
m <- match(sids, sce$pseudobulk)

## Create the sample level metadata by combining the reordered metadata with the number of cells corresponding to each sample.
ei <- data.frame(colData(sce)[m, ], 
                  n_cells, row.names = NULL) %>%
  select(c("case.control", "pseudobulk", "n_cells"))
ei

# Aggregate the counts per sample_id and cluster_id

# Subset metadata to only include the cluster and sample IDs to aggregate across
groups <- colData(sce)[, c("pseudobulk")]

# Aggregate across cluster-sample groups
pb <- aggregate.Matrix(t(counts(sce)), 
                       groupings = groups, fun = "sum") 

class(pb)

dim(pb)

pb[1:6, 1:6]
```

# Write matrices
```{r}
matrix <- as.data.frame(GetAssayData(object = healthy.subtype, slot = "counts"))
matrix <- matrix + 1
write.table(matrix, paste0('/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/healthy', celltype, '_counts.txt'), sep = "\t")

matrix <- as.data.frame(GetAssayData(object = myeloma.subtype, slot = "counts"))
matrix <- matrix + 1
write.table(matrix, paste0('/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/myeloma', celltype, '_counts.txt'), sep = "\t")
```
# Make other metadata
# Need a metadata file with 3 columns...1st is cell names, 2nd is the condition (so in this case, just "myeloma" or "healthy", and 3rd is "paired end")
```{r}
metadata <- as.data.frame(healthy.subtype@meta.data$pseudobulk)
colnames(metadata) <- "Samples"
metadata$V2 <- "Healthy"
metadata$V3 <- "paired-end"
head(metadata)

write.table(metadata, paste0("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/healthy", celltype, "_pseudobulk_Conditions.txt"))

###

metadata <- as.data.frame(myeloma.subtype@meta.data$pseudobulk)
colnames(metadata) <- "Samples"
metadata$condition <- "Myeloma"
metadata$sequencing <- "paired-end"
head(metadata)

write.table(metadata, paste0("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/myeloma", celltype, "_pseudobulk_Conditions.txt"))
```
# TESTTT
```{r}
Heatmap(healthy.subtype@assays$integrated@scale.data, cluster_rows = TRUE, cluster_columns = TRUE,
        column_names_gp = gpar(fontsize = 0), row_names_gp = gpar(fontsize = 0),
        clustering_distance_columns = "euclidean", clustering_method_columns = "average",
        
        )
```

