---
title: "GSEA_RankedList_Generation"
output: html_document
date: "2023-02-02"
---
################################################################################
Created by Dylan Gagler, MSc - dylan.gagler@nyulangone.org or dylangagler@gmail.com

Code used for generating ranked lists for GSEA preranked analysis.

Ranking is done by first performing differential gene expression analysis using wilcoxon rank-sum test via Seurat's FindMarkers() and then multiplying the average logFC by -log10 of the adjusted p values. 

Sometimes this results in infinitely small p-values, in which case no rank-values can be computed. As such, a quick fix can be implemented by manually enforcing ranking values in the order that the genes appear in the output df.

################################################################################

Load libraries
```{r}
library(Seurat)
library(dplyr)
```
Load data. This might be entire dataset or just a subset of cells
```{r}
subtype <- readRDS("/Users/gagled01/morganLab/single-cell/5TGM1_Stromal/objects/5TGM1_MSCOLC_Object.rds")
```
Case vs. Control differential gene expression
```{r}
# If ident.1 = disease and ident.2 = control, then POSITIVE logFC means that the gene is upregulated in disease relative to control and NEGATIVE means the gene is downregulated in disease relative to control
Idents(subtype) <- "case.control"
markers <- FindMarkers(subtype, ident.1 = "myeloma", ident.2 = "control", min.pct = 0.25, logfc.threshold = 0.25)
```
# Generate ranked list
```{r}
# Generate rank score by mulitplying avg log FC by negative log10 of p value.
rank.score <- sign(markers$avg_log2FC) * -log10(markers$p_val_adj)

# Generate gene list
genes <- rownames(markers)

# Put into df
rank.df <- data.frame(genes, rank.score)
rownames(rank.df) <- rank.df$genes

# Optionally fixing infinity values caused by infinitely small zero p values. Just assuming each gene has +1 rank score.
# for (i in 1:72) {
#   rank.df$rank.score[i] <- 373 - i
# }
```
Write out ranked list
```{r}
write.table(rank.df, "/Users/gagled01/morganLab/single-cell/5TGM1_Stromal/GSEA/rankedLists/MSCOLC_CaseControl_RankList_forGSEA.rnk",
            sep = "\t",
            quote = F,
            row.names = F,
            col.names = F)
```

