---
title: "Integration_StandardWorkflow"
output: html_document
date: "2023-07-31"
---

This code takes the QC filtered cells and integrates them together using Seurat. Methodology based off of this vignette: https://satijalab.org/seurat/articles/integration_introduction.html

In short, this involves normalizing and finding variable genes for each library, selecting integration features that are highly variable across libraries, scaling and running a PCA on each library using these shared variable features, finding integration anchors within the shared features, and finally integrating the data together using the integration anchors.

Afterwards, we perform a "standard workflow" on the integrated data, involving scaling, a new PCA, finding neighbors, and clustering.

Load up libraries
```{r}
library(Seurat)
library(ggplot2)
library(dplyr)

options(future.globals.maxSize=1000000000000000)
```
Setting up libraries for integration
```{r}
lib.list <- readRDS("/Users/gagled01/morganLab/single-cell/5TGM1_Stromal/5TGM1_Code/QC_Filtered_Libs.rds")

lib.list <- c(lib1, lib2, lib3, lib5, lib6)

# Normalize and find variable features for each lib
lib.list <- lapply(X = lib.list, FUN = function(x) {
    x <- NormalizeData(x) 
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
})

# select features that are repeatedly variable across datasets for integration run PCA on each
# dataset using these features
features <- SelectIntegrationFeatures(object.list = lib.list)
lib.list <- lapply(X = lib.list, FUN = function(x) {
    x <- ScaleData(x, features = features, verbose = FALSE)
    x <- RunPCA(x, features = features, verbose = FALSE)
})
```
Find anchors
```{r}
lib.anchors <- FindIntegrationAnchors(object.list = lib.list, anchor.features = features, reduction = "cca")
```
Integrate data
```{r}
# This command creates an 'integrated' data assay
lib.combined <- IntegrateData(anchorset = lib.anchors)
```
Create new variable and dump old
```{r}
mouseBM <- lib.combined

rm(lib.combined, lib1, lib2, lib3, lib4, lib5, lib6)
```
Standard workflow time
```{r}
DefaultAssay(mouseBM) <- "integrated"

mouseBM <- ScaleData(mouseBM)
mouseBM <- RunPCA(mouseBM, npcs = 30)

ElbowPlot(
  object = mouseBM, 
  ndims = 30
) +
  geom_abline(
    aes(intercept = 1.75, slope = 0, color = "red"),
    show.legend = FALSE
  )

mouseBM <- FindNeighbors(mouseBM, dims = 1:30)
mouseBM <- FindClusters(mouseBM, resolution = 0.2)
mouseBM <- RunUMAP(mouseBM, dims = 1:30)
DimPlot(mouseBM, label = T, group.by = "integrated_snn_res.0.2")
```
Save out object
```{r}
saveRDS(lib.combined, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/5TGM1_IntegratedObject_StandardWorkflow.rds")
```
