---
title: "Mouse5TG_FinalFigures"
output: html_document
date: "2023-05-01"
---
################################################################################
Code for generating main figures.

Figures 1 and 2 are either conceptual or show flow cyto/mouse experimental data and, as such, do not involve scRNA analysis and are not included here.

Figure 3 is the overall (all cell types and relative abundance) figure

Figure 4 is the MSCOLC result figure

Figure 5 is the BMEC result figure

################################################################################

Load libraries
```{r}
library(Seurat)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ComplexHeatmap)
library(grid)
library(gridExtra)
library(patchwork)
library(flextable)
```
Load in stromal data
```{r}
stromals <- readRDS("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/objects/Rerun_Mouse5TG_StromalsOnly_AnnotatedObject6_21PCs_NoLib_FIXEDLIBRARIES_6.rds")
```
################################################################################
################################ FIGURE 3 ######################################
################################################################################
Figure 3A - UMAP
```{r}
library(scales)
show_col(hue_pal()(7))
# Without cell count legend
Idents(stromals) <- "celltypes_ECsubbed"
figure.3a <- DimPlot(stromals, label = T,  cols = c("#F8766D", "#C49A00", "#53B400", "#00C094", "#00B6EB", "#A58AFF", "#FB61D7")) + NoLegend()
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/StromalCells_UMAP.png", height = 6, width = 6)
# With cell count legend
DimPlot(stromals) + scale_color_manual(labels = c("Sinusoidal ECs (n = 1,881)", "Fibroblasts (n = 3,363)", "MSCs (n = 3,030)", "OLCs (n = 1,248)", "Pericytes (n = 1,615)",
                                                  "Arterial ECs (n = 2,961)", "Chondrocytes (n = 121)"),
                                       values = c("#F8766D", "#C49A00", "#53B400", "#00C094", "#00B6EB", "#A58AFF", "#FB61D7"))
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/StromalCells_UMAP_WithCellCounts.png", height = 6, width = 8)
```
Figure 3B - Relative Abundance of Stromal Pops
```{r fig.height = 4, fig.width = 10}
#healthy <- subset(stromals, subset = case.control == "Healthy")
#myeloma <- subset(stromals, subset = case.control == "Myeloma")

healthy.da <- data.frame(table(healthy@meta.data$celltypes_ECsubbed)/sum(table(healthy@meta.data$celltypes_ECsubbed)))
healthy.da$condition <- "Healthy"
myeloma.da <- data.frame(table(myeloma@meta.data$celltypes_ECsubbed)/sum(table(myeloma@meta.data$celltypes_ECsubbed)))
myeloma.da$condition <- "Myeloma"
combined.da <- rbind(healthy.da, myeloma.da)

library(reshape2)
melted.da.df <- melt(combined.da, id.vars = c("Var1", "condition"))
colnames(melted.da.df) <- c("Celltype", "Condition", "Balls", "RA")
melted.da.df$Celltype <- factor(melted.da.df$Celltype)
levels(melted.da.df$Celltype)
melted.da.df$Celltype <- factor(melted.da.df$Celltype, levels = c("MSCs", "OLCs", "Fibroblasts", "Chondrocytes", "Arterial ECs", "Sinusoidal ECs", "Pericytes"))

# Grouped bar
figure.3b <- ggplot(melted.da.df, aes(fill = Condition, y = RA, x = Celltype)) + geom_bar(position = "dodge", stat = "identity") +
  ylab("") + xlab("") + 
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(axis.title.y = element_text(size = 15),
        axis.text.x = element_text(size = 12)) +
  scale_x_discrete(labels = c("MSCs", "OLCs", "Fibroblasts", "Chondrocytes", "AECs", "SECs", "Pericytes"))
```
Figure 3C - pseudosampled relative abundances
This needs to be updated with more recent pseudotime analysis
```{r}
storage.df <- data.frame(matrix(ncol = 9, nrow = 0))
colnames(storage.df) <- c("cell.ID", "case.control", "MSCs", "OLCs", "Fibroblasts", "Chondrocytes", "Arterial ECs", "Sinusoidal ECs", "Pericytes")

# Loop for subsampling where
# seurat_obj = the object you are subsampling from
# case.control = the string of the condition you will populate your output dataframe with
# iterations = the number of subsamples to generate

subsample_loop <- function(seurat_obj, case.control, iterations) {
  
  # Dataframe to store values
  storage.df <- data.frame(matrix(ncol = 8, nrow = iterations))
  colnames(storage.df) <- c("case.control", "MSCs", "OLCs", "Fibroblasts", "Chondrocytes", "AECs", "SECs", "Pericytes")
  
  # Random downsampling loop
  for (k in 1:iterations) {
    
    # Do the subsampling without replacement to 50% of original size
    subsample <- slice_sample(seurat_obj@meta.data, prop = 0.05) %>% select(celltypes_ECsubbed)
    sum_cells <- ncol(healthy) # get total cell count
    
    # Storage for cell counts
    msc.count <- 0
    olc.count <- 0
    fibroblast.count <- 0
    chondrocyte.count <- 0
    arterialEC.count <- 0
    sinusoidalEC.count <- 0
    pericyte.count <- 0
  
    # Looping through subsample to get cell counts
    for (i in 1:nrow(subsample)) {
      cell.type <- as.vector(subsample$celltypes_ECsubbed[i]) # Get cell type
      
      if (cell.type == "MSCs") {
        msc.count <- msc.count + 1
      } else if (cell.type == "OLCs") {
        olc.count <- olc.count + 1
      } else if (cell.type == "Fibroblasts") {
        fibroblast.count <- fibroblast.count + 1
      } else if (cell.type == "Chondrocytes") {
        chondrocyte.count <- chondrocyte.count + 1
      } else if (cell.type == "Arterial ECs") {
        arterialEC.count <- arterialEC.count + 1
      } else if (cell.type == "Sinusoidal ECs") {
        sinusoidalEC.count <- sinusoidalEC.count + 1
      } else if (cell.type == "Pericytes") {
        pericyte.count <- pericyte.count + 1
      }
    } 
    
    # Add results to storage df
    storage.df$case.control[k] <- case.control
    storage.df$MSCs[k] <- msc.count/nrow(subsample)
    storage.df$OLCs[k] <- olc.count/nrow(subsample)
    storage.df$Fibroblasts[k] <- fibroblast.count/nrow(subsample)
    storage.df$Chondrocytes[k] <- chondrocyte.count/nrow(subsample)
    storage.df$aECs[k] <- arterialEC.count/nrow(subsample)
    storage.df$sECs[k] <- sinusoidalEC.count/nrow(subsample)
    storage.df$Pericytes[k] <- pericyte.count/nrow(subsample)
  }
  return(storage.df)
}

# Run the function
set.seed(99)

healthy.pseudosamples <- subsample_loop(healthy, "Healthy", 5)
myeloma.pseudosamples <- subsample_loop(myeloma, "Myeloma", 5)
```
Pseudosampling continued
```{r fig.height = 3, fig.width = 10}
# Combine dataframes
combined.pseudosamples <- rbind(healthy.pseudosamples, myeloma.pseudosamples)

# Set up dataframe for plotting
pseudosamples.melt <- melt(combined.pseudosamples, id = c("case.control"))
colnames(pseudosamples.melt)[2] <- "celltype"
colnames(pseudosamples.melt)[1] <- "Condition"
head(pseudosamples.melt)

# Boxplot
msc.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "MSCs",], aes(x = celltype, y = value, fill = Condition)) + geom_boxplot() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(size=12)) + NoLegend()
olc.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "OLCs",], aes(x = celltype, y = value, fill = Condition)) + geom_boxplot() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(size=12)) + NoLegend()
fibro.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "Fibroblasts",], aes(x = celltype, y = value, fill = Condition)) + geom_boxplot() +
  theme_minimal() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(size=12)) + NoLegend()
chondro.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "Chondrocytes",], aes(x = celltype, y = value, fill = Condition)) + geom_boxplot() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(size=12)) + NoLegend()
aEC.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "aECs",], aes(x = celltype, y = value, fill = Condition)) + geom_boxplot() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(size=12)) + NoLegend()
sEC.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "sECs",], aes(x = celltype, y = value, fill = Condition)) + geom_boxplot() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(size=12)) + NoLegend()
peri.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "Pericytes",], aes(x = celltype, y = value, fill = Condition)) + geom_boxplot() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(size=12)) + NoLegend()

# Patchwork trying
figure.3c <- msc.box + olc.box + fibro.box + chondro.box + aEC.box + sEC.box + peri.box + plot_layout(ncol = 7, guides = "collect")
gt <- patchwork::patchworkGrob(pseudo.boxplots)
gt.fixed <- gridExtra::grid.arrange(gt, left = "Relative Abundance", bottom = "Cell Type")
```
Supplementary Figure 3D and E - Violin and Feature Plots of Main Marker Genes
```{r}
p1 <- VlnPlot(stromals, features = c("Lepr")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p2 <- VlnPlot(stromals, features = c("Bglap")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p3 <- VlnPlot(stromals, features = c("Fn1")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p4 <- VlnPlot(stromals, features = c("Sox9")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p5 <- VlnPlot(stromals, features = c("Acta2")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p6 <- VlnPlot(stromals, features = c("Cdh5")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()

figure.3d <- p1 | p2 | p3 | p4 | p5 | p6
figure.3d <- p1 / p2 / p3 / p4 / p5 / p6

p1.ft <- FeaturePlot(stromals, features = c("Lepr")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p2.ft <- FeaturePlot(stromals, features = c("Bglap")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p3.ft <- FeaturePlot(stromals, features = c("Fn1")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p4.ft <- FeaturePlot(stromals, features = c("Sox9")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p5.ft <- FeaturePlot(stromals, features = c("Acta2")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p6.ft <- FeaturePlot(stromals, features = c("Cdh5")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()

figure.3e <- p1.ft | p2.ft | p3.ft | p4.ft | p5.ft | p6.ft  
figure.3e <- (p1.ft | p2.ft | p3.ft) / (p4.ft | p5.ft | p6.ft)  
```
# Figure 3 Patchwork
```{r fig.height = 13, fig.width = 22}
layout <- "
AAABBBB
AAABBBB
AAABBBB
CCCC###
CCCC###
CCCC###
"

figure.3a + patchwork.pie + figure.3e +
  plot_layout(design = layout)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Figure3.png", height = 6, width = 10)
```
# Figure 3 GSEA Table
```{r}
fig3.genesets <- c("Interferon Gamma Response", "Interferon Alpha Response", "Oxidative Phosphorylation", "MYC Targets")
fig3.nes <- c("1.51", "1.47", "1.43", "1.41")
fig3.fdrqval <- c("0.000", "0.005", "0.013", "0.023")
fig3.df <- data.frame(fig3.genesets, fig3.nes, fig3.fdrqval)
colnames(fig3.df) <- c("Gene Set", "NES", "FDR q-val")

flextable(fig3.df) |>
  autofit()
```
################################################################################
################################ FIGURE 4 ######################################
################################################################################
Figure 4 - MSC/OLCs
A = BioRender cartoon - no analysis needed
B = Condition Split UMAP
C = Condition Split Pseudotime
D = Feature plots of Differentiation genes = Lepr, Lpl, Adipoq, Wif1, Spp1 (Opn), Bglap
E = Violin plots of differentiation genes
F = GSEA results

Load MSCOLCs
```{r}
mscolcs <- readRDS("/Users/gagled01/morganLab/single-cell/5TGM1_Stromal/objects/5TGM1_MSCOLC_Object.rds")
```
Figure 4B - MSC/OLC Condition Split UMAP
```{r}
figure.4b <- DimPlot(mscolcs, split.by = "case.control", cols = c("#F8766D", "#B79F00", "#00BA38", "#00BFC4", "#619CFF", "#F564E3"))
```
Figure 4C - MSC/OLC Condition Split Pseudotime UMAP
```{r}
cds <- as.cell_data_set(mscolcs)
cds <- estimate_size_factors(cds)

# Assign partitions
recreate.partitions <- c(rep(1, length(cds@colData@rownames)))
names(recreate.partitions) <- cds@colData@rownames
recreate.partitions <- as.factor(recreate.partitions)
cds@clusters@listData[["UMAP"]][["partitions"]] <- recreate.partitions

# Assign cluster info
Idents(subtype) <- "seurat_clusters"
list.cluster <- subtype@active.ident
cds@clusters@listData[["UMAP"]][["clusters"]] <- list.cluster

# Assign UMAP coords
cds@int_colData@listData[["reducedDims"]]@listData[["UMAP"]] <- subtype@reductions$umap@cell.embeddings

# Learn trajectory
cds <- learn_graph(cds, use_partition = T)

# Compute pseudotime
#cds <- order_cells(cds)
cds <- order_cells(cds, reduction_method = "UMAP", root_cells = colnames(cds[, clusters(cds) == "1"]))
figure.4c <- plot_cells(cds, color_cells_by = "pseudotime", label_groups_by_cluster = F,
           label_branch_points = F, label_roots = F, label_leaves = F) + facet_grid(. ~ case.control) 
```
Figure 4D - MSC/OLC Feature plots
```{r}
lepr.ft <- FeaturePlot(subtype, features = c("Lepr")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
lpl.ft <- FeaturePlot(subtype, features = c("Lpl")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
adipoq.ft <- FeaturePlot(subtype, features = c("Adipoq")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
wif1.ft <- FeaturePlot(subtype, features = c("Wif1")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
opn.ft <- FeaturePlot(subtype, features = c("Spp1")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + labs(title = "Spp1 (Opn)") + NoLegend()
bglap.ft <- FeaturePlot(subtype, features = c("Bglap")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()

figure.4d <- lepr.ft / lpl.ft / adipoq.ft / wif1.ft / opn.ft / bglap.ft
```
Figure 4E - Violin plots
```{r}
lepr.vln <- VlnPlot(subtype, features = c("Lepr")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
lpl.vln <- VlnPlot(subtype, features = c("Lpl")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                          plot.title = element_text(size=12)) + NoLegend()
adipoq.vln <- VlnPlot(subtype, features = c("Adipoq")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                              plot.title = element_text(size=12)) + NoLegend()
wif1.vln <- VlnPlot(subtype, features = c("Wif1")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                            plot.title = element_text(size=12)) + NoLegend()
opn.vln <- VlnPlot(subtype, features = c("Spp1")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + labs(title = "Spp1 (Opn)") + NoLegend()
bglap.vln <- VlnPlot(subtype, features = c("Bglap")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                              plot.title = element_text(size=12)) + NoLegend()

figure.4e <- lepr.vln / lpl.vln / adipoq.vln / wif1.vln / opn.vln / bglap.vln
```
Figure 4 Patchwork
```{r fig.height = 10, fig.width = 12.5}
layout <- "
  AAAACD
  AAAACD
  AAAACD
  BBBBCD
  BBBBCD
  BBBBCD
  ####CD
  ####CD
  ####CD
"

panel <- figure.4b + figure.4c + figure.4d + figure.4e +
  plot_layout(design = layout)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Figure4_noHeatmap.png", height = 11, width = 11)
```
################################################################################
################################ FIGURE 5 ######################################
################################################################################
Figure 5 - BMECs
A = BioRender cartoon - no analysis needed
B = Condition Split UMAP
C = Condition Split Pseudotime
D = Feature plots of Differentiation genes = Lepr, Lpl, Adipoq, Wif1, Spp1 (Opn), Bglap
E = Violin plots of differentiation genes
F = GSEA results

Load BMECs
```{r}
bmecs <- readRDS("/Users/gagled01/morganLab/single-cell/5TGM1_Stromal/objects/5TGM1_BMEC_Object.rds")
```
Figure 5B - BMEC UMAP
```{r}
figure.5b <- DimPlot(subtype, split.by = "case.control", cols = c("#F8766D", "#B79F00", "#00BA38", "#00BFC4", "#619CFF", "#F564E3"))
```
Figure 5C - BMEC Pseudotime UMAP
```{r}
cds <- as.cell_data_set(bmecs)
cds <- estimate_size_factors(cds)

# Assign partitions
recreate.partitions <- c(rep(1, length(cds@colData@rownames)))
names(recreate.partitions) <- cds@colData@rownames
recreate.partitions <- as.factor(recreate.partitions)
cds@clusters@listData[["UMAP"]][["partitions"]] <- recreate.partitions

# Assign cluster info
Idents(subtype) <- "seurat_clusters"
list.cluster <- subtype@active.ident
cds@clusters@listData[["UMAP"]][["clusters"]] <- list.cluster

# Assign UMAP coords
cds@int_colData@listData[["reducedDims"]]@listData[["UMAP"]] <- subtype@reductions$umap@cell.embeddings

# Learn trajectory
cds <- learn_graph(cds, use_partition = T)

# Compute pseudotime
#cds <- order_cells(cds)
cds <- order_cells(cds, reduction_method = "UMAP", root_cells = colnames(cds[, clusters(cds) == "1"]))
figure.5c <- plot_cells(cds, color_cells_by = "pseudotime", label_groups_by_cluster = F,
           label_branch_points = F, label_roots = F, label_leaves = F) + facet_grid(. ~ case.control) 
```
Figure 5D - Feature plots
```{r}
sca1.ft <- FeaturePlot(subtype, features = c("Ly6a")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + labs(title = "Ly6a (Sca1)") + NoLegend()
cxcl12.ft <- FeaturePlot(subtype, features = c("Cxcl12")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
vegfr4.ft <- FeaturePlot(subtype, features = c("Flt4")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + labs(title = "Flt4 (Vegfr-4)") + NoLegend()
il6st.ft <- FeaturePlot(subtype, features = c("Il6st")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()

figure.5d <- sca1.ft | cxcl12.ft | vegfr4.ft | il6st.ft 
```
Figure 5E - Violin plots
```{r}
sca1.vln <- VlnPlot(subtype, features = c("Ly6a")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + labs(title = "Ly6a (Sca1)") + NoLegend()
cxcl12.vln <- VlnPlot(subtype, features = c("Cxcl12")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
vegfr4.vln <- VlnPlot(subtype, features = c("Flt4")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + labs(title = "Flt4 (Vegfr-4)") + NoLegend()
il6st.vln <- VlnPlot(subtype, features = c("Il6st")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()

figure.5e <- sca1.vln | cxcl12.vln | vegfr4.vln | il6st.vln 
```
Figure 5 Patchwork
```{r fig.height = 5.6, fig.width = 10}
layout <- "
  AAAACD
  AAAACD
  AAAACD
  BBBBCD
  BBBBCD
  BBBBCD
"

figure.5b + figure.5c + figure.5d + figure.5e +
  plot_layout(design = layout)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Figure5_noHeatmap.png", height = 7.5, width = 10)
```


