---
title: "Monocle3_TrajectoryAnalysis"
output: html_document
date: "2023-03-01"
---

# Load libraries
```{r}
library(Seurat)
library(tidyverse)
library(dplyr)
library(monocle3)
library(monocle)
library(ggplot2)
library(ggridges)
library(SeuratWrappers)
library(DESeq2)

options(future.globals.maxSize=1000000000000000) # Set max global size so we don't run out of memory
```
# Load data
```{r}
stromals <- readRDS("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/objects/Rerun_Mouse5TG_StromalsOnly_AnnotatedObject6_21PCs_NoLib_FIXEDLIBRARIES_6.rds")
```
# Subsetting to MSCs and ECs (because 2 diff trajectories)
```{r}
subtype <- subset(stromals, subset = celltypes_ECsubbed %in% c("MSCs", "OLCs"))
```
# standard Seurat workflow on each lineage
```{r}
# MSCs
subtype <- ScaleData(subtype)
subtype <- FindVariableFeatures(subtype, nfeatures = 2000)
subtype <- RunPCA(subtype, npcs = 10)
subtype <- FindNeighbors(subtype, dims = 1:10)
subtype <- FindClusters(subtype, resolution = 0.2)
subtype <- RunUMAP(subtype, dims = 1:10)

subtype <- subset(subtype, subset = seurat_clusters %in% c("0", "1", "2", "3", "4", "5"))

DimPlot(subtype)
#ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/BMEC_bothConditions_UMAP2.png", height = 6, width = 8)

DimPlot(subtype, group.by = "celltypes_ECsubbed")
#ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCLineage_PreMonocle_UMAP.png", height = 6, width = 6)

```
# Convert Seurat to monocle object
```{r}
msc.cds <- as.cell_data_set(subtype)
msc.cds <- estimate_size_factors(msc.cds)
```
# Check partitions
```{r}
plot_cells(msc.cds, show_trajectory_graph = FALSE, color_cells_by = "partition")
```

# Assign partitions
```{r}
# MSCs
recreate.partitions <- c(rep(1, length(msc.cds@colData@rownames)))
names(recreate.partitions) <- msc.cds@colData@rownames
recreate.partitions <- as.factor(recreate.partitions)

msc.cds@clusters@listData[["UMAP"]][["partitions"]] <- recreate.partitions
```
# Assign cluster information
```{r}
# MSCs 
Idents(subtype) <- "seurat_clusters"
list.cluster <- subtype@active.ident
msc.cds@clusters@listData[["UMAP"]][["clusters"]] <- list.cluster
```
# Assign UMAP coords
```{r}
# MSCs
msc.cds@int_colData@listData[["reducedDims"]]@listData[["UMAP"]] <- subtype@reductions$umap@cell.embeddings
```
# Learn trajectory
```{r}
# MSCs
msc.cds <- learn_graph(msc.cds, use_partition = T)
```
# Plot with trajectory
```{r}
# MSCs
plot_cells(msc.cds, color_cells_by = "cluster", label_groups_by_cluster = F,
           label_branch_points = T, label_roots = T, label_leaves = F, 
           group_label_size = 5) + facet_grid(. ~ case.control)
#ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCOLCs_Monocle_TrajectoryUMAP_ConditionSplit.png", height = 5, width = 8)
```
# Pseudotime
```{r}
# MSCs

msc.cds <- order_cells(msc.cds)
msc.cds <- order_cells(msc.cds, reduction_method = "UMAP", root_cells = colnames(msc.cds[, clusters(msc.cds) == "1"]))
plot_cells(msc.cds, color_cells_by = "pseudotime", label_groups_by_cluster = F,
           label_branch_points = F, label_roots = F, label_leaves = F) + facet_grid(. ~ case.control) 

#ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCOLCs_Monocle_TrajectoryUMAP_Pseudotime_ConditionSplit_Cluster1Root.png", height = 5, width = 8)
```
# Select branch
```{r}
branch <- choose_graph_segments(msc.cds)

```

# Finding roots programmatically
```{r}
get_earliest_principal_node <- function(cds, time_bin="2"){
  cell_ids <- which(colData(cds)[, "seurat_clusters"] == time_bin)
  
  closest_vertex <-
  cds@principal_graph_aux[["UMAP"]]$pr_graph_cell_proj_closest_vertex
  closest_vertex <- as.matrix(closest_vertex[colnames(cds), ])
  root_pr_nodes <-
  igraph::V(principal_graph(cds)[["UMAP"]])$name[as.numeric(names
  (which.max(table(closest_vertex[cell_ids,]))))]
  
  root_pr_nodes
}

msc.cds <- order_cells(msc.cds, root_pr_nodes=get_earliest_principal_node(msc.cds))
plot_cells(msc.cds, color_cells_by = "pseudotime", label_groups_by_cluster = F, label_cell_groups = F,
           label_branch_points = F, label_roots = F, label_leaves = F) + facet_grid(. ~ case.control)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/BMECs_ONLY_Monocle_Pseudotime_ConditionSplit_ProgrammaticRootChoose.png", height = 5, width = 8)
```
# Fix gene name thing
```{r}
rowData(msc.cds)$gene_name <- rownames(msc.cds)
rowData(msc.cds)$gene_short_name <- rowData(msc.cds)$gene_name
```
# Plot?
```{r}
"Lepr" %in% row.names(rowData(msc.cds))

plot_cells(msc.cds, genes = c("Bglap", "Lpl", "Lepr", "Car3"),
           label_cell_groups = F,
           show_trajectory_graph = F,
           min_expr = 3)
```
# Find genes by pseudotime
```{r}
cds_pt_res <- graph_test(msc.cds, neighbor_graph = "principal_graph", cores = 4)
cds_pt_res_subset <- na.omit(cds_pt_res)
cds_pt_res_subset <- cds_pt_res_subset[cds_pt_res_subset$p_value < 0.05 & cds_pt_res_subset$status == "OK",]
```

# Plot genes by pseudotime
```{r}
my_genes <- rownames(subset(fData(msc.cds), subset = gene_short_name %in% c("Lepr", "Lpl", "Bglap")))
cds_subset <- msc.cds[my_genes,]

plot_genes_in_pseudotime(msc.cds,
                         min_expr = 0.25)
```

# Genes by pseudotime
```{r}
ciliated_genes <- c("Bglap",
                    "Col1a1",
                    "Sparc",
                    "Bglap2",
                    "Lepr",
                    "Adipoq",
                    "Lpl",
                    "Cebpa")
msc.cds_subset <- msc.cds[rownames(msc.cds) %in% ciliated_genes,]

gene_fits <- fit_models(msc.cds_subset, model_formula_str = "~seurat_clusters")

fit_coefs <- coefficient_table(gene_fits)

cluster_time_terms <- fit_coefs %>% filter(term == "seurat_clusters")

cluster_time_terms <- fit_coefs %>% filter (q_value < 0.05) %>%
         select(gene_id, term, q_value, estimate)

plot_genes_violin(msc.cds_subset, ncol=2) +
      theme(axis.text.x=element_text(angle=45, hjust=1))
```


# Cells ordered by pseudotime
```{r}
# MSCs
msc.cds$monocle3_pseudotime <- pseudotime(msc.cds)
msc.pseudotime.data <- as.data.frame(colData(msc.cds))
ggplot(msc.pseudotime.data, aes(monocle3_pseudotime, reorder(seurat_clusters, monocle3_pseudotime), fill = seurat_clusters)) + geom_boxplot()
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCLineage_Monocle_Pseudotime_OrderedByClusters.png", height = 6, width = 6)

# ECs
ec.cds$monocle3_pseudotime <- pseudotime(ec.cds)
ec.pseudotime.data <- as.data.frame(colData(ec.cds))
ggplot(ec.pseudotime.data, aes(monocle3_pseudotime, reorder(seurat_clusters, monocle3_pseudotime), fill = seurat_clusters)) + geom_boxplot()
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/ECLineage_Monocle_Pseudotime_OrderedByClusters.png", height = 6, width = 6)
```
# Checking some genes
```{r}
# ECs
FeaturePlot(ec.lineage, features = c("Lpl", "Atf3", "Socs3", "Timp4", "Ly6c1", "Vcam1"), ncol = 3)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/ECLineage_SomeSubcluster_Genes.png", height = 6, width = 9)

# Fibroblasts
FeaturePlot(subtype, features = c("Cytl1", "Chad", "Acan"), ncol = 3)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCLineage_FibroblastSubcluster_Genes.png", height = 6, width = 9)

# MSCs
FeaturePlot(subtype, features = c("Lepr", "Wif1", "Flt1", "Atf3", "Btg2", "Cd200r3"), ncol = 3)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCLineage_MSCSubcluster_Genes.png", height = 6, width = 9)

# OLCs
FeaturePlot(subtype, features = c("Bglap", "Col1a1", "Sparc", "Spp1", "Ogn", "Mmp13"), ncol = 3)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCLineage_OLCSubcluster_Genes.png", height = 6, width = 9)
 
```

# Genes that change as a function of pseudotime
```{r}
msc.deg <- graph_test(msc.cds, neighbor_graph = "principal_graph")
msc.deg %>% arrange(q_value) %>% filter(status == "OK") %>% head()
```

