---
title: "Mouse5TG_FinalFigures"
output: html_document
date: "2023-05-01"
---

# Load libraries
```{r}
library(Seurat)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(grid)
library(gridExtra)
library(patchwork)
library(flextable)
```
# Load in stromal data
```{r}
stromals <- readRDS("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/objects/Rerun_Mouse5TG_StromalsOnly_AnnotatedObject6_21PCs_NoLib_FIXEDLIBRARIES_6.rds")
```
# Subset celltype
```{r}
subtype <- subset(stromals, subset = celltypes_ECsubbed %in% c("Arterial ECs", "Sinusoidal ECs"))
```
# Standard workflow
# Doing this because we want the most variable genes for the fibroblasts specifically, not for the entire dataset
```{r}
subtype <- ScaleData(subtype)
subtype <- FindVariableFeatures(subtype, nfeatures = 2000)
subtype <- RunPCA(subtype, npcs = 7)
subtype <- FindNeighbors(subtype, dims = 1:7)
subtype <- FindClusters(subtype, resolution = 0.15)
subtype <- RunUMAP(subtype, dims = 1:7)
DimPlot(subtype)

#subtype <- subset(subtype, subset = seurat_clusters %in% c("0", "1", "2", "3", "4", "5"))
DimPlot(subtype)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/BMEC_Subclustered_UMAP_ABL.png", height = 5, width = 5)
```
# Figure 3A
```{r}
# A - UMAP
library(scales)
show_col(hue_pal()(7))
subtype <- stromals
# Without cell count legend
Idents(stromals) <- "celltypes_ECsubbed"
figure.3a <- DimPlot(stromals, label = T,  cols = c("#F8766D", "#C49A00", "#53B400", "#00C094", "#00B6EB", "#A58AFF", "#FB61D7")) + NoLegend()
#ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/StromalCells_UMAP.png", height = 6, width = 6)
# With cell count legend
DimPlot(stromals) + scale_color_manual(labels = c("Sinusoidal ECs (n = 1,881)", "Fibroblasts (n = 3,363)", "MSCs (n = 3,030)", "OLCs (n = 1,248)", "Pericytes (n = 1,615)",
                                                  "Arterial ECs (n = 2,961)", "Chondrocytes (n = 121)"),
                                       values = c("#F8766D", "#C49A00", "#53B400", "#00C094", "#00B6EB", "#A58AFF", "#FB61D7"))
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/StromalCells_UMAP_WithCellCounts.png", height = 6, width = 8)
```
Figure 3B - Relative Abundance of Stromal Pops
```{r fig.height = 4, fig.width = 10}
# B - Relative abundances
#healthy <- subset(stromals, subset = case.control == "Healthy")
#myeloma <- subset(stromals, subset = case.control == "Myeloma")

healthy.da <- data.frame(table(healthy@meta.data$celltypes_ECsubbed)/sum(table(healthy@meta.data$celltypes_ECsubbed)))
healthy.da$condition <- "Healthy"
myeloma.da <- data.frame(table(myeloma@meta.data$celltypes_ECsubbed)/sum(table(myeloma@meta.data$celltypes_ECsubbed)))
myeloma.da$condition <- "Myeloma"
combined.da <- rbind(healthy.da, myeloma.da)

library(reshape2)
melted.da.df <- melt(combined.da, id.vars = c("Var1", "condition"))
colnames(melted.da.df) <- c("Celltype", "Condition", "Balls", "RA")
melted.da.df$Celltype <- factor(melted.da.df$Celltype)
levels(melted.da.df$Celltype)
melted.da.df$Celltype <- factor(melted.da.df$Celltype, levels = c("MSCs", "OLCs", "Fibroblasts", "Chondrocytes", "Arterial ECs", "Sinusoidal ECs", "Pericytes"))

# Grouped bar
figure.3b <- ggplot(melted.da.df, aes(fill = Condition, y = RA, x = Celltype)) + geom_bar(position = "dodge", stat = "identity") +
  ylab("") + xlab("") + 
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(axis.title.y = element_text(size = 15),
        axis.text.x = element_text(size = 12)) +
  scale_x_discrete(labels = c("MSCs", "OLCs", "Fibroblasts", "Chondrocytes", "AECs", "SECs", "Pericytes"))
```
Figure 3C - pseudosampled relative abundances
# Create storage df
```{r}
storage.df <- data.frame(matrix(ncol = 9, nrow = 0))
colnames(storage.df) <- c("cell.ID", "case.control", "MSCs", "OLCs", "Fibroblasts", "Chondrocytes", "Arterial ECs", "Sinusoidal ECs", "Pericytes")

# Loop for subsampling where
# seurat_obj = the object you are subsampling from
# case.control = the string of the condition you will populate your output dataframe with
# iterations = the number of subsamples to generate

subsample_loop <- function(seurat_obj, case.control, iterations) {
  
  # Dataframe to store values
  storage.df <- data.frame(matrix(ncol = 8, nrow = iterations))
  colnames(storage.df) <- c("case.control", "MSCs", "OLCs", "Fibroblasts", "Chondrocytes", "AECs", "SECs", "Pericytes")
  
  # Random downsampling loop
  for (k in 1:iterations) {
    
    # Do the subsampling without replacement to 50% of original size
    subsample <- slice_sample(seurat_obj@meta.data, prop = 0.05) %>% select(celltypes_ECsubbed)
    sum_cells <- ncol(healthy) # get total cell count
    
    # Storage for cell counts
    msc.count <- 0
    olc.count <- 0
    fibroblast.count <- 0
    chondrocyte.count <- 0
    arterialEC.count <- 0
    sinusoidalEC.count <- 0
    pericyte.count <- 0
  
    # Looping through subsample to get cell counts
    for (i in 1:nrow(subsample)) {
      cell.type <- as.vector(subsample$celltypes_ECsubbed[i]) # Get cell type
      
      if (cell.type == "MSCs") {
        msc.count <- msc.count + 1
      } else if (cell.type == "OLCs") {
        olc.count <- olc.count + 1
      } else if (cell.type == "Fibroblasts") {
        fibroblast.count <- fibroblast.count + 1
      } else if (cell.type == "Chondrocytes") {
        chondrocyte.count <- chondrocyte.count + 1
      } else if (cell.type == "Arterial ECs") {
        arterialEC.count <- arterialEC.count + 1
      } else if (cell.type == "Sinusoidal ECs") {
        sinusoidalEC.count <- sinusoidalEC.count + 1
      } else if (cell.type == "Pericytes") {
        pericyte.count <- pericyte.count + 1
      }
    } 
    
    # Add results to storage df
    storage.df$case.control[k] <- case.control
    storage.df$MSCs[k] <- msc.count/nrow(subsample)
    storage.df$OLCs[k] <- olc.count/nrow(subsample)
    storage.df$Fibroblasts[k] <- fibroblast.count/nrow(subsample)
    storage.df$Chondrocytes[k] <- chondrocyte.count/nrow(subsample)
    storage.df$aECs[k] <- arterialEC.count/nrow(subsample)
    storage.df$sECs[k] <- sinusoidalEC.count/nrow(subsample)
    storage.df$Pericytes[k] <- pericyte.count/nrow(subsample)
  }
  return(storage.df)
}

# Run the function
set.seed(99)

healthy.pseudosamples <- subsample_loop(healthy, "Healthy", 5)
myeloma.pseudosamples <- subsample_loop(myeloma, "Myeloma", 5)
```

```{r fig.height = 3, fig.width = 10}
# Combine dataframes
combined.pseudosamples <- rbind(healthy.pseudosamples, myeloma.pseudosamples)

# Set up dataframe for plotting
pseudosamples.melt <- melt(combined.pseudosamples, id = c("case.control"))
colnames(pseudosamples.melt)[2] <- "celltype"
colnames(pseudosamples.melt)[1] <- "Condition"
head(pseudosamples.melt)

# Boxplot
msc.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "MSCs",], aes(x = celltype, y = value, fill = Condition)) + geom_boxplot() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(size=12)) + NoLegend()
olc.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "OLCs",], aes(x = celltype, y = value, fill = Condition)) + geom_boxplot() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(size=12)) + NoLegend()
fibro.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "Fibroblasts",], aes(x = celltype, y = value, fill = Condition)) + geom_boxplot() +
  theme_minimal() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(size=12)) + NoLegend()
chondro.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "Chondrocytes",], aes(x = celltype, y = value, fill = Condition)) + geom_boxplot() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(size=12)) + NoLegend()
aEC.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "aECs",], aes(x = celltype, y = value, fill = Condition)) + geom_boxplot() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(size=12)) + NoLegend()
sEC.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "sECs",], aes(x = celltype, y = value, fill = Condition)) + geom_boxplot() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(size=12)) + NoLegend()
peri.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "Pericytes",], aes(x = celltype, y = value, fill = Condition)) + geom_boxplot() +
  theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.x=element_text(size=12)) + NoLegend()

# Patchwork trying
figure.3c <- msc.box + olc.box + fibro.box + chondro.box + aEC.box + sEC.box + peri.box + plot_layout(ncol = 7, guides = "collect")
gt <- patchwork::patchworkGrob(pseudo.boxplots)
gt.fixed <- gridExtra::grid.arrange(gt, left = "Relative Abundance", bottom = "Cell Type")
#ggsave(plot = gt.fixed, filename = "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/CellTypeRelativeAbundance_Boxplot_Condition_Pseudosampled_05PercDownSample.png", width = 12, height = 3)
```
# Supplementary Figure 3D and E - Violin and Feature Plots of Main Marker Genes
```{r}
# D - Violin Plots
library(cowplot)
p1 <- VlnPlot(stromals, features = c("Lepr")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p2 <- VlnPlot(stromals, features = c("Bglap")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p3 <- VlnPlot(stromals, features = c("Fn1")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p4 <- VlnPlot(stromals, features = c("Sox9")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p5 <- VlnPlot(stromals, features = c("Acta2")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p6 <- VlnPlot(stromals, features = c("Cdh5")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
#cowplot::plot_grid(p1, p2, p3, p4, p5, p6,
#                   nrow = 6,
#                   labels = "auto")

figure.3d <- p1 | p2 | p3 | p4 | p5 | p6
figure.3d <- p1 / p2 / p3 / p4 / p5 / p6
#ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/StromalMarkerGene_VlnPlots_Tall.png", height = 18, width = 4)

# E - Feature Plots
p1.ft <- FeaturePlot(stromals, features = c("Lepr")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p2.ft <- FeaturePlot(stromals, features = c("Bglap")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p3.ft <- FeaturePlot(stromals, features = c("Fn1")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p4.ft <- FeaturePlot(stromals, features = c("Sox9")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p5.ft <- FeaturePlot(stromals, features = c("Acta2")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
p6.ft <- FeaturePlot(stromals, features = c("Cdh5")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank()) + NoLegend()
#cowplot::plot_grid(p1.ft, p2.ft, p3.ft, p4.ft, p5.ft, p6.ft,
#                   nrow = 6,
#                   labels = "auto")

figure.3e <- p1.ft | p2.ft | p3.ft | p4.ft | p5.ft | p6.ft  
figure.3e <- (p1.ft | p2.ft | p3.ft) / (p4.ft | p5.ft | p6.ft)  

#ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/StromalMarkerGene_FeaturePlots_Tall.png", height = 18, width = 4)
```
# Figure 3 Patchwork
```{r fig.height = 13, fig.width = 22}
layout <- "
AAABBBB
AAABBBB
AAABBBB
CCCC###
CCCC###
CCCC###
"

figure.3a + patchwork.pie + figure.3e +
  plot_layout(design = layout)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Figure3.png", height = 6, width = 10)
```
# Figure 3 GSEA Table
```{r}
fig3.genesets <- c("Interferon Gamma Response", "Interferon Alpha Response", "Oxidative Phosphorylation", "MYC Targets")
fig3.nes <- c("1.51", "1.47", "1.43", "1.41")
fig3.fdrqval <- c("0.000", "0.005", "0.013", "0.023")
fig3.df <- data.frame(fig3.genesets, fig3.nes, fig3.fdrqval)
colnames(fig3.df) <- c("Gene Set", "NES", "FDR q-val")

flextable(fig3.df) |>
  autofit()
```
# Figure 4 - MSC/OLCs
```{r}
# A = Complex heatmap
# B = Condition Split UMAP
# C = Condition Split Pseudotime
# D = Feature plots of Differentiation genes = Lepr, Lpl, Adipoq, Wif1, Spp1 (Opn), Bglap
# E = Violin plots of differentiation genes
```
# Figure 4A - MSC/OLC Subcluster Heatmap
```{r}
# Copy and pasted from ComplexHeatmap markdown. Assumes all preprocessing + analysis is complete
heatmap <- Heatmap(matrix.roworder.subset, cluster_rows = F, cluster_columns = TRUE,
        row_order = output_gene_order_sub,
        show_column_dend = F,
        col = col_fun,
        column_names_gp = gpar(fontsize = 0), row_names_gp = gpar(fontsize = 4.5),
        clustering_distance_columns = "euclidean", clustering_method_columns = "average",
        column_dend_height = unit(40,"mm"), row_dend_width = unit(40, "mm"),
        top_annotation = annotation,
        heatmap_legend_param = list(title = "Scaled Expression")
        )
library(tidyHeatmap)
```
# Figure 4B - MSC/OLC Condition Split UMAP
```{r}
figure.4b <- DimPlot(subtype, split.by = "case.control", cols = c("#F8766D", "#B79F00", "#00BA38", "#00BFC4", "#619CFF", "#F564E3"))
```
# Figure 4C - MSC/OLC Condition Split Pseudotime UMAP
```{r}
# Copy and pasted from TrajectoryAnalysis markdown. Assumes all preprocessing + analysis is complete
get_earliest_principal_node <- function(cds, time_bin="1"){
  cell_ids <- which(colData(cds)[, "seurat_clusters"] == time_bin)
  
  closest_vertex <-
  cds@principal_graph_aux[["UMAP"]]$pr_graph_cell_proj_closest_vertex
  closest_vertex <- as.matrix(closest_vertex[colnames(cds), ])
  root_pr_nodes <-
  igraph::V(principal_graph(cds)[["UMAP"]])$name[as.numeric(names
  (which.max(table(closest_vertex[cell_ids,]))))]
  
  root_pr_nodes
}

msc.cds <- order_cells(msc.cds, root_pr_nodes=get_earliest_principal_node(msc.cds))
figure.4c <- plot_cells(msc.cds, color_cells_by = "pseudotime", label_groups_by_cluster = F, label_cell_groups = F,
           label_branch_points = F, label_roots = F, label_leaves = F) + facet_grid(. ~ case.control)
#ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCOLCs_ONLY_Monocle_Pseudotime_ConditionSplit_ProgrammaticRootChoose.png", height = 5, width = 8)
```
# Figure 4D - MSC/OLC Feature plots
```{r}
lepr.ft <- FeaturePlot(subtype, features = c("Lepr")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
lpl.ft <- FeaturePlot(subtype, features = c("Lpl")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
adipoq.ft <- FeaturePlot(subtype, features = c("Adipoq")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
wif1.ft <- FeaturePlot(subtype, features = c("Wif1")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
opn.ft <- FeaturePlot(subtype, features = c("Spp1")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + labs(title = "Spp1 (Opn)") + NoLegend()
bglap.ft <- FeaturePlot(subtype, features = c("Bglap")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
### Currently adapted for ABL presentation structure !!! ###
figure.4d <- (lepr.ft | lpl.ft | adipoq.ft | wif1.ft | opn.ft | bglap.ft)
figure.4d
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/MSCOLC_ClusterAssignment_FeaturePlotsHorizontal.png", width = 15, height = 3)
```
# Figure 4E - Violin plots
```{r}
lepr.vln <- VlnPlot(subtype, features = c("Lepr")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
lpl.vln <- VlnPlot(subtype, features = c("Lpl")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                          plot.title = element_text(size=12)) + NoLegend()
adipoq.vln <- VlnPlot(subtype, features = c("Adipoq")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                              plot.title = element_text(size=12)) + NoLegend()
wif1.vln <- VlnPlot(subtype, features = c("Wif1")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                            plot.title = element_text(size=12)) + NoLegend()
opn.vln <- VlnPlot(subtype, features = c("Spp1")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + labs(title = "Spp1 (Opn)") + NoLegend()
bglap.vln <- VlnPlot(subtype, features = c("Bglap")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                              plot.title = element_text(size=12)) + NoLegend()

figure.4e <- (lepr.vln | lpl.vln | adipoq.vln | wif1.vln | opn.vln | bglap.vln)
figure.4e
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/MSCOLC_ClusterAssignment_VlnPlotsHorizontal.png", width = 15, height = 3)
```
# Figure 4 Patchwork
```{r fig.height = 10, fig.width = 12.5}
layout <- "
  AAAACD
  AAAACD
  AAAACD
  BBBBCD
  BBBBCD
  BBBBCD
  ####CD
  ####CD
  ####CD
"

panel <- figure.4b + figure.4c + figure.4d + figure.4e +
  plot_layout(design = layout)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Figure4_noHeatmapEEE.png", height = 11, width = 11)
```
# Figure 5 - BMECs
```{r}
# A = Complex heatmap
# B = Condition Split UMAP
# C = Condition Split Pseudotime
# D = Feature plots of Differentiation genes = Lepr, Lpl, Adipoq, Wif1, Spp1 (Opn), Bglap
# E = Violin plots of differentiation genes
```
# Figure 5B - BMEC UMAP
```{r}
figure.5b <- DimPlot(subtype, split.by = "case.control", cols = c("#F8766D", "#B79F00", "#00BA38", "#00BFC4", "#619CFF", "#F564E3"))
```
# Figure 5C - BMEC Pseudotime UMAP
```{r}
# Ran prelim code in the TrajectoryAnalysis markdown
msc.cds <- order_cells(msc.cds, root_pr_nodes=get_earliest_principal_node(msc.cds))
figure.5c <- plot_cells(msc.cds, color_cells_by = "pseudotime", label_groups_by_cluster = F, label_cell_groups = F,
           label_branch_points = F, label_roots = F, label_leaves = F) + facet_grid(. ~ case.control)
```
# Figure 5D - Feature plots
```{r}
sca1.ft <- FeaturePlot(subtype, features = c("Ly6a")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + labs(title = "Ly6a (Sca1)") + NoLegend()
cxcl12.ft <- FeaturePlot(subtype, features = c("Cxcl12")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
vegfr4.ft <- FeaturePlot(subtype, features = c("Flt4")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + labs(title = "Flt4 (Vegfr-4)") + NoLegend()
il6st.ft <- FeaturePlot(subtype, features = c("Il6st")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()

figure.5d <- sca1.ft | cxcl12.ft | vegfr4.ft | il6st.ft 
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/BMEC_FeaturePlots_ABL.png", height = 3, width = 12)

DimPlot(subtype)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/BMEC_UMAP_ABL.png", height = 5, width = 5)

```
# Figure 5E - Violin plots
```{r}
sca1.vln <- VlnPlot(subtype, features = c("Ly6a")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + labs(title = "Ly6a (Sca1)") + NoLegend()
cxcl12.vln <- VlnPlot(subtype, features = c("Cxcl12")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()
vegfr4.vln <- VlnPlot(subtype, features = c("Flt4")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + labs(title = "Flt4 (Vegfr-4)") + NoLegend()
il6st.vln <- VlnPlot(subtype, features = c("Il6st")) + theme(axis.title.x = element_blank(),
                                                             axis.title.y = element_blank(),
                                                           plot.title = element_text(size=12)) + NoLegend()

figure.5e <- sca1.vln | cxcl12.vln | vegfr4.vln | il6st.vln 
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/BMEC_ViolinPlots_ABL.png", height = 3, width = 12)
```
# Figure 5 Patchwork
```{r fig.height = 5.6, fig.width = 10}
layout <- "
  AAAACD
  AAAACD
  AAAACD
  BBBBCD
  BBBBCD
  BBBBCD
"

# layout <- "
#   AAAAAABBBBBB
#   AAAAAABBBBBB
#   AAAAAABBBBBB
#   ######CCCCCC
#   ######DDDDDD
# "

figure.5b + figure.5c + figure.5d + figure.5e +
  plot_layout(design = layout)
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Figure5_noHeatmap.png", height = 7.5, width = 10)
```
#
```{r fig.height = 8, fig.width = 12.5}
Heatmap(matrix.roworder.subset, cluster_rows = F, cluster_columns = TRUE,
        row_order = output_gene_order_sub,
        show_column_dend = F,
        col = col_fun,
        column_names_gp = gpar(fontsize = 0), row_names_gp = gpar(fontsize = 4.5),
        clustering_distance_columns = "euclidean", clustering_method_columns = "average",
        column_dend_height = unit(40,"mm"), row_dend_width = unit(40, "mm"),
        top_annotation = annotation,
        heatmap_legend_param = list(title = "Scaled Expression")
        )
```

```{r}
# D - Violin Plots
library(cowplot)
p1 <- VlnPlot(stromals, features = c("Lepr")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + NoLegend()
p2 <- VlnPlot(stromals, features = c("Bglap")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + NoLegend()
p3 <- VlnPlot(stromals, features = c("Fn1")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + NoLegend()
p4 <- VlnPlot(stromals, features = c("Sox9")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + NoLegend()
p5 <- VlnPlot(stromals, features = c("Acta2")) + theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + NoLegend()
p6 <- VlnPlot(stromals, features = c("Cdh5")) + theme(axis.title.x = element_blank())+ NoLegend()
cowplot::plot_grid(p1, p2, p3, p4, p5, p6,
                   nrow = 6,
                   labels = "auto")
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/StromalMarkerGene_VlnPlots_Tall.png", height = 18, width = 4)

# E - Feature Plots
p1.ft <- FeaturePlot(stromals, features = c("Lepr")) + theme(axis.title.x = element_blank()) + NoLegend()
p2.ft <- FeaturePlot(stromals, features = c("Bglap")) + theme(axis.title.x = element_blank()) + NoLegend()
p3.ft <- FeaturePlot(stromals, features = c("Fn1")) + theme(axis.title.x = element_blank()) + NoLegend()
p4.ft <- FeaturePlot(stromals, features = c("Sox9")) + theme(axis.title.x = element_blank()) + NoLegend()
p5.ft <- FeaturePlot(stromals, features = c("Acta2")) + theme(axis.title.x = element_blank()) + NoLegend()
p6.ft <- FeaturePlot(stromals, features = c("Cdh5")) + NoLegend()
cowplot::plot_grid(p1.ft, p2.ft, p3.ft, p4.ft, p5.ft, p6.ft,
                   nrow = 6,
                   labels = "auto")
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/StromalMarkerGene_FeaturePlots_Tall.png", height = 18, width = 4)

VlnPlot(stromals, features = c("Lepr", "Bglap", "Fn1", "Sox9", "Acta2", "Cdh5"), ncol = 1)

```
# Marker Genes of Main Stromal Populations - Supplementary Figure 3A Violin Plots
```{r}
Idents(stromals) <- "celltypes_ECsubbed"
stromals <- RenameIdents(stromals, "Sinusoidal ECs" = "SECs")
stromals <- RenameIdents(stromals, "Arterial ECs" = "AECs")
stromals@meta.data$celltypes_ECsubbed2 <- Idents(stromals)
Idents(stromals) <- "celltypes_ECsubbed2"
# MSCs
msc.markers1 <- VlnPlot(stromals, features = c("Lepr")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20), axis.text.x = element_blank()) + ylab("MSCs") + NoLegend()
msc.markers2 <- VlnPlot(stromals, features = c("Adipoq")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_blank()) + NoLegend()
msc.markers3 <- VlnPlot(stromals, features = c("Cxcl12")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_blank()) + NoLegend()

# OLCs
olc.markers1 <- VlnPlot(stromals, features = c("Bglap")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20), axis.text.x = element_blank()) + ylab("OLCs") + NoLegend()
olc.markers2 <- VlnPlot(stromals, features = c("Spp1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_blank(), plot.title = element_text(hjust = 0.5)) + labs(title = "Spp1 (Opn)") + NoLegend()
olc.markers3 <- VlnPlot(stromals, features = c("Sp7")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_blank()) + NoLegend()

# Fibroblasts
fibro.markers1 <- VlnPlot(stromals, features = c("S100a4")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20), axis.text.x = element_blank()) + ylab("Fibroblasts") + NoLegend()
fibro.markers2 <- VlnPlot(stromals, features = c("Fn1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_blank()) + NoLegend()
fibro.markers3 <- VlnPlot(stromals, features = c("Dcn")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_blank()) + NoLegend()

# Chondrocytes
chondro.markers1 <- VlnPlot(stromals, features = c("Col2a1")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20), axis.text.x = element_blank()) + ylab("Chondrocytes") + NoLegend()
chondro.markers2 <- VlnPlot(stromals, features = c("Sox9")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_blank()) + NoLegend()
chondro.markers3 <- VlnPlot(stromals, features = c("Acan")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_blank()) + NoLegend()

# Pericytes
peri.markers1 <- VlnPlot(stromals, features = c("Acta2")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20), axis.text.x = element_blank()) + ylab("Pericytes") + NoLegend()
peri.markers2 <- VlnPlot(stromals, features = c("Myh11")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_blank()) + NoLegend()
peri.markers3 <- VlnPlot(stromals, features = c("Mcam")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.x = element_blank()) + NoLegend()

# ECs
ec.markers1 <- VlnPlot(stromals, features = c("Cdh5")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20)) + ylab("ECs") + NoLegend()
ec.markers2 <- VlnPlot(stromals, features = c("Cd34")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
ec.markers3 <- VlnPlot(stromals, features = c("Pecam1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

supp3.patchwork <- (msc.markers1 | msc.markers2 | msc.markers3) /
  (olc.markers1 | olc.markers2 | olc.markers3) /
  (fibro.markers1 | fibro.markers2 | fibro.markers3) /
  (chondro.markers1 | chondro.markers2 | chondro.markers3) /
  (peri.markers1 | peri.markers2 | peri.markers3) /
  (ec.markers1 | ec.markers2 | ec.markers3) + theme(plot.tag = element_text(size = 10))

gt <- patchwork::patchworkGrob(supp3.patchwork)
pdf("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MainStromalPopulations_MarkerGene_VlnPlots_SupplementaryFigure3.pdf", width = 15, height = 18)
gridExtra::grid.arrange(gt, left = textGrob("Scaled Expression", gp=gpar(fontsize=24), rot = 90), bottom = textGrob("Subcluster", gp=gpar(fontsize=24)))
dev.off()
```
# Supplementary Figure 4 - Main Stromal Marker Gene Feature Plots
```{r}
# MSCs
msc.markers1 <- FeaturePlot(stromals, features = c("Lepr")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20)) + ylab("MSCs")+ NoLegend()
msc.markers2 <- FeaturePlot(stromals, features = c("Adipoq")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
msc.markers3 <- FeaturePlot(stromals, features = c("Cxcl12")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

# OLCs
olc.markers1 <- FeaturePlot(stromals, features = c("Bglap")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20)) + ylab("MSCs")  + NoLegend()
olc.markers2 <- FeaturePlot(stromals, features = c("Spp1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), plot.title = element_text(hjust = 0.5)) + labs(title = "Spp1 (Opn)") + NoLegend()
olc.markers3 <- FeaturePlot(stromals, features = c("Sp7")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

# Fibroblasts
fibro.markers1 <- FeaturePlot(stromals, features = c("S100a4")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20)) + ylab("MSCs")  + NoLegend()
fibro.markers2 <- FeaturePlot(stromals, features = c("Fn1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
fibro.markers3 <- FeaturePlot(stromals, features = c("Dcn")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

# Chondrocytes
chondro.markers1 <- FeaturePlot(stromals, features = c("Col2a1")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20)) + ylab("MSCs")  + NoLegend()
chondro.markers2 <- FeaturePlot(stromals, features = c("Sox9")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
chondro.markers3 <- FeaturePlot(stromals, features = c("Acan")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

# Pericytes
peri.markers1 <- FeaturePlot(stromals, features = c("Acta2")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20)) + ylab("MSCs")  + NoLegend()
peri.markers2 <- FeaturePlot(stromals, features = c("Myh11")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
peri.markers3 <- FeaturePlot(stromals, features = c("Mcam")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

# ECs
ec.markers1 <- FeaturePlot(stromals, features = c("Cdh5")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20)) + ylab("MSCs")  + NoLegend()
ec.markers2 <- FeaturePlot(stromals, features = c("Cd34")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
ec.markers3 <- FeaturePlot(stromals, features = c("Pecam1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

supp4.patchwork <- (msc.markers1 | msc.markers2 | msc.markers3) /
  (olc.markers1 | olc.markers2 | olc.markers3) /
  (fibro.markers1 | fibro.markers2 | fibro.markers3) /
  (chondro.markers1 | chondro.markers2 | chondro.markers3) /
  (peri.markers1 | peri.markers2 | peri.markers3) /
  (ec.markers1 | ec.markers2 | ec.markers3) + theme(plot.tag = element_text(size = 8))

gt <- patchwork::patchworkGrob(supp4.patchwork)
pdf("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MainStromalPopulations_MarkerGene_FeaturePlots_SupplementaryFigure4.pdf", width = 12, height = 18)
gridExtra::grid.arrange(gt, left = textGrob("UMAP 1", gp=gpar(fontsize=24), rot = 90), bottom = textGrob("UMAP 2", gp=gpar(fontsize=24)))
dev.off()
```
# Supplementary Figure 5 - MSC/OLC subcluster highlight genes
```{r}
# Cluster 0
cluster0.1 <- VlnPlot(subtype, features = c("Lepr")) + theme(axis.title.x = element_blank(), axis.text.y = element_text(size = 15)) + ylab("Cluster 0") + NoLegend()
cluster0.2 <- VlnPlot(subtype, features = c("Lpl")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
cluster0.3 <- VlnPlot(subtype, features = c("Mgp")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
cluster0.4 <- VlnPlot(subtype, features = c("Bmp6")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

# Cluster 1
cluster1.1 <- VlnPlot(subtype, features = c("Wif1")) + theme(axis.title.x = element_blank(), axis.text.y = element_text(size = 15)) + ylab("Cluster 1") + NoLegend()
cluster1.2 <- VlnPlot(subtype, features = c("Limch1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
cluster1.3 <- VlnPlot(subtype, features = c("Sfrp1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
cluster1.4 <- VlnPlot(subtype, features = c("Tnc")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

# Cluster 2
cluster2.1 <- VlnPlot(subtype, features = c("Spp1")) + theme(axis.title.x = element_blank(), axis.text.y = element_text(size = 15)) + ylab("Cluster 2") + NoLegend()
cluster2.2 <- VlnPlot(subtype, features = c("Mmp13")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
cluster2.3 <- VlnPlot(subtype, features = c("Postn")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
cluster2.4 <- VlnPlot(subtype, features = c("Col6a3")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

# Cluster 3
cluster3.1 <- VlnPlot(subtype, features = c("Bglap")) + theme(axis.title.x = element_blank(), axis.text.y = element_text(size = 15)) + ylab("Cluster 3") + NoLegend()
cluster3.2 <- VlnPlot(subtype, features = c("Bglap2")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
cluster3.3 <- VlnPlot(subtype, features = c("Col11a1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
cluster3.4 <- VlnPlot(subtype, features = c("Car3")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

# Cluster 4
cluster4.1 <- VlnPlot(subtype, features = c("Btg2")) + theme(axis.title.x = element_blank(), axis.text.y = element_text(size = 15)) + ylab("Cluster 4") + NoLegend()
cluster4.2 <- VlnPlot(subtype, features = c("Atf3")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
cluster4.3 <- VlnPlot(subtype, features = c("Zfp36")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
cluster4.4 <- VlnPlot(subtype, features = c("Socs3")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

# Cluster 5
cluster5.1 <- VlnPlot(subtype, features = c("Ppbp")) + theme(axis.title.x = element_blank(), axis.text.y = element_text(size = 15), plot.title = element_text(hjust = 0.5)) + labs(title = "Ppbp (Cxcl7)") + ylab("Cluster 5") + NoLegend()
cluster5.2 <- VlnPlot(subtype, features = c("Pf4")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), plot.title = element_text(hjust = 0.5)) + labs(title = "Pf4 (Cxcl4)") + NoLegend()
cluster5.3 <- VlnPlot(subtype, features = c("Adamts4")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
cluster5.4 <- VlnPlot(subtype, features = c("Treml1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

msc.subcluster.patchwork <- (cluster0.1 | cluster0.2 | cluster0.3 | cluster0.4) /
  (cluster1.1 | cluster1.2 | cluster1.3 | cluster1.4) /
  (cluster2.1 | cluster2.2 | cluster2.3 | cluster2.4) /
  (cluster3.1 | cluster3.2 | cluster3.3 | cluster3.4) /
  (cluster4.1 | cluster4.2 | cluster4.3 | cluster4.4) /
  (cluster5.1 | cluster5.2 | cluster5.3 | cluster5.4) + theme(plot.tag = element_text(size = 8))

gt <- patchwork::patchworkGrob(msc.subcluster.patchwork)
pdf("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCOLCs_Subclusters_VlnPlot_SupplementaryFigure5.pdf", width = 18, height = 18)
gridExtra::grid.arrange(gt, left = textGrob("Scaled Expression", gp=gpar(fontsize=20), rot = 90), bottom = textGrob("Subcluster", gp=gpar(fontsize=20)))
dev.off()
```
# Supplementary figure 6 - BMEC subcluster violin plots
```{r}
bmec1 <- VlnPlot(subtype, features = c("Lpl")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20)) + ylab("Cluster 0") + NoLegend()
bmec2 <- VlnPlot(subtype, features = c("Bmp6")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
bmec3 <- VlnPlot(subtype, features = c("Sparcl1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

bmec4 <- VlnPlot(subtype, features = c("Sparcl1")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20)) + ylab("Cluster 1") + NoLegend()
bmec5 <- VlnPlot(subtype, features = c("Timp4")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
bmec6 <- VlnPlot(subtype, features = c("Bmp6")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

bmec7 <- VlnPlot(subtype, features = c("Fbln5")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20)) + ylab("Cluster 2") + NoLegend()
bmec8 <- VlnPlot(subtype, features = c("Sema3g")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
bmec9 <- VlnPlot(subtype, features = c("Sox17")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

bmec10 <- VlnPlot(subtype, features = c("Atf3")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20)) + ylab("Cluster 3") + NoLegend()
bmec11 <- VlnPlot(subtype, features = c("Btg2")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
bmec12 <- VlnPlot(subtype, features = c("Socs3")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

bmec13 <- VlnPlot(subtype, features = c("Vcam1")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20)) + ylab("Cluster 4") + NoLegend()
bmec14 <- VlnPlot(subtype, features = c("Sema3a")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
bmec15 <- VlnPlot(subtype, features = c("Sele")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

bmec16 <- VlnPlot(subtype, features = c("Cenpf")) + theme(axis.title.x = element_blank(), axis.title.y = element_text(size = 20)) + ylab("Cluster 5") + NoLegend()
bmec17 <- VlnPlot(subtype, features = c("Prc1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
bmec18 <- VlnPlot(subtype, features = c("Knl1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

bmec.subcluster.patchwork <- (bmec1 | bmec2 | bmec3) /
  (bmec4 | bmec5 | bmec6) /
  (bmec7 | bmec8 | bmec9) /
  (bmec10 | bmec11 | bmec12) /
  (bmec13 | bmec14 | bmec15) /
  (bmec16 | bmec17 | bmec18) + theme(plot.tag = element_text(size = 8))

gt <- patchwork::patchworkGrob(bmec.subcluster.patchwork)
pdf("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/BMEC_Subclusters_VlnPlot_SupplementaryFigure6.pdf", width = 12, height = 18)
gridExtra::grid.arrange(gt, left = textGrob("Scaled Expression", gp=gpar(fontsize=20), rot = 90), bottom = textGrob("Subcluster", gp=gpar(fontsize=20)))
dev.off()
```
# Supplementary Figure 5 and 6 = Case/Control Heatmaps
# MSC/OLC Supplementary Figures - Secreted - Supplementary Figure 7
```{r}
p1.secrete <- VlnPlot(subtype, features = c("Cxcl12")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p2.secrete <- VlnPlot(subtype, features = c("Ppbp")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), plot.title = element_text(hjust = 0.5)) + labs(title = "Ppbp (Cxcl7)") + NoLegend()
p3.secrete <- VlnPlot(subtype, features = c("Pf4")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), plot.title = element_text(hjust = 0.5)) + labs(title = "Pf4 (Cxcl4)") + NoLegend()
p4.secrete <- VlnPlot(subtype, features = c("Tnfsf13b")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), plot.title = element_text(hjust = 0.5)) + labs(title = "Tnfsf13b (Baff)") + NoLegend()

p5.secrete <- VlnPlot(subtype, features = c("Bglap")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p6.secrete <- VlnPlot(subtype, features = c("Spp1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), plot.title = element_text(hjust = 0.5)) + labs(title = "Spp1 (Opn)") + NoLegend()
p7.secrete <- VlnPlot(subtype, features = c("Col1a1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p8.secrete <- VlnPlot(subtype, features = c("Alpl")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

p9.secrete <- VlnPlot(subtype, features = c("Lpl")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p10.secrete <- VlnPlot(subtype, features = c("Adipoq")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p11.secrete <- VlnPlot(subtype, features = c("Mgp")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p12.secrete <- VlnPlot(subtype, features = c("Wif1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

p13.secrete <- VlnPlot(subtype, features = c("Fgf7")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p14.secrete <- VlnPlot(subtype, features = c("Igf1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p15.secrete <- VlnPlot(subtype, features = c("Il7")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p16.secrete <- VlnPlot(subtype, features = c("Tnfsf11")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), plot.title = element_text(hjust = 0.5)) + labs(title = "Tnfsf11 (Rankl)") + NoLegend()

p17.secrete <- VlnPlot(subtype, features = c("Tgfb1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p18.secrete <- VlnPlot(subtype, features = c("Tgfb3")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p19.secrete <- VlnPlot(subtype, features = c("Sp7")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p20.secrete <- VlnPlot(subtype, features = c("Vegfa")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

secrete.patchwork <- (p1.secrete | p2.secrete | p3.secrete | p4.secrete) /
  (p5.secrete | p6.secrete | p7.secrete | p8.secrete) /
  (p9.secrete | p10.secrete | p11.secrete | p12.secrete) /
  (p13.secrete | p14.secrete | p15.secrete | p16.secrete) /
  (p17.secrete | p18.secrete | p19.secrete | p20.secrete) + theme(plot.tag = element_text(size = 8))

gt <- patchwork::patchworkGrob(secrete.patchwork)
pdf("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCOLCs_SecretedMolecules_VlnPlot_ConditionSplit_SupplementaryFigure8.pdf", width = 12, height = 15)
gridExtra::grid.arrange(gt, left = textGrob("Scaled Expression", gp=gpar(fontsize=24), rot = 90), bottom = textGrob("Subcluster", gp=gpar(fontsize=24)))
dev.off()
```
# MSC/OLC Supplementary Figures - Receptors - Supplementary Figure 8
```{r}
library(patchwork)
library(grid)

p1.recep <- VlnPlot(subtype, features = c("Lepr")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p2.recep <- VlnPlot(subtype, features = c("Tgfbr1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p3.recep <- VlnPlot(subtype, features = c("Tgfbr2")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p4.recep <- VlnPlot(subtype, features = c("Tgfbr3")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p5.recep <- VlnPlot(subtype, features = c("Tnfrsf1a")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), plot.title = element_text(hjust = 0.5)) + labs(title = "Tnfrsf1a (Tnfr1)") + NoLegend()
p6.recep <- VlnPlot(subtype, features = c("Tnfrsf1b")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), plot.title = element_text(hjust = 0.5)) + labs(title = "Tnfrsf1b (Tnfr2)") + NoLegend()
p7.recep <- VlnPlot(subtype, features = c("Acvr1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p8.recep <- VlnPlot(subtype, features = c("Acvr1b")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p9.recep <- VlnPlot(subtype, features = c("Fgfr1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p10.recep <- VlnPlot(subtype, features = c("Fgfr2")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p11.recep <- VlnPlot(subtype, features = c("Fgfr3")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p12.recep <- VlnPlot(subtype, features = c("Pth1r")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p13.recep <- VlnPlot(subtype, features = c("Bmpr1a")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p14.recep <- VlnPlot(subtype, features = c("Bmpr2")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p15.recep <- VlnPlot(subtype, features = c("Lrp5")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p16.recep <- VlnPlot(subtype, features = c("Lrp6")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p1.adhese <- VlnPlot(subtype, features = c("Vcam1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p2.adhese <- VlnPlot(subtype, features = c("Icam1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p3.adhese <- VlnPlot(subtype, features = c("Jag1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()
p4.adhese <- VlnPlot(subtype, features = c("Notch1")) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + NoLegend()

receptor.patchwork <- (p1.recep | p2.recep | p3.recep | p4.recep) /
  (p5.recep | p6.recep | p7.recep | p8.recep) /
  (p9.recep | p10.recep | p11.recep | p12.recep) /
  (p13.recep | p14.recep | p15.recep | p16.recep) /
  (p1.adhese | p2.adhese | p3.adhese | p4.adhese) + theme(plot.tag = element_text(size = 8))

gt <- patchwork::patchworkGrob(receptor.patchwork)
pdf("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSCOLCs_ReceptorAdhesionMolecules_VlnPlot_ConditionSplit_SupplementaryFigure8.pdf", width = 12, height = 16)
gridExtra::grid.arrange(gt, left = textGrob("Scaled Expression", gp=gpar(fontsize=20), rot = 90), bottom = textGrob("Subcluster", gp=gpar(fontsize=20)))
dev.off()
```
# MSC/OLC Supplementary Figures - Adhesion Molecules - Supplementary Figure 10

# BMEC Supplementary Figure Genes
Secreted = Adamts4, Adamts5, Hif1a, Eno1, Tjp1, Vegfa, Sparc, Il15, Tgfb1, Ly96, Flna, Vim, Ywhah, Ywhaz
Receptors = Flt1, Kdr, Flt4, Tnfrsf1a, Aplnr, Itgav, Tlr3, Tlr4
Adhesion = Selp, Sele, Vcam1, Icam1

# Temptest 
# Csf3 = Gcsf
# Csf2 = Gmcsf
```{r}
# Secreted

# Receptors

# For immune figure
VlnPlot(subtype, features = c("Ccl5", "Ccl6", "Ccl19", "Cxcl9", "Cxcl10"))
```

# Supplementary Figure 12 - BMEC Secreted Molecules
```{r}
secrete1 <- VlnPlot(subtype, features = c("Vcan")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete2 <- VlnPlot(subtype, features = c("Vegfa")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete3 <- VlnPlot(subtype, features = c("Angpt1")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete4 <- VlnPlot(subtype, features = c("Sparc")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete5 <- VlnPlot(subtype, features = c("Tgfb1")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete6 <- VlnPlot(subtype, features = c("Ly96")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete7 <- VlnPlot(subtype, features = c("Eno1")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete8 <- VlnPlot(subtype, features = c("Hif1a")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete9 <- VlnPlot(subtype, features = c("Tjp1")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete10 <- VlnPlot(subtype, features = c("Il15")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete11 <- VlnPlot(subtype, features = c("Bmp4")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete12 <- VlnPlot(subtype, features = c("Egfl7")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete13 <- VlnPlot(subtype, features = c("Flna")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete14 <- VlnPlot(subtype, features = c("Vim")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete15 <- VlnPlot(subtype, features = c("Adamts4")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
secrete16 <- VlnPlot(subtype, features = c("Adamts5")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()

secrete.patchwork <- (secrete1 | secrete2 | secrete3 | secrete4) /
  (secrete5 | secrete6 | secrete7 | secrete8) /
  (secrete9 | secrete10 | secrete11 | secrete12) /
  (secrete13 | secrete14 | secrete15 | secrete16) + theme(plot.tag = element_text(size = 8))

gt <- patchwork::patchworkGrob(secrete.patchwork)
pdf("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/BMEC_SecretedMolecules_VlnPlot_SupplementaryFigure12.pdf", width = 16, height = 16)
gridExtra::grid.arrange(gt, left = textGrob("Scaled Expression", gp=gpar(fontsize=16), rot = 90), bottom = textGrob("Subcluster", gp=gpar(fontsize=16)))
dev.off()
```
# Supplementary Figure 13 - Receptor Molecules
```{r}
receptor1 <- VlnPlot(subtype, features = c("Flt1")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + labs(title = "Flt (Vegfr-1)") + NoLegend()
receptor2 <- VlnPlot(subtype, features = c("Kdr")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + labs(title = "Kdr (Vegfr-2)") + NoLegend()
receptor3 <- VlnPlot(subtype, features = c("Flt4")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + labs(title = "Flt4 (Vegfr-4)") + NoLegend()
receptor4 <- VlnPlot(subtype, features = c("Tnfrsf1a")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + labs(title = "Tnfrsf1a (Tnfr1)") + NoLegend()
receptor5 <- VlnPlot(subtype, features = c("Jag1")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
receptor6 <- VlnPlot(subtype, features = c("Notch1")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
receptor7 <- VlnPlot(subtype, features = c("Ifnar1")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
receptor8 <- VlnPlot(subtype, features = c("Ifngr1")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
receptor9 <- VlnPlot(subtype, features = c("Tlr3")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
receptor10 <- VlnPlot(subtype, features = c("Tlr4")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
receptor11 <- VlnPlot(subtype, features = c("Itgav")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
receptor12 <- VlnPlot(subtype, features = c("Aplnr")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
adhese1 <- VlnPlot(subtype, features = c("Selp")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
adhese2 <- VlnPlot(subtype, features = c("Sele")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
adhese3 <- VlnPlot(subtype, features = c("Vcam1")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
adhese4 <- VlnPlot(subtype, features = c("Icam1")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()

receptor.patchwork <- (receptor1 | receptor2 | receptor3 | receptor4) /
  (receptor5 | receptor6 | receptor7 | receptor8) /
  (receptor9 | receptor10 | receptor11 | receptor12) /
  (adhese1 | adhese2 | adhese3 | adhese4) + theme(plot.tag = element_text(size = 8))

gt <- patchwork::patchworkGrob(receptor.patchwork)
pdf("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/BMEC_ReceptorAdheseMolecules_VlnPlot_SupplementaryFigure13.pdf", width = 16, height = 16)
gridExtra::grid.arrange(gt, left = textGrob("Scaled Expression", gp=gpar(fontsize=16), rot = 90), bottom = textGrob("Subcluster", gp=gpar(fontsize=16)))
dev.off()
```
# Supplementary Figure 14 - Proinflammatory Cytokines in BMECs
```{r}
cytokine1 <- VlnPlot(subtype, features = c("Il1b")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
cytokine2 <- VlnPlot(subtype, features = c("Il6")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
cytokine3 <- VlnPlot(subtype, features = c("Il6st")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
cytokine4 <- VlnPlot(subtype, features = c("Il11")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
cytokine5 <- VlnPlot(subtype, features = c("Il15")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
cytokine6 <- VlnPlot(subtype, features = c("Ccl2")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
cytokine7 <- VlnPlot(subtype, features = c("Ccl5")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
cytokine8 <- VlnPlot(subtype, features = c("Ccl6")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
cytokine9 <- VlnPlot(subtype, features = c("Ccl19")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
cytokine10 <- VlnPlot(subtype, features = c("Cxcl1")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
cytokine11 <- VlnPlot(subtype, features = c("Cxcl9")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()
cytokine12 <- VlnPlot(subtype, features = c("Cxcl10")) + theme(axis.title.x = element_blank(), 
                                                            axis.title.y = element_blank()) + NoLegend()

cytokine.patchwork <- (cytokine1 | cytokine2 | cytokine3 | cytokine4) /
  (cytokine5 | cytokine6 | cytokine7 | cytokine8) /
  (cytokine9 | cytokine10 | cytokine11 | cytokine12) + theme(plot.tag = element_text(size = 8))

gt <- patchwork::patchworkGrob(cytokine.patchwork)
pdf("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/BMEC_ProinflammatoryCytokines_VlnPlot_SupplementaryFigure14.pdf", width = 16, height = 12)
gridExtra::grid.arrange(gt, left = textGrob("Scaled Expression", gp=gpar(fontsize=16), rot = 90), bottom = textGrob("Subcluster", gp=gpar(fontsize=16)))
dev.off()
```


# Supplementary Table 1-3. GSEA Top Gene tables
```{r}
# Function to clean frames
clean_frame <- function(df, drop.cols){
  df <- df[, !names(df) %in% drop.cols]
  df <- head(df, n = 10)
  df$`RUNNING ES` <- round(df$`RUNNING ES`, 4)
  df$PATHWAY <- "ALLOGRAFT REJECTION"
  df <- df %>%
    relocate(PATHWAY)
  df$PATHWAY[2:10] <- ""
  return(df)
}
```
Supplementary Table 1 - All Stromal GSEA Top Genes
```{r}
require(data.table)
allograft <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/allStromal_HALLMARK_ALLOGRAFT_REJECTION.tsv"))
coagulation <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/allStromal_HALLMARK_COAGULATION.tsv"))
myc <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/allStromal_HALLMARK_MYC_TARGETS_V1.tsv"))
oxi.phos <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/allStromal_HALLMARK_OXIDATIVE_PHOSPHORYLATION.tsv"))
angio <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/allStromal_HALLMARK_ANGIOGENESIS.tsv"))
inter.alpha <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/allStromal_HALLMARK_INTERFERON_ALPHA_RESPONSE.tsv"))
inter.gamma <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/allStromal_HALLMARK_INTERFERON_GAMMA_RESPONSE.tsv"))

drop.cols <- c("NAME", "TITLE", "RANK METRIC SCORE", "CORE ENRICHMENT", "V8") # Columns to drop

allo.df <- clean_frame(allograft, drop.cols)
coag.df <- clean_frame(coagulation, drop.cols)
myc.df <- clean_frame(myc, drop.cols)
oxi.df <- clean_frame(oxi.phos, drop.cols)
angio.df <- clean_frame(angio, drop.cols)
interA.df <- clean_frame(inter.alpha, drop.cols)
interG.df <- clean_frame(inter.gamma, drop.cols)

# Combine frames we want for final output
stromal.combined <- rbind(interA.df, interG.df, oxi.df, myc.df)
colnames(stromal.combined) <- c("Pathway", "Gene", "Gene Rank", "Running ES") # Rename columns
table_caption <- c("Top GSEA Genes in MM Stromal Cells", "Supplementary Table 1") # Define captions

# Run it
stromal.table <- stromal.combined |>
  flextable() |>
  add_header_lines(values = rev(table_caption)) |>
  padding(padding = 0, part = "all") |>
  padding(padding.right = 10, part = "all") |>
  autofit() |>
  bold(part = "header", i = 1) |>
  italic(part = "header", i = c(2:length(table_caption))) |>
  align(part = "header", i = c(1:length(table_caption)), align = "left") |>
  border(part = "head", i = c(1:length(table_caption)),
         border = list("width" = 0, color = "black", style = "solid"))
save_as_docx(stromal.table, path = "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/allStromal_GSEA_TopGenes_SupplementaryTable1.docx", align = "center")
```
# Supplementary Table 2 - MSC/OLC GSEA Top Genes
```{r}
require(data.table)
complement <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/MSCOLC_HALLMARK_COMPLEMENT.tsv"))
allograft <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/MSCOLC_HALLMARK_ALLOGRAFT_REJECTION.tsv"))
inter.alpha <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/MSCOLC_HALLMARK_INTERFERON_ALPHA_RESPONSE.tsv"))
inter.gamma <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/MSCOLC_HALLMARK_INTERFERON_GAMMA_RESPONSE.tsv"))
oxi.phos <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/MSCOLC_HALLMARK_OXIDATIVE_PHOSPHORYLATION.tsv"))

drop.cols <- c("NAME", "TITLE", "RANK METRIC SCORE", "CORE ENRICHMENT", "V8") # Columns to drop

comp.df <- clean_frame(complement, drop.cols)
allo.df <- clean_frame(allograft, drop.cols)
interA.df <- clean_frame(inter.alpha, drop.cols)
interG.df <- clean_frame(inter.gamma, drop.cols)
oxi.df <- clean_frame(oxi.phos, drop.cols)

# Combine frames we want for final output
mscolc.combined <- rbind(interA.df, interG.df, oxi.df)
colnames(mscolc.combined) <- c("Pathway", "Gene", "Gene Rank", "Running ES") # Rename columns
table_caption <- c("Top GSEA Genes in MM MSC/OLC", "Supplementary Table 2") # Define captions

# Run it
mscolc.table <- mscolc.combined |>
  flextable() |>
  add_header_lines(values = rev(table_caption)) |>
  padding(padding = 0, part = "all") |>
  padding(padding.right = 10, part = "all") |>
  autofit() |>
  bold(part = "header", i = 1) |>
  italic(part = "header", i = c(2:length(table_caption))) |>
  align(part = "header", i = c(1:length(table_caption)), align = "left") |>
  border(part = "head", i = c(1:length(table_caption)),
         border = list("width" = 0, color = "black", style = "solid"))
save_as_docx(mscolc.table, path = "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/MSCOLC_GSEA_TopGenes_SupplementaryTable2.docx", align = "center")
```
# Supplementary Table 3 - BMEC GSEA Top Genes
```{r}
# Load in tables
require(data.table)
allo.reject <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/BMEC_HALLMARK_ALLOGRAFT_REJECTION.tsv"))
inflam.respo <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/BMEC_HALLMARK_INFLAMMATORY_RESPONSE.tsv"))
inter.alpha <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/BMEC_HALLMARK_INTERFERON_ALPHA_RESPONSE.tsv"))
inter.gamma <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/BMEC_HALLMARK_INTERFERON_GAMMA_RESPONSE.tsv"))
myc.targets <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/BMEC_HALLMARK_MYC_TARGETS_V1.tsv"))
oxi.phos <- as.data.frame(fread("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/BMEC_HALLMARK_OXIDATIVE_PHOSPHORYLATION.tsv"))

drop.cols <- c("NAME", "TITLE", "RANK METRIC SCORE", "CORE ENRICHMENT", "V8") # Columns to drop

allo.df <- clean_frame(allo.reject, drop.cols)
inflam.df <- clean_frame(inflam.respo, drop.cols)
interA.df <- clean_frame(inter.alpha, drop.cols)
interG.df <- clean_frame(inter.gamma, drop.cols)
myc.df <- clean_frame(myc.targets, drop.cols)
oxi.df <- clean_frame(oxi.phos, drop.cols)

# Combine frames we want for final output
bmec.combined <- rbind(interA.df, interG.df, oxi.df, inflam.df)
colnames(bmec.combined) <- c("Pathway", "Gene", "Gene Rank", "Running ES") # Rename columns
table_caption <- c("Top GSEA Genes in MM BMEC", "Supplementary Table 3") # Define captions

# Run it
bmec.table <- bmec.combined |>
  flextable() |>
  add_header_lines(values = rev(table_caption)) |>
  padding(padding = 0, part = "all") |>
  padding(padding.right = 10, part = "all") |>
  autofit() |>
  bold(part = "header", i = 1) |>
  italic(part = "header", i = c(2:length(table_caption))) |>
  align(part = "header", i = c(1:length(table_caption)), align = "left") |>
  border(part = "head", i = c(1:length(table_caption)),
         border = list("width" = 0, color = "black", style = "solid"))
  save_as_docx(bmec.table, path = "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/GSEA/BMEC_GSEA_TopGenes_SupplementaryTable3.docx", align = "center")
show_col(hue_pal()(7))

DimPlot(stromals)
```
# MSCOLC subcluster relative abundances
```{r}
healthy.subtype <- subset(subtype, subset = case.control == "Healthy")
myeloma.subtype <- subset(subtype, subset = case.control == "Myeloma")

healthy.value <- as.numeric(table(healthy.subtype@meta.data$seurat_clusters))[1:6]
myeloma.value <- as.numeric(table(myeloma.subtype@meta.data$seurat_clusters))[1:6]

healthy.frac <- healthy.value/sum(healthy.value)
myeloma.frac <- myeloma.value/sum(myeloma.value)

healthy.labels <- paste(as.character(round(healthy.frac, digits=4)*100), "%", sep="")
myeloma.labels <- paste(as.character(round(myeloma.frac, digits=4)*100), "%", sep="")

healthy.pie <- data.frame(group=c("Cluster 0", "Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5"), 
                          value = healthy.value, 
                          perc = healthy.frac, 
                          labels = healthy.labels)

myeloma.pie <- data.frame(group=c("Cluster 0", "Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5"), 
                          value = myeloma.value, 
                          perc = myeloma.frac, 
                          labels = myeloma.labels)

healthy.pie <- healthy.pie %>%
  arrange(desc(group)) %>%
  mutate(prop = value / sum(healthy.pie$value)*100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)

myeloma.pie <- myeloma.pie %>%
  arrange(desc(group)) %>%
  mutate(prop = value / sum(myeloma.pie$value)*100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)

healthy.pie.plot <- ggplot(healthy.pie, aes(x = "", y = perc, fill = group)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels), color = rep(c("white")),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  guides(fill = guide_legend(title = "")) + 
  scale_fill_manual(values = stromal.colors) +
  coord_polar(theta = "y") +
  ggtitle("Healthy") +
  theme_void() + NoLegend() + theme(plot.title = element_text(hjust = 0.5))

myeloma.pie.plot <- ggplot(myeloma.pie, aes(x = "", y = perc, fill = group)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels), color = rep(c("white")),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  guides(fill = guide_legend(title = "")) + 
  scale_fill_manual(values = stromal.colors) +
  coord_polar(theta = "y") +
  ggtitle("Myeloma") + theme_void() + theme(plot.title = element_text(hjust = 0.5))
  
patchwork.pie <- healthy.pie.plot | myeloma.pie.plot
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/SupplementaryFigure8_MSCOLC_SubclusterPieCharts.png", width = 10, height = 6)
```
#
```{r}
healthy.subtype <- subset(subtype, subset = case.control == "Healthy")
myeloma.subtype <- subset(subtype, subset = case.control == "Myeloma")

healthy.value <- as.numeric(table(healthy.subtype@meta.data$seurat_clusters))[1:6]
myeloma.value <- as.numeric(table(myeloma.subtype@meta.data$seurat_clusters))[1:6]

healthy.frac <- healthy.value/sum(healthy.value)
myeloma.frac <- myeloma.value/sum(myeloma.value)

healthy.labels <- paste(as.character(round(healthy.frac, digits=4)*100), "%", sep="")
myeloma.labels <- paste(as.character(round(myeloma.frac, digits=4)*100), "%", sep="")

healthy.pie <- data.frame(group=c("Cluster 0", "Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5"), 
                          value = healthy.value, 
                          perc = healthy.frac, 
                          labels = healthy.labels)

myeloma.pie <- data.frame(group=c("Cluster 0", "Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5"), 
                          value = myeloma.value, 
                          perc = myeloma.frac, 
                          labels = myeloma.labels)

healthy.pie <- healthy.pie %>%
  arrange(desc(group)) %>%
  mutate(prop = value / sum(healthy.pie$value)*100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)

myeloma.pie <- myeloma.pie %>%
  arrange(desc(group)) %>%
  mutate(prop = value / sum(myeloma.pie$value)*100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)

healthy.pie.plot <- ggplot(healthy.pie, aes(x = "", y = perc, fill = group)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels), color = rep(c("white")),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  guides(fill = guide_legend(title = "")) + 
  scale_fill_manual(values = stromal.colors) +
  coord_polar(theta = "y") +
  ggtitle("Healthy") +
  theme_void() + NoLegend() + theme(plot.title = element_text(hjust = 0.5))

myeloma.pie.plot <- ggplot(myeloma.pie, aes(x = "", y = perc, fill = group)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels), color = rep(c("white")),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  guides(fill = guide_legend(title = "")) + 
  scale_fill_manual(values = stromal.colors) +
  coord_polar(theta = "y") +
  ggtitle("Myeloma") + theme_void() + theme(plot.title = element_text(hjust = 0.5))
  
patchwork.pie <- healthy.pie.plot | myeloma.pie.plot
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/SupplementaryFigure11_BMEC_SubclusterPieCharts.png", width = 10, height = 6)
```

# Temp testing pie chart for Fig 3 overall stromal populations
```{r}
stromal.colors <- c("#F8766D", "#C49A00", "#53B400", "#00C094", "#00B6EB", "#A58AFF", "#FB61D7")

right.levels <- levels(stromals@meta.data$celltypes_ECsubbed)
healthy.values <- as.numeric(table(healthy@meta.data$celltypes_ECsubbed))
myeloma.values <- as.numeric(table(myeloma@meta.data$celltypes_ECsubbed))

healthy.frac <- healthy.values/sum(healthy.values)
myeloma.frac <- myeloma.values/sum(myeloma.values)

healthy.labels <- paste(as.character(round(healthy.frac, digits=4)*100), "%", sep="")
myeloma.labels <- paste(as.character(round(myeloma.frac, digits=4)*100), "%", sep="")

healthy.pie <- data.frame(group=c("SECs", "AECs", "Fibroblasts", "MSCs", "OLCs", "Pericytes", "Chondrocytes"), 
                          value = healthy.values, 
                          perc = healthy.frac, 
                          labels = healthy.labels)

myeloma.pie <- data.frame(group=c("SECs", "AECs", "Fibroblasts", "MSCs", "OLCs", "Pericytes", "Chondrocytes"), 
                          value = myeloma.values, 
                          perc = myeloma.frac, 
                          labels = myeloma.labels)

healthy.pie <- healthy.pie %>%
  arrange(desc(group)) %>%
  mutate(prop = value / sum(healthy.pie$value)*100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)

myeloma.pie <- myeloma.pie %>%
  arrange(desc(group)) %>%
  mutate(prop = value / sum(myeloma.pie$value)*100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop)

healthy.pie.plot <- ggplot(healthy.pie, aes(x = "", y = perc, fill = group)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels), color = rep(c("white")),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  guides(fill = guide_legend(title = "")) + 
  scale_fill_manual(values = stromal.colors) +
  coord_polar(theta = "y") +
  ggtitle("Healthy") +
  theme_void() + NoLegend() + theme(plot.title = element_text(hjust = 0.5))

myeloma.pie.plot <- ggplot(myeloma.pie, aes(x = "", y = perc, fill = group)) +
  geom_col(color = "black") +
  geom_label(aes(label = labels), color = rep(c("white")),
             position = position_stack(vjust = 0.5),
             show.legend = FALSE) +
  guides(fill = guide_legend(title = "")) + 
  scale_fill_manual(values = stromal.colors) +
  coord_polar(theta = "y") +
  ggtitle("Myeloma") + theme_void() + theme(plot.title = element_text(hjust = 0.5))
  
patchwork.pie <- healthy.pie.plot | myeloma.pie.plot
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Figure3_PieCharts.png", width = 10, height = 6)
```


