---
title: "PrelimWM_Signac_ATAC_Analysis"
output: html_document
date: "2023-04-20"
---

# Load libraries
```{r}
library(Signac)
library(Seurat)
library(GenomeInfoDb)
library(EnsDb.Hsapiens.v86)
library(ggplot2)
library(patchwork)
set.seed(1234)
options(future.globals.maxSize=10000000000000)
```
# Load data
```{r}
# Counts
p1.data <- Read10X("/Users/gagled01/morganLab/Waldenstroms/singlecell/data/01-190/filtered_feature_bc_matrix")
p2.data <- Read10X("/Users/gagled01/morganLab/Waldenstroms/singlecell/data/04-003/filtered_feature_bc_matrix")

# Get ATAC counts specifically
p1.counts <- p1.data$Peaks
p2.counts <- p2.data$Peaks

# Metadata
p1.metadata <- read.csv("/Users/gagled01/morganLab/Waldenstroms/singlecell/data/01-190/per_barcode_metrics.csv", header = TRUE, row.names = 1)
p2.metadata <- read.csv("/Users/gagled01/morganLab/Waldenstroms/singlecell/data/04-003/per_barcode_metrics.csv", header = TRUE, row.names = 1)

# Subsetting and cleaning metadata
p1.metadata <- p1.metadata[p1.metadata$is_cell == "1",] # only want non-failed cells
p2.metadata <- p2.metadata[p2.metadata$is_cell == "1",]

#rownames(p1.metadata) <- p1.metadata$barcode # set cell names to ATAC barcodes...not RNA
#rownames(p2.metadata) <- p2.metadata$barcode

p1.metadata <- p1.metadata[21:30] # only want ATAC metadata...not RNA
p2.metadata <- p2.metadata[21:30]

# Create chromatin assays w fragment files
p1.chrom_assay <- CreateChromatinAssay(
  counts = p1.counts,
  sep = c(":", "-"),
  genome = 'hg38',
  fragments = '/Users/gagled01/morganLab/Waldenstroms/singlecell/data/01-190/atac_fragments.tsv.gz',
  min.cells = 10,
  min.features = 200
)

p2.chrom_assay <- CreateChromatinAssay(
  counts = p2.counts,
  sep = c(":", "-"),
  genome = 'hg38',
  fragments = '/Users/gagled01/morganLab/Waldenstroms/singlecell/data/04-003/atac_fragments.tsv.gz',
  min.cells = 10,
  min.features = 200
)

# Final metadata clean so that it matches size of chromatin assay
p1.metadata <- p1.metadata[rownames(p1.metadata) %in% colnames(p1.chrom_assay),]
p2.metadata <- p2.metadata[rownames(p2.metadata) %in% colnames(p2.chrom_assay),]

# Create Seurat objects
p01_190 <- CreateSeuratObject(
  counts = p1.chrom_assay,
  assay = "peaks",
  meta.data = p1.metadata
)

p04_003 <- CreateSeuratObject(
  counts = p2.chrom_assay,
  assay = "peaks",
  meta.data = p2.metadata
)
```
# Adding annotations and metadata
```{r}
annotations <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
seqlevelsStyle(annotations) <- 'UCSC'

Annotation(p01_190) <- annotations
Annotation(p04_003) <- annotations

# Get ratio of fragments > 147bp and < 147bp (avg histone wrapping length is 147 so we're getting ratio of fragments w nucleosomes vs not)
p01_190 <- NucleosomeSignal(object = p01_190)
p04_003 <- NucleosomeSignal(object = p04_003)

# Get transcription start site enrichment
# Note, this will take a while unless you set fast = TRUE, but that will not allow you to plot cells using TSSPlot()
p01_190 <- TSSEnrichment(object = p01_190, fast = TRUE)
p04_003 <- TSSEnrichment(object = p04_003, fast = TRUE)

# Counting fragments in genome blacklist regions
p01_190$blacklist_fraction <- FractionCountsInRegion(
  object = p01_190, 
  assay = 'peaks',
  regions = blacklist_hg38
)

p04_003$blacklist_fraction <- FractionCountsInRegion(
  object = p04_003, 
  assay = 'peaks',
  regions = blacklist_hg38
)

# Counting fraction of reads in peaks
p01_190$pct_reads_in_peaks <- p01_190$atac_peak_region_fragments / p01_190$atac_fragments * 100
p04_003$pct_reads_in_peaks <- p04_003$atac_peak_region_fragments / p04_003$atac_fragments * 100
```
# Computing QC Metrics

# Nucleosome banding pattern - the histogram of DNA fragment sizes should exhibit a strong pattern corresopnding to the length of DNA wrapped around a single nucleosome (that's 147bp for us)
# We calculate the ratio of fragments with 1 nucleosome to those with 0

# Plotting TSS score
# TSS score is essentially the ratio of fragments centered at the TSS to fragments in TSS-flanked regions. Poor ATAC data has low TSS score

# Total fragments in peaks - a measure of sequencing depth. Exclude those with too few or too many

# Fraction of fragments in peaks - fraction of fragments that fall within peaks. Cells with low values (< 15-20%) are low quality

# Ratio of reads in blacklist regions aka areas known to be assocaited with artifact signals
```{r}
VlnPlot(p01_190, features = c("atac_peak_region_fragments", "pct_reads_in_peaks", "blacklist_fraction", "nucleosome_signal","TSS.enrichment"), ncol = 5)
ggsave("/Users/gagled01/morganLab/Waldenstroms/singlecell/figures/WMPrelim_p01_190_ATAC_QC_BeforeFilter.png", height = 4, width = 15)

VlnPlot(p04_003, features = c("atac_peak_region_fragments", "pct_reads_in_peaks", "blacklist_fraction", "nucleosome_signal","TSS.enrichment"), ncol = 5)
ggsave("/Users/gagled01/morganLab/Waldenstroms/singlecell/figures/WMPrelim_p04_003_ATAC_QC_BeforeFilter.png", height = 4, width = 15)
```
# Filtering
```{r}
low_prf <- quantile(p01_190[["atac_peak_region_fragments"]]$atac_peak_region_fragments, probs = 0.02)
high_prf <- quantile(p01_190[["atac_peak_region_fragments"]]$atac_peak_region_fragments, probs = 0.98)

low_prp <- quantile(p01_190[["pct_reads_in_peaks"]]$pct_reads_in_peaks, probs = 0.02)
high_prp <- quantile(p01_190[["pct_reads_in_peaks"]]$pct_reads_in_peaks, probs = 0.98)

high_blr <- quantile(p01_190[["blacklist_fraction"]]$blacklist_fraction, probs = 0.98)
high_ns <- quantile(p01_190[["nucleosome_signal"]]$nucleosome_signal, probs = 0.98)
high_tss <- quantile(p01_190[["TSS.enrichment"]]$TSS.enrichment, probs = 0.98)
low_tss <- quantile(p01_190[["TSS.enrichment"]]$TSS.enrichment, probs = 0.02)

p01_190_filtered <- subset(p01_190, subset = atac_peak_region_fragments > 3000 &
                       atac_peak_region_fragments < 20000 &
                       pct_reads_in_peaks > 20 &
                       blacklist_fraction < 0.05 &
                       nucleosome_signal < 4 &
                       TSS.enrichment < 10 &
                       TSS.enrichment > 2)
VlnPlot(p01_190_filtered, features = c("atac_peak_region_fragments", "pct_reads_in_peaks", "blacklist_fraction", "nucleosome_signal","TSS.enrichment"), ncol = 5)
ggsave("/Users/gagled01/morganLab/Waldenstroms/singlecell/figures/WMPrelim_p01_190_ATAC_QC_ManuallyFiltered.png", height = 4, width = 15)

p04_003_filtered <- subset(p04_003, subset = atac_peak_region_fragments > 3000 &
                       atac_peak_region_fragments < 20000 &
                       pct_reads_in_peaks > 20 &
                       blacklist_fraction < 0.05 &
                       nucleosome_signal < 4 &
                       TSS.enrichment < 10 &
                       TSS.enrichment > 2)
VlnPlot(p04_003_filtered, features = c("atac_peak_region_fragments", "pct_reads_in_peaks", "blacklist_fraction", "nucleosome_signal","TSS.enrichment"), ncol = 5)
ggsave("/Users/gagled01/morganLab/Waldenstroms/singlecell/figures/WMPrelim_p04_003_ATAC_QC_ManuallyFiltered.png", height = 4, width = 15)
```
# Merge patients
```{r}
p01_190_filtered$patient <- "p01_190"
p04_003_filtered$patient <- "p04_003"
merged <- merge(p01_190_filtered, p04_003_filtered)
```
# Normalization and dimensional reduction
Normalization is using frequency-inverse document frequency (TF-IDF) normalization which normalizes across cells to correct for sequencing depth and also across peaks to give more weight to rare peaks

# Feature selection
It's hard to find features in ATAC data so we set the cutoff to include all features

# Dimensional reduction
Use singular value decomposition (SVD), this is basically the ATAC version of PCA
```{r}
merged <- RunTFIDF(merged)
merged <- FindTopFeatures(merged, min.cutoff = 'q0')
merged <- RunSVD(merged)
```


# UMAP, Neighbors, Clustering with standard Seurat functions
```{r}
merged <- RunUMAP(merged, reduction = 'lsi', dims = 2:30)
merged <- FindNeighbors(merged, reduction = 'lsi', dims = 2:30)
merged <- FindClusters(merged, verbose = FALSE, algorithm = 3)
DimPlot(merged, label = TRUE) + NoLegend()
ggsave("/Users/gagled01/morganLab/Waldenstroms/singlecell/figures/WMPrelim_p04_003_ATAC_MergedUMAP_byCluster.png", height = 4, width = 4)
DimPlot(merged, label = TRUE, group.by = "patient") + NoLegend()
ggsave("/Users/gagled01/morganLab/Waldenstroms/singlecell/figures/WMPrelim_p04_003_ATAC_MergedUMAP_byPatient.png", height = 4, width = 4)

```
# Estimate gene activity
```{r}
gene.activities <- GeneActivity(merged)

# add the gene activity matrix to the Seurat object as a new assay and normalize it
merged[['RNA']] <- CreateAssayObject(counts = gene.activities)
merged <- NormalizeData(
  object = merged,
  assay = 'RNA',
  normalization.method = 'LogNormalize',
  scale.factor = median(merged$nCount_RNA)
)
```

```{r}
DefaultAssay(merged) <- 'RNA'

FeaturePlot(
  object = merged,
  features = c('MS4A1', 'MZB1', 'IGHM'),
  pt.size = 0.1,
  max.cutoff = 'q95',
  ncol = 3
)
```


