---
title: "PseudoSampling_DA"
output: html_document
date: "2023-03-03"
---
################################################################################

# Script which attempts to develop a method to pseudosample the Mouse 5TG data due to failed hashtags.
# The general workflow is to take isolate the control and myeloma libraries (~6k and 8k samples each) and then randomly subsample them, without replacement, to 50% of their original size
# For each subsample, store the relative abundance (%) of each cell type
# Repeat this 1000x
# This data will then be used to generate bar and box plots of relative abundance for each celltype between both cases and controls, equipped with confidence intervals and possibly even p-values.
################################################################################

# Load libraries
```{r}
rm(list = ls()) # clear environment
options(warn=-1) # turn off warning messages globally

library(Seurat)
library(dplyr)
library(reshape2)
library(gridExtra)
library(ggpubr)

options(future.globals.maxSize=1000000000000000) # Set max global size so we don't run out of memory
```
# Load data
```{r}
stromals <- readRDS("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Rerun_Mouse5TG_StromalsOnly_AnnotatedObject6_21PCs_NoLib_FIXEDLIBRARIES_6.rds")
```
# Subset by condition
```{r}
healthy <- subset(stromals, subset = case.control == "Healthy")
myeloma <- subset(stromals, subset = case.control == "Myeloma")
```
# RA
```{r}
table(healthy@meta.data$celltypes_ECsubbed)/sum(table(healthy@meta.data$celltypes_ECsubbed)) * 100
table(myeloma@meta.data$celltypes_ECsubbed)/sum(table(myeloma@meta.data$celltypes_ECsubbed)) * 100
```
# Create storage df
```{r}
storage.df <- data.frame(matrix(ncol = 9, nrow = 0))
colnames(storage.df) <- c("cell.ID", "case.control", "MSCs", "OLCs", "Fibroblasts", "Chondrocytes", "Arterial ECs", "Sinusoidal ECs", "Pericytes")
```
# Subsampling for loop
```{r}
# Loop for subsampling where
# seurat_obj = the object you are subsampling from
# case.control = the string of the condition you will populate your output dataframe with
# iterations = the number of subsamples to generate

subsample_loop <- function(seurat_obj, case.control, iterations) {
  
  # Dataframe to store values
  storage.df <- data.frame(matrix(ncol = 8, nrow = iterations))
  colnames(storage.df) <- c("case.control", "MSCs", "OLCs", "Fibroblasts", "Chondrocytes", "aECs", "sECs", "Pericytes")
  
  # Random downsampling loop
  for (k in 1:iterations) {
    
    # Do the subsampling without replacement to 50% of original size
    subsample <- slice_sample(seurat_obj@meta.data, prop = 0.05) %>% select(celltypes_ECsubbed)
    sum_cells <- ncol(healthy) # get total cell count
    
    # Storage for cell counts
    msc.count <- 0
    olc.count <- 0
    fibroblast.count <- 0
    chondrocyte.count <- 0
    arterialEC.count <- 0
    sinusoidalEC.count <- 0
    pericyte.count <- 0
  
    # Looping through subsample to get cell counts
    for (i in 1:nrow(subsample)) {
      cell.type <- as.vector(subsample$celltypes_ECsubbed[i]) # Get cell type
      
      if (cell.type == "MSCs") {
        msc.count <- msc.count + 1
      } else if (cell.type == "OLCs") {
        olc.count <- olc.count + 1
      } else if (cell.type == "Fibroblasts") {
        fibroblast.count <- fibroblast.count + 1
      } else if (cell.type == "Chondrocytes") {
        chondrocyte.count <- chondrocyte.count + 1
      } else if (cell.type == "Arterial ECs") {
        arterialEC.count <- arterialEC.count + 1
      } else if (cell.type == "Sinusoidal ECs") {
        sinusoidalEC.count <- sinusoidalEC.count + 1
      } else if (cell.type == "Pericytes") {
        pericyte.count <- pericyte.count + 1
      }
    } 
    
    # Add results to storage df
    storage.df$case.control[k] <- case.control
    storage.df$MSCs[k] <- msc.count/nrow(subsample)
    storage.df$OLCs[k] <- olc.count/nrow(subsample)
    storage.df$Fibroblasts[k] <- fibroblast.count/nrow(subsample)
    storage.df$Chondrocytes[k] <- chondrocyte.count/nrow(subsample)
    storage.df$aECs[k] <- arterialEC.count/nrow(subsample)
    storage.df$sECs[k] <- sinusoidalEC.count/nrow(subsample)
    storage.df$Pericytes[k] <- pericyte.count/nrow(subsample)
  }
  return(storage.df)
}
```
# Run the function
```{r}
set.seed(99)

healthy.pseudosamples <- subsample_loop(healthy, "Healthy", 5)
head(healthy.pseudosamples)

myeloma.pseudosamples <- subsample_loop(myeloma, "Myeloma", 5)
head(myeloma.pseudosamples)

mean(healthy.pseudosamples$MSCs)
mean(healthy.pseudosamples$OLCs)
mean(healthy.pseudosamples$Fibroblasts)
mean(healthy.pseudosamples$Chondrocytes)
mean(healthy.pseudosamples$aECs)
mean(healthy.pseudosamples$sECs)
mean(healthy.pseudosamples$Pericytes)

mean(myeloma.pseudosamples$MSCs)
mean(myeloma.pseudosamples$OLCs)
mean(myeloma.pseudosamples$Fibroblasts)
mean(myeloma.pseudosamples$Chondrocytes)
mean(myeloma.pseudosamples$aECs)
mean(myeloma.pseudosamples$sECs)
mean(myeloma.pseudosamples$Pericytes)

```
# Generate pie charts
```{r}

```

# Combine dataframes
```{r}
combined.pseudosamples <- rbind(healthy.pseudosamples, myeloma.pseudosamples)
```
# Set up dataframe for plotting
```{r}
head(combined.pseudosamples)

pseudosamples.melt <- melt(combined.pseudosamples, id = c("case.control"))
colnames(pseudosamples.melt)[2] <- "celltype"
head(pseudosamples.melt)
```
#
```{r}
healthy.subtype <- subset(subtype, subset = case.control == "Healthy")
myeloma.subtype <- subset(subtype , subset = case.control == "Myeloma")

healthy.frac <- as.numeric(table(healthy.subtype@meta.data$seurat_clusters))/sum(table(healthy.subtype@meta.data$seurat_clusters))
healthy.frac <- healthy.frac[1:6]

myeloma.frac <- as.numeric(table(myeloma.subtype@meta.data$seurat_clusters))/sum(table(myeloma.subtype@meta.data$seurat_clusters))
myeloma.frac <- myeloma.frac[1:6]
```

# Start plotting
```{r}
# Barplot
pseudo.bar <- ggplot(pseudosamples.melt, aes(x = celltype, y = value, fill = case.control)) + geom_bar(position = "dodge", stat = "identity") + theme_minimal() + ggtitle("Relative Abundance of Cell Types between Conditions in Random Pseudosamples") + ylab("Relative Abundance") + xlab("Cell Type") + theme_minimal()
ggsave(plot = pseudo.bar, filename = "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/CellTypeRelativeAbundance_Barplot_Condition_Pseudosampled_05PercDownSample.png", width = 12, height = 3)

library(patchwork)
# Boxplot
msc.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "MSCs",], aes(x = celltype, y = value, fill = case.control)) + geom_boxplot() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank()) + theme_minimal()
olc.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "OLCs",], aes(x = celltype, y = value, fill = case.control)) + geom_boxplot() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank()) + theme_minimal()
fibro.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "Fibroblasts",], aes(x = celltype, y = value, fill = case.control)) + geom_boxplot() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank()) + theme_minimal()
chondro.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "Chondrocytes",], aes(x = celltype, y = value, fill = case.control)) + geom_boxplot() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank()) + theme_minimal()
aEC.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "aECs",], aes(x = celltype, y = value, fill = case.control)) + geom_boxplot() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank()) + theme_minimal()
sEC.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "sECs",], aes(x = celltype, y = value, fill = case.control)) + geom_boxplot() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank()) + theme_minimal()
peri.box <- ggplot(pseudosamples.melt[pseudosamples.melt$celltype == "Pericytes",], aes(x = celltype, y = value, fill = case.control)) + geom_boxplot() +
  theme(axis.title.x=element_blank(),
        axis.title.y=element_blank()) + theme_minimal()

# Patchwork trying
pseudo.boxplots <- msc.box + olc.box + fibro.box + chondro.box + aEC.box + sEC.box + peri.box + plot_layout(ncol = 7, guides = "collect")
gt <- patchwork::patchworkGrob(pseudo.boxplots)
gt.fixed <- gridExtra::grid.arrange(gt, left = "Relative Abundance", bottom = "Cell Type")
ggsave(plot = gt.fixed, filename = "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/CellTypeRelativeAbundance_Boxplot_Condition_Pseudosampled_05PercDownSample.png", width = 12, height = 3)
```
# Stats testing
```{r}
msc.adj.p <- p.adjust(t.test(healthy.pseudosamples$MSCs, myeloma.pseudosamples$MSCs, alternative = "two.sided")$p.value, method = "bonferroni", n = 5)
olc.adj.p <- p.adjust(t.test(healthy.pseudosamples$OLCs, myeloma.pseudosamples$OLCs, alternative = "two.sided")$p.value, method = "bonferroni", n = 5)
fibro.adj.p <- p.adjust(t.test(healthy.pseudosamples$Fibroblasts, myeloma.pseudosamples$Fibroblasts, alternative = "two.sided")$p.value, method = "bonferroni", n = 5)
chondro.adj.p <- p.adjust(t.test(healthy.pseudosamples$Chondrocytes, myeloma.pseudosamples$Chondrocytes, alternative = "two.sided")$p.value, method = "bonferroni", n = 5)
aEC.adj.p <- p.adjust(t.test(healthy.pseudosamples$aECs, myeloma.pseudosamples$aECs, alternative = "two.sided")$p.value, method = "bonferroni", n = 5)
sEC.adj.p <- p.adjust(t.test(healthy.pseudosamples$sECs, myeloma.pseudosamples$sECs, alternative = "two.sided")$p.value, method = "bonferroni", n = 5)
peri.adj.p <- p.adjust(t.test(healthy.pseudosamples$Pericytes, myeloma.pseudosamples$Pericytes, alternative = "two.sided")$p.value, method = "bonferroni", n = 5)

p.values <- c(msc.adj.p, olc.adj.p, fibro.adj.p, chondro.adj.p, aEC.adj.p, sEC.adj.p, peri.adj.p)
p.values
```

# Incorporating p values to boxplots
```{r}
pvalue_forPlot <- tibble(
  "values" = c(healthy.pseudosamples$MSCs,
                myeloma.pseudosamples$MSCs),
  "case.control" = c(rep("Healthy", length(healthy.pseudosamples$MSCs)),
              rep("Myeloma", length(healthy.pseudosamples$MSCs)))) %>%
  rstatix::wilcox_test(formula = values ~ case.control,
                       alternative = "two.sided",
                       p.adjust.method = "bonferroni",
                       mu = 0,
                       paired = FALSE) %>%
  rstatix::add_significance() %>%
  rstatix::add_xy_position(x = "case.control")

boxplot_plot <- ggplot(tibble(
  "values" = c(healthy.pseudosamples$MSCs,
                myeloma.pseudosamples$MSCs),
  "case.control" = c(rep("Healthy", length(healthy.pseudosamples$MSCs)),
              rep("Myeloma", length(healthy.pseudosamples$MSCs))))) +
  # geom_jitter(aes(x = case.control, y = values),
  #             color = "black",
  #             size = 1.15,
  #             alpha = 0.65,
  #             width = 0.15) +
  geom_boxplot(aes(x = case.control, y = values),
               width = 0.60,
               fill = c("#F8776D", "#00BFC4"),
               color = "black",
               alpha = 0.75,
               outlier.shape = NA) +
  stat_boxplot(aes(x = case.control, y = values),
               geom = "errorbar",
               width = 0.40) +
  stat_pvalue_manual(data = pvalue_forPlot,
                     label = "{p} {p.signif}",
                     bracket.nudge.y = 0.01) +
  #scale_y_continuous(breaks = c(-6,-3,0,3,6)) +
  ylab("Relative Abundance") +
  xlab("") + ggtitle("5% Downsampled MSCs") + 
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5))

boxplot_plot
```

# Dev zone
```{r}
boxplot_loop <- function(healthy_vector, myeloma_vector,) {
  plotting_df <- tibble(
    values = c(healthy_vector,
               myeloma_vector),
    condition = c(rep("Healthy"), length(healthy_vector)),
    rep("Myeloma", length(myeloma_vector))) %>%
    rstatix::t_test(formula =
  )
  
}
pvalue_forPlot <- tibble(
  "values" = c(healthy.pseudosamples$MSCs,
                myeloma.pseudosamples$MSCs),
  "case.control" = c(rep("Healthy", length(healthy.pseudosamples$MSCs)),
              rep("Myeloma", length(healthy.pseudosamples$MSCs)))) %>%
  rstatix::wilcox_test(formula = values ~ case.control,
                       alternative = "two.sided",
                       p.adjust.method = "bonferroni",
                       mu = 0,
                       paired = FALSE) %>%
  rstatix::add_significance() %>%
  rstatix::add_xy_position(x = "case.control")

boxplot_plot <- ggplot(tibble(
  "values" = c(healthy.pseudosamples$MSCs,
                myeloma.pseudosamples$MSCs),
  "case.control" = c(rep("Healthy", length(healthy.pseudosamples$MSCs)),
              rep("Myeloma", length(healthy.pseudosamples$MSCs))))) +
  geom_boxplot(aes(x = case.control, y = values),
               width = 0.60,
               fill = c("#F8776D", "#00BFC4"),
               color = "black",
               alpha = 0.75,
               outlier.shape = NA) +
  stat_boxplot(aes(x = case.control, y = values),
               geom = "errorbar",
               width = 0.40) +
  stat_pvalue_manual(data = pvalue_forPlot,
                     label = "{p} {p.signif}",
                     bracket.nudge.y = 0.01) +
  #scale_y_continuous(breaks = c(-6,-3,0,3,6)) +
  ylab("Relative Abundance") +
  xlab("") + ggtitle("5% Downsampled MSCs") + 
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
```

