---
title: "Rerun_5TG_DiffusionMapping"
output: html_document
date: "2022-12-01"
---

```{r}
library(Seurat)
library(SingleCellExperiment)
library(destiny)
library(scater)
library(gam)
library(corrplot)
library(ggplot2)
library(ggthemes)
library(ggbeeswarm)
library(dplyr)
library(cowplot)
library(RColorBrewer)
options(future.globals.maxSize=1000000000000000) # Set max global size so we don't run out of memory
```
# load in data
```{r}
stromals <- readRDS("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Rerun_Mouse5TG_StromalsOnly_AnnotatedObject6_21PCs_NoLib_FIXEDLIBRARIES_6.rds")
```
# Library cell counts
```{r}
lib1 <- subset(stromals, subset = library == "lib1")
lib2 <- subset(stromals, subset = library == "lib2")
lib3 <- subset(stromals, subset = library == "lib3")
lib5 <- subset(stromals, subset = library == "lib5")
lib6 <- subset(stromals, subset = library == "lib6")

table(lib1@meta.data$celltypes_ECsubbed)
table(lib2@meta.data$celltypes_ECsubbed)
table(lib3@meta.data$celltypes_ECsubbed)
table(lib5@meta.data$celltypes_ECsubbed)
table(lib6@meta.data$celltypes_ECsubbed)


table(lib1@meta.data$celltypes_ECsubbed)/sum(table(lib1@meta.data$celltypes_ECsubbed))
table(lib2@meta.data$celltypes_ECsubbed)/sum(table(lib2@meta.data$celltypes_ECsubbed))
table(lib3@meta.data$celltypes_ECsubbed)/sum(table(lib3@meta.data$celltypes_ECsubbed))
table(lib5@meta.data$celltypes_ECsubbed)/sum(table(lib5@meta.data$celltypes_ECsubbed))
table(lib6@meta.data$celltypes_ECsubbed)/sum(table(lib6@meta.data$celltypes_ECsubbed))
```

# Update metadata
```{r}
outlier.down.list <- read.csv("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSC_Myeloma_OutlierDown_CellNames.csv")

stromals@meta.data$celltypes_ECsubbed2 <- stromals@meta.data$celltypes_ECsubbed
stromals@meta.data <- stromals@meta.data %>% mutate(celltypes_ECsubbed2 = ifelse(rownames(stromals@meta.data) %in% outlier.down.list$x, "Osteoclast-like", celltypes_ECsubbed2))
table(stromals@meta.data$celltypes_ECsubbed2)

stromals@meta.data <- stromals@meta.data %>% 
  mutate(celltype_cond = case_when(
    case.control == "Myeloma" & celltypes_ECsubbed == "Fibroblasts" ~ "MM Fibroblasts",
    case.control == "Myeloma" & celltypes_ECsubbed == "Chondrocytes" ~ "MM Chondrocytes",
    case.control == "Myeloma" & celltypes_ECsubbed == "MSCs" ~ "MM MSCs",
    case.control == "Myeloma" & celltypes_ECsubbed == "OLCs" ~ "MM OLCs",
    case.control == "Myeloma" & celltypes_ECsubbed == "Sinusoidal ECs" ~ "MM sECs",
    case.control == "Myeloma" & celltypes_ECsubbed == "Arterial ECs" ~ "MM aECs",
    case.control == "Myeloma" & celltypes_ECsubbed == "Pericytes" ~ "MM Pericytes",
    case.control == "Healthy" & celltypes_ECsubbed == "Fibroblasts" ~ "Healthy Fibroblasts",
    case.control == "Healthy" & celltypes_ECsubbed == "Chondrocytes" ~ "Healthy Chondrocytes",
    case.control == "Healthy" & celltypes_ECsubbed == "MSCs" ~ "Healthy MSCs",
    case.control == "Healthy" & celltypes_ECsubbed == "OLCs" ~ "Healthy OLCs",
    case.control == "Healthy" & celltypes_ECsubbed == "Sinusoidal ECs" ~ "Healthy sECs",
    case.control == "Healthy" & celltypes_ECsubbed == "Arterial ECs" ~ "Healthy aECs",
    case.control == "Healthy" & celltypes_ECsubbed == "Pericytes" ~ "Healthy Pericytes"
  ))

table(stromals@meta.data$celltype_cond)
sum(table(stromals@meta.data$celltype_cond))

# Backup which includes outlier osteolike fibro info
stromals@meta.data$celltype_cond2 <- stromals@meta.data$celltype_cond
stromals@meta.data <- stromals@meta.data %>% mutate(celltype_cond2 = ifelse(rownames(stromals@meta.data) %in% outlier.down.list$x, "Osteoclast-like Fibroblasts", celltype_cond2))
table(stromals@meta.data$celltype_cond2)

#saveRDS(stromals, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Rerun_Mouse5TG_StromalsOnly_AnnotatedObject6_21PCs_NoLib6.rds")
```
# Exploring outlier cells~!
```{r}
# Discovering that the weird myeloma MSC cells were from lib 6!
weird.myelomas <- subset(myelomas, cells = outlier.cells)
DimPlot(weird.myelomas, group.by = "celltypes_ECsubbed")
DimPlot(weird.myelomas, group.by = "library")
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Mouse_5TG_Stromals_WeirdCells_LibraryUMAP.png")
```


# Separating MSCs and ECs as they have different cellular origins
```{r}
# MSCs
msc.lineage <- subset(stromals, subset = celltype_cond2 %in% c("Healthy Fibroblasts", "Healthy MSCs", "Healthy OLCs", "Healthy Chondrocytes",
                                                               "MM Fibroblasts", "MM MSCs", "MM OLCs", "MM Chondrocytes",
                                                               "Osteoclast-like Fibroblasts"))
table(msc.lineage@meta.data$celltype_cond2)

#ECs
ec.lineage <- subset(stromals, subset = celltype_cond2 %in% c("Healthy aECs", "Healthy sECs", "MM aECs", "MM sECs"))
table(ec.lineage@meta.data$celltype_cond2)
```
# Breaking down for disease/control
```{r}
msc.healthy <- subset(msc.lineage, subset = case.control == "Healthy")
msc.myeloma <- subset(msc.lineage, subset = case.control == "Myeloma")

ec.healthy <- subset(ec.lineage, subset = case.control == "Healthy")
ec.myeloma <- subset(ec.lineage, subset = case.control == "Myeloma")

DimPlot(msc.healthy, group.by = "celltypes_ECsubbed")
DimPlot(msc.myeloma, group.by = "celltypes_ECsubbed")


# Convert to SCE
msc.healthy.sce <- as.SingleCellExperiment(msc.healthy)
msc.myeloma.sce <- as.SingleCellExperiment(msc.myeloma)

ec.healthy.sce <- as.SingleCellExperiment(ec.healthy)
ec.myeloma.sce <- as.SingleCellExperiment(ec.myeloma)
```
# Load in diffusion maps
```{r}
ec.diffusions <- readRDS("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/EC_Myeloma_DiffusionMap.rds")
```

# PCs? Why? Skip for the time being
```{r}
# PCA 
# msc.healthy.sce <- runPCA(msc.healthy.sce, ncomponents = 10)
# msc.healthy.pca <- reducedDim(msc.healthy.sce, "PCA")
# msc.healthy.sce$PC1 <- msc.healthy.pca[,1]
# msc.healthy.sce$PC2 <- msc.healthy.pca[,2]
# 
# msc.myeloma.sce <- runPCA(msc.myeloma.sce, ncomponents = 10)
# msc.myeloma.pca <- reducedDim(msc.myeloma.sce, "PCA")
# msc.myeloma.sce$PC1 <- msc.myeloma.pca[,1]
# msc.myeloma.sce$PC2 <- msc.myeloma.pca[,2]
# 
# ec.healthy.sce <- runPCA(ec.healthy.sce, ncomponents = 10)
# ec.healthy.pca <- reducedDim(ec.healthy.sce, "PCA")
# ec.healthy.sce$PC1 <- ec.healthy.pca[,1]
# ec.healthy.sce$PC2 <- ec.healthy.pca[,2]
# 
# ec.myeloma.sce <- runPCA(ec.myeloma.sce, ncomponents = 10)
# ec.myeloma.pca <- reducedDim(ec.myeloma.sce, "PCA")
# ec.myeloma.sce$PC1 <- ec.myeloma.pca[,1]
# ec.myeloma.sce$PC2 <- ec.myeloma.pca[,2]
```
# Run diffusion maps
# MSCs healthy
```{r}
msc.healthy.matrix <- as.matrix(logcounts(msc.healthy.sce))  # access log-transformed counts matrix
msc.healthy.cellLabels <- msc.healthy.sce$celltypes_ECsubbed
colnames(msc.healthy.matrix) <- msc.healthy.cellLabels
msc.healthy.dm <- DiffusionMap(data = t(msc.healthy.matrix), n_pcs = 50)

saveRDS(msc.healthy.dm, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/HealthyMSCs_DiffusionMap.rds")

```
# MSCs myeloma
```{r}
msc.myeloma.matrix <- as.matrix(logcounts(msc.myeloma.sce))  # access log-transformed counts matrix
msc.myeloma.cellLabels <- msc.myeloma.sce$celltypes_ECsubbed
msc.myeloma.cellNames <- colnames(msc.myeloma.sce)
colnames(msc.myeloma.matrix) <- msc.myeloma.cellLabels
msc.myeloma.dm <- DiffusionMap(data = t(msc.myeloma.matrix), n_pcs = 50)

saveRDS(msc.myeloma.dm, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MyelomaMSCs_DiffusionMap.rds")
```
# Passing on the ECs for now as they have already been done
# ECs healthy
```{r}
ec.healthy.matrix <- as.matrix(logcounts(ec.healthy.sce))  # access log-transformed counts matrix
ec.healthy.cellLabels <- ec.healthy.sce$celltypes_ECsubbed
colnames(ec.healthy.matrix) <- ec.healthy.cellLabels
ec.healthy.dm <- DiffusionMap(data = t(ec.healthy.matrix), n_pcs = 50)

# save out
saveRDS(ec.healthy.dm, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/HealthyECs_DiffusionMap.rds")
```
# ECs myeloma
```{r}
ec.myeloma.matrix <- as.matrix(logcounts(ec.myeloma.sce))  # access log-transformed counts matrix
ec.myeloma.cellLabels <- ec.myeloma.sce$celltypes_ECsubbed
colnames(ec.myeloma.matrix) <- ec.myeloma.cellLabels
ec.myeloma.dm <- DiffusionMap(data = t(ec.myeloma.matrix), n_pcs = 50)

# save out
saveRDS(ec.myeloma.dm, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MyelomaECs_DiffusionMap.rds")
```
# Plot disease/controls
```{r}
tmp <- data.frame(DC1 = eigenvectors(msc.healthy.dm)[, 1],
                  DC2 = eigenvectors(msc.healthy.dm)[, 2],
                  Timepoint = msc.healthy.cellLabels)
ggplot(tmp, aes(x = DC1, y = DC2, colour = Timepoint)) +
    geom_point() + scale_color_manual(values = c("#F8766D", "#00BA38", "#00BFC4", "#B79F00")) +
    ggtitle("Control - MSC Lineage Cells") +
    xlab("Diffusion component 1") + 
    ylab("Diffusion component 2") +
    theme_classic()
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/MSC_Lineage_Controls_DiffusionMap.png", height = 6, width = 8)

#fibro.outliers <- tmp[tmp$DC2 > 0.04,]$Timepoint
#write.csv(fibro.outliers, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSC_Myeloma_OutlierFibro_CellNames.csv")

tmp <- data.frame(DC1 = eigenvectors(msc.myeloma.dm)[, 1],
                  DC2 = eigenvectors(msc.myeloma.dm)[, 2],
                  Timepoint = msc.myeloma.cellLabels)
ggplot(tmp, aes(x = DC1, y = DC2, colour = Timepoint)) +
    geom_point() + scale_color_manual(values = c("#F8766D", "#00BA38", "#00BFC4", "#B79F00")) +
    ggtitle("Myeloma - MSC Lineage Cells") +
    xlab("Diffusion component 1") + 
    ylab("Diffusion component 2") +
    theme_classic()
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/MSC_Lineage_Myeloma_DiffusionMap.png", height = 6, width = 8)

outlier.cells1 <- tmp[tmp$DC1 > 0.05,]$Timepoint # These are the cells with the large DC1 which drives the diffusion behavior in the myeloma MSCs
write.csv(outlier.cells, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Myeloma_MSCs_DiffusionMap_OutlierCells.csv")

```
# Fibro exploration
```{r}
outlier.cells1 <- tmp[tmp$DC2 > 0.04,]$Timepoint
write.csv(outlier.cells1, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSC_Myeloma_OutlierUp_CellNames.csv")
outlier.cells2 <- tmp[tmp$DC2 < -0.04,]$Timepoint
write.csv(outlier.cells2, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSC_Myeloma_OutlierDown_CellNames.csv")


```

# ECs
```{r}
### ECs
tmp <- data.frame(DC1 = eigenvectors(ec.healthy.dm)[, 1],
                  DC2 = eigenvectors(ec.healthy.dm)[, 2],
                  Timepoint = ec.healthy.cellLabels)
ggplot(tmp, aes(x = DC1, y = DC2, colour = Timepoint)) +
    geom_point() + scale_color_manual(values = c("#F564E3", "#C77CFF")) + 
    ggtitle("Control - Endothelial Cells") +
    xlab("Diffusion component 1") + 
    ylab("Diffusion component 2") +
    theme_classic()
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/EC_Lineage_Controls_DiffusionMap.png", height = 6, width = 8)

tmp <- data.frame(DC1 = eigenvectors(ec.myeloma.dm)[, 1],
                  DC2 = eigenvectors(ec.myeloma.dm)[, 2],
                  Timepoint = ec.myeloma.cellLabels)
ggplot(tmp, aes(x = DC1, y = DC2, colour = Timepoint)) +
    geom_point() + scale_color_manual(values = c("#F564E3", "#C77CFF")) +  
    ggtitle("Myeloma - Endothelial Cells") +
    xlab("Diffusion component 1") + 
    ylab("Diffusion component 2") +
    theme_classic()
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/EC_Lineage_Myeloma_DiffusionMap.png", height = 6, width = 8)
```
# Pseudotime
```{r}
msc.myeloma@meta.data$pseudotime_diffusionmap <- rank(eigenvectors(msc.myeloma.dm)[,1])

index <- 1:length(msc.myeloma@meta.data$pseudotime_diffusionmap)
dpt <- DPT(msc.myeloma.dm, tips = index[msc.myeloma@meta.data$pseudotime_diffusionmap == 1])
msc.myeloma@meta.data$dpt <- dpt$DPT1

msc.myeloma.dpt <- DPT(msc.myeloma.dm)

set.seed(4)
msc.healthy.dpt_random <- DPT(msc.myeloma.dm, tips = colnames())

plot(msc.myeloma.dpt, root = 2, paths_to = c(1,3), col_by = 'branch')
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/Myeloma_Fibros_PseudotimeKinda.png", height = 6, width = 8)

```


##################### NOT SPLIT UP BY CONDITION #######################

# Add osteoclast-like annotation
```{r}
stromals <- readRDS("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/Rerun_Mouse5TG_StromalsOnly_AnnotatedObject6_21PCs_NoLib_FIXEDLIBRARIES_6.rds")

# Osteoclast-like cells
outlier.down.list <- read.csv("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSC_Myeloma_OutlierDown_CellNames.csv")

stromals@meta.data <- stromals@meta.data %>% mutate(celltypes_ECsubbed = ifelse(rownames(stromals@meta.data) %in% outlier.down.list$x, "Osteoclast-like", celltypes_ECsubbed))
table(stromals@meta.data$celltypes_ECsubbed)
```

# Get right cells
```{r}
# MSCs
msc.lineage <- subset(stromals, subset = celltypes_ECsubbed %in% c("Fibroblasts", "Chondrocytes", "MSCs", "OLCs"))
table(msc.lineage@meta.data$celltypes_ECsubbed)

#ECs
ec.lineage <- subset(stromals, subset = celltypes_ECsubbed %in% c("Arterial ECs", "Sinusoidal ECs"))
table(ec.lineage@meta.data$celltypes_ECsubbed)
```
# Add osteoclast-like to mscs
```{r}
msc.lineage@meta.data <- msc.lineage@meta.data %>% mutate(celltypes_ECsubbed = ifelse(rownames(msc.lineage@meta.data) %in% outlier.down.list$x, "Osteoclast-like", celltypes_ECsubbed))
table(msc.lineage@meta.data$celltypes_ECsubbed)

Idents(msc.lineage) <- "celltypes_ECsubbed"
msc.lineage <- RenameIdents(msc.lineage, "2" = "Fibroblasts")
msc.lineage <- RenameIdents(msc.lineage, "3" = "MSCs")
msc.lineage <- RenameIdents(msc.lineage, "4" = "OLCs")
msc.lineage <- RenameIdents(msc.lineage, "7" = "Chondrocytes")

msc.lineage@meta.data$celltypes_ECsubbed <- Idents(msc.lineage)
table(msc.lineage@meta.data$celltypes_ECsubbed)

```

# Convert to SCE
```{r}
msc.sce <- as.SingleCellExperiment(msc.lineage)
ec.sce <- as.SingleCellExperiment(ec.lineage)
```
# First test run
# All cells (including ECs, which we want to do separately)
```{r}
msc.matrix <- as.matrix(logcounts(msc.sce))  # access log-transformed counts matrix
msc.cellLabels <- msc.sce$case.control
colnames(msc.matrix) <- msc.cellLabels

# Make a diffusion map.
msc.dm <- DiffusionMap(data = t(msc.matrix), n_pcs = 50)

saveRDS(msc.dm, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/MSC_AllCells_DiffusionMap.rds")
```
# ECs now
```{r}
ec.matrix <- as.matrix(logcounts(ec.sce))  # access log-transformed counts matrix
ec.cellLabels <- ec.sce$celltypes_ECsubbed
colnames(ec.matrix) <- ec.cellLabels

# Make a diffusion map.
ec.dm <- DiffusionMap(data = t(ec.matrix), n_pcs = 50)

saveRDS(ec.dm, "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/EC_AllCells_DiffusionMap.rds")
```

# Plot
```{r}
tmp <- data.frame(DC1 = eigenvectors(msc.dm)[, 1],
                  DC2 = eigenvectors(msc.dm)[, 2],
                  Timepoint = msc.sce$celltypes_ECsubbed,
                  Condition = msc.sce$case.control)
ggplot(tmp, aes(x = DC1, y = DC2, colour = Timepoint)) +
    geom_point() + scale_color_manual(values = c("#F8766D", "#00BA38", "#00BFC4", "#B79F00", "#619CFF")) +
    facet_grid(. ~ Condition) +
    xlab("Diffusion component 1") + 
    ylab("Diffusion component 2") +
    theme_classic()
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/MSC_DiffusionMap_ConditionSplit_Osteoclastlike.png", height = 6, width = 12)
```

# Now EC plot
```{r}
tmp <- data.frame(DC1 = eigenvectors(ec.dm)[, 1],
                  DC2 = eigenvectors(ec.dm)[, 2],
                  Timepoint = ec.sce$celltypes_ECsubbed,
                  Condition = ec.sce$case.control)
ggplot(tmp, aes(x = DC1, y = DC2, colour = Timepoint)) +
    geom_point() + scale_color_manual(values = c("#F564E3", "#C77CFF")) +
    facet_grid(. ~ Condition) +
    xlab("Diffusion component 1") + 
    ylab("Diffusion component 2") +
    theme_classic()
ggsave("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/figures/EC_DiffusionMap_ConditionSplit.png", height = 6, width = 12)
```


