---
title: "DESeq2 Modular Workflow"
author: "Alex Li and Patrick Blaney. Edited by DG on 1/31/2023"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

# DG edit notes: 

I replaced most of the table reading steps in the deseq2_exe() function. Specifically, I made it so that the table reading steps involve reading the table WITH the headers, replacing a character in the cell names for the counts matrix so that it matches the content of the conditions file, and reading in the conditions file and doing some manipulations to give it the proper format. Furthermore, I had to add a +1 pseudocount when generating the matrix originally due to the zeroes causing a downstream error in the DESeq() function. 

# Environment Setup

Load in necessary packages, set scientific notation and ggplot theme
```{r set-env, message=FALSE, warning=FALSE, results='hide'}
rm(list=ls())
library(DESeq2)
library(apeglm)
library(rtracklayer)
library(tidyverse)
library(ComplexHeatmap)
library(glue)
library(ggrepel)
library(EnhancedVolcano)
library(ggpubr)
library(paletteer)

options(scipen = 999)

# Set figure base theme
theme_set(
  theme_gray() +
  theme(axis.line = element_line(size = 0.5,  color = "black"),
        panel.background = element_rect(fill = NA, size = rel(14)),
        panel.grid.minor = element_line(color = NA),
        axis.text = element_text(size = 12,  color = "black"),
        axis.title = element_text(size = 14),
        axis.ticks = element_line(size = 0.75),
        title = element_text(size = 16),
        plot.title = element_text(hjust = 0.5)
        )
  )
```

# Function Loadout

```{r}
# Run DESeq2 on featureCounts data
deseq2_exe <- function(x_counts_input, y_counts_input,
                       x_conditions_input, y_conditions_input,
                       gencode_gene_list = "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/gencode.v42.primary_assembly.basic.annotation.gtf.gz") {
  
  x_counts <- read.table(file = x_counts_input) # deleted header = F
  colnames(x_counts) <- gsub('\\.', '-', colnames(x_counts)) # need to do this replace because for some reason the - is being replaced before this??? no idea where???

  y_counts <- read.table(file = y_counts_input)
  colnames(y_counts) <- gsub('\\.', '-', colnames(y_counts)) # need to do this replace because for some reason the - is being replaced before this??? no idea where???

  x_conditions <- read.table(x_conditions_input)
  colnames(x_conditions) <- c("sample", "condition", "type")
  rownames(x_conditions) <- x_conditions$sample
  x_conditions <- x_conditions[,-1]
  
  y_conditions <- read.table(y_conditions_input)
  colnames(y_conditions) <- c("sample", "condition", "type")
  rownames(y_conditions) <- y_conditions$sample
  y_conditions <- y_conditions[,-1]

  # Create condition comparison matrix and factorize condition matrix variables
  y_vs_x_conditions <- rbind(x_conditions, y_conditions)
  y_vs_x_conditions$condition <- factor(y_vs_x_conditions$condition,
                                        levels = c(x_conditions[1,1], y_conditions[1,1]))
  y_vs_x_conditions$type <- factor(y_vs_x_conditions$type)
  
  # Create count comparison matrix
  y_vs_x_counts <- cbind(x_counts, y_counts)

  # To refine analysis, restrict the counts matrix to CDS genes in GENCODE v42 GTF
  # and then remove any mitochondrial genes
  gencode_genes <- import(gencode_gene_list)
  gencode_genes <- unique(gencode_genes[gencode_genes$type == "CDS"]$gene_name)
  gencode_genes <- gencode_genes[!str_detect(string = gencode_genes, pattern = "^MT-")]
  
  y_vs_x_counts <- y_vs_x_counts %>%
                     dplyr::filter(rownames(y_vs_x_counts) %in% gencode_genes)
  
  # Quick first hard filter to remove genes with little expression across all samples
  # Remove all genes with mean expression < median(all_per_gene_exp_mean) / 10
  per_gene_exp_mean <- y_vs_x_counts %>%
                         dplyr::mutate(exp_mean = round(rowSums(y_vs_x_counts) / ncol(y_vs_x_counts),
                                                        digits = 0)) %>%
                         dplyr::select(exp_mean) %>%
                         purrr::as_vector() %>%
                         as.numeric()
  
  y_vs_x_counts <- y_vs_x_counts %>%
                     dplyr::filter(rowSums(y_vs_x_counts) / ncol(y_vs_x_counts) > round(median(per_gene_exp_mean) / 10,
                                                                                        digits = 0))
  
  # Convert counts matrix and conditions file into DESeq2 object
  y_vs_x_counts_deseq_obj <- DESeqDataSetFromMatrix(countData = y_vs_x_counts,
                                                    colData = y_vs_x_conditions,
                                                    design = ~ condition)
  
  # Run DESeq2
  y_vs_x_counts_deseq_obj <- DESeq(object = y_vs_x_counts_deseq_obj)

  return(y_vs_x_counts_deseq_obj)
}

# Run the LFC shrink transform on the DESeq2 obj and produce friendly tibble
deseq2_lfc_shrink_transform <- function(deseq2_obj) {
  
  # Next, utilize the log fold change shrinkage to more accurately visualize and then rank the results
  y_vs_x_lfc_results <- lfcShrink(dds = deseq2_obj,
                                  coef = 2,
                                  type = "apeglm")
  
  # Visualize the improved LFC values with annotations for top 25 genes by absolute Log2FC value
  friendly_data <- tibble("gene" = rownames(y_vs_x_lfc_results),
                          "base_mean" = round(y_vs_x_lfc_results$baseMean, digits = 2),
                          "log2_fc" = round(y_vs_x_lfc_results$log2FoldChange, digits = 4),
                          "p_val" = y_vs_x_lfc_results$pvalue,
                          "p_val_adj" = y_vs_x_lfc_results$padj)
  
  return(friendly_data)
}

# Generate volcano plot, p_val_adj < 0.05 and Log2FC > 1.0, top 30 sig. genes
deseq2_volcano_plot <- function(deseq2_df, top_n_neg_gene_labels, top_n_pos_gene_labels) {
  
  volcano_plot <- EnhancedVolcano::EnhancedVolcano(toptable = deseq2_df,
                                                   lab = deseq2_df$gene,
                                                   selectLab = c(top_n_neg_gene_labels[1:30], top_n_pos_gene_labels[1:30]),
                                                   drawConnectors = T,
                                                   maxoverlapsConnectors = 100,
                                                   labSize = 3.75,
                                                   labFace = "bold",
                                                   x = "log2_fc",
                                                   y = "p_val_adj",
                                                   pCutoff = 0.05,
                                                   FCcutoff = 1.0,
                                                   gridlines.major = F,
                                                   gridlines.minor = F,
                                                   raster = T)
  
  volcano_plot <- volcano_plot +
                    annotate(geom = "label",
                             x = c(max(volcano_plot$data$xvals, na.rm = T) - 1.5,
                                   min(volcano_plot$data$xvals, na.rm = T) + 1.5),
                             y = max(-log10(volcano_plot$data$p_val_adj), na.rm = T) + 4,
                             label = c(paste("Upregulated in", levels(diff_exp_obj$condition)[2], "samples", sep = " "),
                                       paste("Upregulated in", levels(diff_exp_obj$condition)[1], "samples", sep = " ")),
                             size = 4.75,
                             fontface = "bold")
  
  return(volcano_plot)
}


#
# Load PCA execution and visualization function originally written by Igor Dolgalev,
# minor adaptations made Run the code block to initialize the functions for use
# (<https://github.com/igordot/sns/blob/main/scripts/deseq2-pca.R>)
#

deseq2_pca <- function(deseq2_obj, ntop = 1000) {
  # run PCA
  rv <- rowVars(assay(deseq2_obj))
  select <- order(rv, decreasing = TRUE)[seq_len(min(ntop, length(rv)))]
  pca <- prcomp(t(assay(deseq2_obj)[select, ]))
  return(pca)
}

deseq2_pca_plotter <- function(deseq2_obj, deseq2_pca_obj, intgroup = "condition", point_labels = TRUE) {

  # determine the proportion of variance for the PCs
  pca_variance <- (deseq2_pca_obj$sdev ^ 2) / (sum(deseq2_pca_obj$sdev ^ 2))
  pca_variance <- round(pca_variance, 3) * 100

  if (!all(intgroup %in% names(colData(deseq2_obj)))) {
    stop("the argument 'intgroup' should specify columns of colData()")
  }

  # add the intgroup factors together to create a new grouping factor
  intgroup_df = as.data.frame(colData(deseq2_obj)[, intgroup, drop = FALSE])
  if (length(intgroup) > 1) {
    fac <- factor(apply(intgroup_df, 1, paste, collapse = " : "))
  } else {
    fac <- colData(deseq2_obj)[[intgroup]]
    # set factor levels if they do not already exist (may never happen in practice)
    if (is.null(levels(fac))) {
      levels(fac) <- unique(as.character(fac))
    }
  }

  # set color sheme based on number of groups
  if (nlevels(fac) > 20) {
    colors <- rainbow(nlevels(fac))
  } else if (nlevels(fac) > 2) {
    colors <- c(brewer.pal(5, "Set1"), brewer.pal(8, "Dark2"), brewer.pal(8, "Accent"))
    colors <- unique(colors)
    colors <- head(colors, nlevels(fac))
  } else {
    colors <- c("forestgreen", "darkorchid4")
  }

  # set text font size based on number of samples (measured in mm, not points)
  num_samples <- ncol(deseq2_obj)
  if (num_samples > 50) {
    font_size <- 1.5
  } else if (num_samples > 20) {
    font_size <- 2
  } else if (num_samples > 10) {
    font_size <- 3
  } else {
    font_size <- 4
  }

  # PCA data frame for plotting
  pca_df <- data.frame(PC1 = deseq2_pca_obj$x[,1], PC2 = deseq2_pca_obj$x[,2], group = fac, stringsAsFactors = FALSE)

  # randomize sample order
  set.seed(99)
  pca_df <- pca_df[sample(rownames(pca_df)), ]

  # plot
  p <- ggplot(pca_df, aes(x = PC1, y = PC2)) +
         geom_point(aes(color = group), size = 4) +
         labs(title = "PCA",
              x = glue("PC1 ({pca_variance[1]}% Variance)"),
              y = glue("PC2 ({pca_variance[2]}% Variance)")) +
         scale_color_manual(values = colors) +
         cowplot::theme_cowplot() +
         theme(aspect.ratio = 1,
               plot.background = element_rect(fill = "white"),
               plot.title = element_text(hjust = 0.5),
               plot.subtitle = element_text(hjust = 0.5),
               legend.title = element_blank())

  # add sample labels
  if (point_labels) {
    p <- p +
           geom_text_repel(aes(label = rownames(pca_df)),
                           size = font_size,
                           point.padding = unit(0.5, "lines"),
                           color = "black")
  }
  
  # return plot
  p
}



# Generate PCA plot of VST data
deseq2_pca_plot_wrapper <- function(untransformed_deseq2_obj) {
  # Perform variance stabilization transform to visualize via heatmap and PCA
   y_vs_x_vst <- varianceStabilizingTransformation(object = untransformed_deseq2_obj,
                                                   blind = TRUE)
   
   # Now generate PCA and visualize
   y_vs_x_vst_pca_obj <- deseq2_pca(deseq2_obj = y_vs_x_vst,
                                    ntop = 1000)
   pca_plot <- deseq2_pca_plotter(deseq2_obj = y_vs_x_vst,
                                  deseq2_pca_obj = y_vs_x_vst_pca_obj,
                                  point_labels = F)
   
   return(pca_plot)
}

# Calculate the row wide Z-score, used for heatmap of top 100 up/down regulated genes
row_wide_zscore_calc <- function(matrix) {
  row_mean_vector <- apply(matrix, 1, mean)
  row_sd_vector <- apply(matrix, 1, sd)
  
  zscores_matrix <- matrix(nrow = nrow(matrix), ncol = ncol(matrix))
  for(i in 1:nrow(matrix)) {
    
    for(j in 1:ncol(matrix)) {
      zscores_matrix[i,j] <- (matrix[i,j] - row_mean_vector[i]) / row_sd_vector[i]
    }
  }

  return(zscores_matrix)
}

# Generate heatmap with clustered dendrogram of top 100 up/down regulated genes
deseq2_heatmap_plot <- function(untransformed_deseq2_obj, top_n_neg_gene_labels, top_n_pos_gene_labels) {
  
  # Perform variance stabilization transform to visualize via heatmap and PCA
  y_vs_x_vst <- varianceStabilizingTransformation(object = untransformed_deseq2_obj,
                                                   blind = TRUE)
   
  # Find top 200 most variable genes based on log2FC
  heatmap_variable_genes <- assay(y_vs_x_vst)[rownames(assay(y_vs_x_vst)) %in% c(top_n_pos_gene_labels, top_n_neg_gene_labels),]

  heatmap_variable_genes_zscores <- row_wide_zscore_calc(matrix = heatmap_variable_genes)
  colnames(heatmap_variable_genes_zscores) <- colnames(heatmap_variable_genes)
  rownames(heatmap_variable_genes_zscores) <- rownames(heatmap_variable_genes)
   
  # Defining sample colors vector + naming for input into HeatmapAnnotation
  color_labels <- c("forestgreen","darkorchid4")
  names(color_labels) <- c(levels(diff_exp_obj$condition)[1], levels(diff_exp_obj$condition)[2])
  
  # Grab sample names to use as annotations
  group_anno <- HeatmapAnnotation("Sample Type" = diff_exp_obj$condition,
                                  col = list("Sample Type" = color_labels))

  heatmap_plot <- Heatmap(matrix = heatmap_variable_genes_zscores,
                          clustering_distance_columns = "euclidean",
                          clustering_method_columns = "average",
                          col = rev(paletteer_c("ggthemes::Orange-Blue-White Diverging", 7)),
                          column_dend_height = unit(25, "mm"),
                          column_names_gp = gpar(fontsize = 0),
                          row_dend_width = unit(25, "mm"),
                          row_names_gp = gpar(fontsize = 2.7, fontface = "bold"),
                          top_annotation = group_anno,
                          heatmap_legend_param = list(title = "Row Wide\nZ-Score"))
  
  return(heatmap_plot)
}

# Generate significance ranked gene expression list for importing to GSEA 
rank_for_gsea <- function(deseq2_df, x_counts_input, y_counts_input) {
  
  # Create ranked list of genes based on the sign of the logFC and the -log10 p-adj
  ranked_list_for_gsea <- deseq2_df %>%
                            dplyr::mutate(rank_score = sign(log2_fc) * -log10(p_val_adj))
  
  # Must be changed per writer based on desired output location
  # Parses name before .counts.txt in file name to create y_vs_x format rank list naming
  rnk_output <- paste(strsplit(tail(strsplit(y_counts_input,"/")[[1]],1),"\\.")[[1]][1],
                      "_vs_",
                      strsplit(tail(strsplit(x_counts_input,"/")[[1]],1),"\\.")[[1]][1],
                      "_gsea_data.rnk", sep = "")
  
  # Write rank list
  write.table(x = ranked_list_for_gsea %>% dplyr::select(gene, rank_score),
              file = rnk_output,
              sep = "\t",
              row.names = F,
              quote = F,
              col.names = F)
  
  return(ranked_list_for_gsea)
}
```

# Workflow Execution

Stage input files here, the output will reflect enrichment in Y versus X.
```{r}
x_counts <- "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/healthyFibroblasts_counts.txt"
y_counts <-"/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/myelomaFibroblasts_counts.txt"
  
x_conditions <- "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/healthyFibroblasts_pseudobulk_Conditions.txt"
y_conditions <- "/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/myelomaFibroblasts_pseudobulk_Conditions.txt"


output_basename <- paste(strsplit(tail(strsplit(y_counts,"/")[[1]],1),"\\.")[[1]][1],
                         "_vs_",
                         strsplit(tail(strsplit(x_counts,"/")[[1]],1),"\\.")[[1]][1],
                         sep = "")

diff_exp_obj <- deseq2_exe(x_counts_input = x_counts,
                           y_counts_input = y_counts,
                           x_conditions_input = x_conditions,
                           y_conditions_input = y_conditions)

diff_exp_transformed_df <- deseq2_lfc_shrink_transform(diff_exp_obj)

top100_pos_lfc_genes <- diff_exp_transformed_df %>%
                          dplyr::arrange(p_val_adj) %>%
                          dplyr::filter(log2_fc > 0) %>%
                          dplyr::select(gene) %>%
                          head(n = 100) %>%
                          purrr::as_vector()

top100_neg_lfc_genes <- diff_exp_transformed_df %>%
                          dplyr::arrange(p_val_adj) %>%
                          dplyr::filter(log2_fc < 0) %>%
                          dplyr::select(gene) %>%
                          head(n = 100) %>%
                          purrr::as_vector()

vplot <- deseq2_volcano_plot(diff_exp_transformed_df, top100_neg_lfc_genes, top100_pos_lfc_genes)
ggsave(filename = paste0("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/", output_basename, ".volcano.pdf"),
       plot = vplot,
       scale = 0.85,
       width = 35,
       height = 28,
       units = "cm")

pplot <- deseq2_pca_plot_wrapper(untransformed_deseq2_obj = diff_exp_obj)
ggsave(filename = paste0("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/", output_basename, ".pca.pdf"),
       plot = pplot,
       scale = 0.85,
       width = 15,
       height = 13,
       units = "cm")

pdf(file = paste0(output_basename, ".heatmap.pdf"))
hplot <- deseq2_heatmap_plot(untransformed_deseq2_obj = diff_exp_obj, top100_neg_lfc_genes, top100_pos_lfc_genes)
draw(hplot)

diff_exp_transformed_ranked_df <- rank_for_gsea(diff_exp_transformed_df,
                                                x_counts_input = x_counts, y_counts_input = y_counts)
write.table(x = diff_exp_transformed_ranked_df,
            file = paste0("/Users/gagled01/morganLab/single-cell/5TG_Mouse_Rerun/", output_basename, ".diffexp.txt"),
            sep = "\t",
            row.names = F,
            quote = F,
            col.names = T)
```

